### 2020-06-17 output_scripts/Null_Windows.Rmd

---
title: "Null Window Correction"
author: "Owen Ward"
date: "6/17/2020"
output: html_document
---

This document briefly reviews the corrections for including empty time
windows in both fitting and evaluating the Hawkes process type models.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/owenw/Documents/Github_Local/MMHP_Latent/")
```

```{r setup_data, include=FALSE,message=FALSE,echo=FALSE}

# specify the path where the data is saved locally
save_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"

no_segments <- 500

library(rstan)
options(mc.cores = parallel::detectCores())

library(compete)
#library(RColorBrewer)
#library(Hmisc)
#library(wCorr)
#library(tidyverse)
library(dplyr)
#library(R.utils)
#library(fields)

source('lib/naiveRankHierarchy.R')
source('lib/expertRankHierarchy.R')
source('lib/cleanData.R') 
source('lib/prepareDataStan.R') 
source('lib/inferLatentMmhp.R')
source('lib/plotUtil.R') 
source('lib/mmhp.R')
source('lib/uniHawkes.R')
source('lib/simulatePrediction.R')
source('lib/myGlicko.R')
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")
source('lib/matrixPlotParameter.R')
source('lib/residualStructureScore.R')

#### load the data ####
full_data <- readRDS("data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

model1_fn <- list(alpha.fun = function(x,y,eta1,eta2,eta3){return(eta1*x*y*exp(-eta2*abs(x-y))/(1+exp(-eta3*(x-y))))})

model3_fn <- list(alpha.fun = function(x,y,eta1,eta2){return(eta1*x*y*exp(-eta2*abs(x-y)))},
                  q1.fun = function(x,y,eta3){return(exp(-eta3*x))},
                  q0.fun = function(x,y,eta3){return(exp(-eta3*y))})

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <- expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
}

```

## Compare Pearson residual matrices

```{r,include=FALSE,echo=FALSE}
current_cohort <- 9
m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

```

```{r, include=FALSE,echo=FALSE}
range(m1_pr)
range(m2_pr)
range(m3_pr)
range(indep_pr)
```

Then we can plot each of these, for each cohort.

```{r}
#current_cohort <- 1
cohorts <- 1:10
cohorts <- cohorts[-3]
for(current_cohort in cohorts){
  cat("Cohort",current_cohort,"\n",sep = " ")
  m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))
matrix_lst_plot <- list(
                        indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                        m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]]) 
myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                  xLabels=expert_rank_10[[current_cohort]],
                  yLabels=rev(expert_rank_10[[current_cohort]]),
                  min=-150,max=150,axis_cex=2,title_cex = 1.8,
                  colorPalette="RdBu",if.emp=FALSE,
                  #legend.mar=c(0.5,.5,0.5,0.5),
                  title_lst=list("I-MMHP","C-HP","C-DCHP","C-MMHP"), 
                  #title_lst = list("I-MMHP"),
                  col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                  matrix.mar=c(2.5,2.5,2.5,1)
                  )

}

```

Here pairs with no colour (white squares)
indicate values larger than the limits chosen. One cohort is
missing due to not fitting on the cluster in time.
The independent MMHP model now appears to be slightly better than the C-MMHP but the prediction
results shouldn't change, so overall this appears reasonable.
Also, this is using the old priors which can likely lead to a better fit.

## Predictions

For a single cohort now.

```{r}
current_cohort <- 1
load(paste(data_path,cohort_names[current_cohort],
                  "/predict_simulation_",cohort_names[current_cohort],
                  ".RData",sep=''))
```