### 2020-06-22 output_scripts/New_Priors.Rmd
improved examples

---
title: "Priors for Latent Hawkes"
author: "Owen G. Ward"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior/"

library(rstan)
```

```{r mice_setup,include=FALSE,echo=FALSE}
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12
```

# First Prior

To remove some of the issues with the previously specified priors,
we reparameterise `lambda1` and choose more appropriate priors
for `w_lambda,lambda0` and `beta`.

```{r eval=FALSE}

lambda1 = lambda0*(1+w_lambda);
// now w_lambda only has to be positive

//priors
lambda0 ~ gamma(0.1,1); // prior has mean 0.1, var 0.1
w_lambda ~ lognormal(0,2);
eta_1 ~ lognormal(0,1);
eta_2 ~ lognormal(0,1);
eta_3 ~ lognormal(0,1);
beta ~ lognormal(0,2);

```

## Examine `stan` diagnostics

```{r}
fit_cohorts <- 1:10
fit_cohorts <- fit_cohorts[-7]
for(current_cohort in fit_cohorts){
  print(current_cohort)
  load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
  rstan::check_hmc_diagnostics(fit_cohort_mmhp)
}

```

This leads to better results in terms of divergent transitions and
max tree depth, which were both issues previously.

## Param Histograms

We can check the traceplots/posterior distributions
based on these prior choices.

```{r}
current_cohort <- 1
print(current_cohort)
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Priors")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Priors")

```

We can similarly look at the original fits (with and without considering the windows with 
no events).

```{r}

orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/"

new_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/debug_copy/"

load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("Original, no null windows")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("Original, no null windows")

fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_fit_",cohort_names[current_cohort],
         ".RDS",sep=''))

rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("Original, null windows")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("Original, null windows")

```

## Latent Rankings

```{r}
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Priors")

p1

```

We can compare this to the original priors (both with and excluding the null windows).

```{r}
load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
p2 <- rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Original, no null windows")

fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_fit_",cohort_names[current_cohort],
         ".RDS",sep=''))

p3 <- rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Original, null windows")

p2
p3

```

```{r,fig.height=9}
ggpubr::ggarrange(p1,p2,p3,nrow=3)
```

These seem similar either way, and maybe tighter with these new priors.

# Lower bound on $\lambda_0$

We run into some issues with this where `w_lambda` can explode,
likely caused by extremely small draws for `lambda0`. To circumvent this
we place a lower bound on `lambda0` of 0.01 and refit this model.

```{r}
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"

print(current_cohort)
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Priors")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Priors")

```

This avoids the previous issue with parameters blowing up however one chain
seems to be stuck.

```{r}
  print(current_cohort)
  load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
  rstan::check_hmc_diagnostics(fit_cohort_mmhp)
```
```{r}
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound on lambda")

p1_new
```

We can compare this to the previous fit.

These fits appear to be slightly tighter overall for some cohorts, but slightly
larger for others.

```{r, fig.height=10}
print(current_cohort)
ggpubr::ggarrange(p1,p1_new,nrow=2)
```

We can repeat this for another cohort.

```{r, fig.height=10}
current_cohort <- 2

print(current_cohort)
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior")

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound lambda0")

ggpubr::ggarrange(p1,p1_new,nrow=2)

```