### 2020-06-22 output_scripts/mod_cmmhp.Rmd
correction so that plot shown

---
title: "Exploring C-MMHP"
author: "Owen Ward"
output:
  html_document:
    toc: yes
    df_print: paged
    toc_float: yes
    code_folding: hide
    number_sections: true
---

<!-- The aim of this notebook is to explore some properties of the current -->
<!-- C-MMHP model to ensure it captures the correct social dynamics present -->
<!-- in animal interactions. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height = 4,fig.width = 5)

```

```{r load_mice_data and packages, include=FALSE, echo=FALSE}
library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())

func_dir <- paste(here::here(),"/lib/",sep = "")

output_dir <- paste(here::here(),"/output/",sep = "")

orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/"

library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

theme_set(theme_minimal())

full_data <- readRDS(paste(here::here(),"/data/mice.RData",sep = ""))
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <- expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
})

```

# Plot rankings against degree corrected rates

The first goal is to plot the latent ranks against the corresponding
degree corrected rate estimates for the `C-DCHP`.

```{r load_dchp,include=FALSE,echo=FALSE}
current_cohort <- 3

load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_dchp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

#fit_cohort_dchp
#sim_cohort_dchp

#dim(sim_cohort_dchp$f)
mean_rank <- apply(sim_cohort_dchp$f,2,mean)

mean_gamma <- apply(sim_cohort_dchp$gamma,2,mean)
mean_zeta <- apply(sim_cohort_dchp$zeta, 2, mean)

mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                      isi = expert_rank_10[[current_cohort]])

# mean_values %>%
#   rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
#   pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
#   #rename(in_out = name) %>%
#   ggplot(aes(mean_rank,value,color = out_in)) + geom_point()  +
#   labs(title = "") + theme(legend.title = element_blank())
# 
# 
# 
# mean_values %>%
#   rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
#   pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
#   ggplot(aes(isi,value,color =  out_in))  + geom_point()  +
#   theme(legend.title = element_blank()) +
#   scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking")

mean_values %>%
  rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
  pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
  mutate(out_in = as.factor(out_in)) %>%
  ggplot(aes(isi,value))  + geom_point()  +
  theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
  scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking")

```

```{r}
for(current_cohort in 1:10){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_dchp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank <- apply(sim_cohort_dchp$f,2,mean)

  mean_gamma <- apply(sim_cohort_dchp$gamma,2,mean)
  mean_zeta <- apply(sim_cohort_dchp$zeta, 2, mean)

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,value))  + geom_point()  +
    theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

}
```

This does not appear to indicate any clear relationship between these degree
corrected estimates for the out/in baseline rates and the expert
ranking.

# The Limiting State

## Current Model

We wish to better understand the properties captured by the current form
of the C-MMHP model.

Currently, we have that for each pair $i,j$, $q_1^{i,j}/q_{0}^{i,j}$ is the probability 
that pair transitions out of the active/inactive state. In particular,

$$
q_1^{i,j} = \exp(-\eta_3 f_i)\hspace{1cm}   \ q_{0}^{i,j} = \exp(-\eta_3 f_j),
$$
and the limiting proportion of 
time spent in the active state is $logistic(\eta_3 (f_i-f_j))$,
i.e

$$
\frac{q_{0}^{i,j}}{q_{0}^{i,j}+q_1^{i,j}}  \equiv \frac{\exp(-\eta_3 f_j)}{\exp(-\eta_3 f_j)+\exp(-\eta_3 f_i)}.
$$

To illustrate this, lets look at the this for a $f_i$.

```{r, include=FALSE,echo=FALSE,eval=FALSE}
f_i <- 0.1
f_x <- function(x) {
  (exp(-eta_3*x))/( exp(-eta_3*x) + exp(-eta_3*f_i) )
}

curve(f_x,from = 0,to = 1)
f_i

```

```{r,fig.show="hold", out.width="50%"}
f_curr <- function(x,f_i,eta_3) {
  (exp(-eta_3*x))/( exp(-eta_3*x) + exp(-eta_3*f_i) )
}

plot_limit <- function(f_i,eta_3) {
  plt <- ggplot(data.frame(x = c(0,1)),aes(x)) + 
    stat_function(fun = f_curr,
                  args = list(f_i = f_i, eta_3 = eta_3)) + 
    ylab("Limiting Proportion Active") +
    xlab("f_j (Receiver Rank)") + ggtitle(paste("f_i (Send Rank) of ",f_i,sep = ""))
  print(plt)
}

plot_limit(f_i = 0.9,eta_3 = 3)
plot_limit(f_i = 0.1,eta_3 = 3)

```

This indicates a high probability of being in the active state when the attacking mouse
is highly ranked. He is likely to fight with all mice.

When the attacking mouse is poorly ranked, there is quite a high probability he
will attack mice around him, but this is unlikely for mice ranked much higher.

## Alternative Model

We consider some alternative formulations for this limiting behaviour.

One such modification is
$$
q_1^{i,j} = f_i\exp(-\eta_3 f_i) \hspace{1cm}   \ q_{0}^{i,j} = f_j\exp(-\eta_3 f_j),
$$

```{r,fig.show="hold", out.width="50%"}
f_mod <- function(x,f_i,eta_3) {
  (x*exp(-eta_3*x))/( x*exp(-eta_3*x) + f_i*exp(-eta_3*f_i) )
}

plot_limit_1 <- function(f_i,eta_3) {
  plt <- ggplot(data.frame(x = c(0,1)),aes(x)) + 
    stat_function(fun = f_mod,
                  args = list(f_i = f_i, eta_3 = eta_3)) + 
    ylab("Limiting Proportion Active") +
    xlab("f_j (Receiver Rank)") + ggtitle(paste("f_i (Send Rank) of ",f_i,sep = ""))
  print(plt)
}

plot_limit_1(f_i = 0.9,eta_3 = 3)
plot_limit_1(f_i = 0.1,eta_3 = 3)

```

This does not seem appropriate, in that a dominant mouse is very unlikely 
to fight a very weak mouse. Similarly, a weak mouse is more likely to fight a strong 
mouse than an even weaker one than itself.

## Alternative 2

Another alternative is

$$
q_1^{i,j} = f_j\exp(-\eta_3 f_i) \hspace{1cm}   \ q_{0}^{i,j} = f_i\exp(-\eta_3 f_j),
$$

```{r, fig.show="hold", out.width="50%"}
f_mod_2 <- function(x,f_i,eta_3) {
  (f_i*exp(-eta_3*x))/( f_i*exp(-eta_3*x) + x*exp(-eta_3*f_i) )
}

plot_limit_2 <- function(f_i,eta_3) {
  plt <- ggplot(data.frame(x = c(0,1)),aes(x)) + 
    stat_function(fun = f_mod_2,
                  args = list(f_i = f_i, eta_3 = eta_3)) + 
    ylab("Limiting Proportion Active") +
    xlab("f_j (Receiver Rank)") + ggtitle(paste("f_i (Send Rank) of ",f_i,sep = ""))
  print(plt)
}

plot_limit_2(f_i = 0.9,eta_3 = 3)
plot_limit_2(f_i = 0.1,eta_3 = 3)

```

This looks more plausible, but is perhaps similar
(if more complex) to the first model. Here the limiting state of a weak mouse fighting
a stronger one is more likely to be inactive, which seems to be
an improvement.

# The excitation parameter

The excitation parameter is currently defined by

$$
\alpha^{i,j} = \eta_1 f_i f_j \exp(-\eta_2 |f_i-f_j|)
$$

We can look at a contour plot of this for a range of values of $f_i,f_j$,
as is currently in the paper.

```{r, contour_alpha}
f_alpha <- function(f_i,f_j,eta_1,eta_2) {
  eta_1*f_i*f_j * exp(-eta_2*abs(f_i-f_j))
}
plot_alpha <- function(fun = f_alpha, grid_size = 0.1,eta_1 = 1,eta_2 = 1, ...) {
  plt <- tibble(x = seq(0,1,by = grid_size),y = seq(0,1,by = grid_size)) %>%
    expand(x,y) %>%
    mutate(value = map2_dbl(x,y,fun,eta_1,eta_2,...)) %>%
    ggplot(aes(y,x)) + geom_contour_filled(aes(z = value))  +
    scale_x_reverse()+ 
    xlab("f_j (Receive)") + ylab("f_i (Send)") + geom_abline(slope = -1, intercept = 0,
                                                             col = "red")
  print(plt)
}

plot_alpha(f_alpha,eta_1 = 3, eta_2 = 1,grid_size = 0.01)
```

This appears to capture desirable properties present in animal interactions.

We want this to take larger values when $f_i>f_j$ than when $f_i < f_j$ although it is 
hard to claim it is not symmetric currently. This could be due to using a different
plot function than originally used by Jing.

## Incorporate the latent state

We can also plot $\alpha Z(t)$, which will capture the overall structure
which will be seen.

```{r}

latent_state <- function(f_i,f_j,eta_3) {
  exp(-eta_3*f_j)/(exp(-eta_3*f_j) + exp(-eta_3*f_i)  )
}

f_alpha_z <- function(f_i,f_j,eta_1,eta_2,eta_3) {
  eta_1*f_i*f_j * exp(-eta_2*abs(f_i-f_j)) * latent_state(f_i,f_j,eta_3)
}

plot_alpha(fun = f_alpha_z,eta_1 = 3, eta_2 = 1,grid_size = 0.01,eta_3 = 4)

```

This much more clearly captures the upper triangular structure we wish to have.

## With modified latent state

We can repeat this with the second proposed modification above, which seemed 
to capture the dynamics well.

```{r}

latent_state_2 <- function(f_i,f_j,eta_3) {
  f_i*exp(-eta_3*f_j)/(f_i*exp(-eta_3*f_j) + f_j*exp(-eta_3*f_i)  )
}

f_alpha_z <- function(f_i,f_j,eta_1,eta_2,eta_3) {
  eta_1*f_i*f_j * exp(-eta_2*abs(f_i-f_j)) * latent_state_2(f_i,f_j,eta_3)
}

plot_alpha(fun = f_alpha_z,eta_1 = 3, eta_2 = 1,grid_size = 0.01,eta_3 = 4)

```

This is almost identical to the previous plot, with perhaps a slight shift when
both $f_i$ and $f_j$ are large.