### 2020-06-24 output_scripts/New_Priors.Rmd
lower bound for lambda0 in dc fit also

---
title: "Priors for Latent Hawkes"
author: "Owen G. Ward"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior/"

func_dir <- "../lib/"

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

theme_set(theme_minimal())
```

```{r mice_setup,include=FALSE,echo=FALSE}
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <- expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
}

```

<!-- # First Prior -->

<!-- To remove some of the issues with the previously specified priors, -->
<!-- we reparameterise `lambda1` and choose more appropriate priors -->
<!-- for `w_lambda,lambda0` and `beta`. -->

<!-- ```{r eval=FALSE} -->

<!-- lambda1 = lambda0*(1+w_lambda); -->
<!-- // now w_lambda only has to be positive -->

<!-- //priors -->
<!-- lambda0 ~ gamma(0.1,1); // prior has mean 0.1, var 0.1 -->
<!-- w_lambda ~ lognormal(0,2); -->
<!-- eta_1 ~ lognormal(0,1); -->
<!-- eta_2 ~ lognormal(0,1); -->
<!-- eta_3 ~ lognormal(0,1); -->
<!-- beta ~ lognormal(0,2); -->

<!-- ``` -->

<!-- ## Examine `stan` diagnostics -->

<!-- ```{r} -->
<!-- fit_cohorts <- 1:10 -->
<!-- fit_cohorts <- fit_cohorts[-7] -->
<!-- for(current_cohort in fit_cohorts){ -->
<!--   print(current_cohort) -->
<!--   load(paste(data_path,cohort_names[current_cohort], -->
<!--            "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--            ".RData",sep='')) -->
<!--   rstan::check_hmc_diagnostics(fit_cohort_mmhp) -->
<!-- } -->

<!-- ``` -->

<!-- This leads to better results in terms of divergent transitions and -->
<!-- max tree depth, which were both issues previously. -->

<!-- ## Param Histograms -->

<!-- We can check the traceplots/posterior distributions -->
<!-- based on these prior choices. -->

<!-- ```{r} -->
<!-- current_cohort <- 1 -->
<!-- print(current_cohort) -->
<!-- load(paste(data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->
<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("New Priors") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("New Priors") -->

<!-- ``` -->

<!-- We can similarly look at the original fits (with and without considering the windows with  -->
<!-- no events). -->

<!-- ```{r} -->

<!-- orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/" -->

<!-- new_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/debug_copy/" -->

<!-- load(paste(orig_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->

<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("Original, no null windows") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("Original, no null windows") -->

<!-- fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_fit_",cohort_names[current_cohort], -->
<!--          ".RDS",sep='')) -->

<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("Original, null windows") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("Original, null windows") -->

<!-- ``` -->

<!-- ## Latent Rankings -->

```{r, include=FALSE}
current_cohort <- 1
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) +
  ggtitle("New Priors")

p1

```

<!-- We can compare this to the original priors (both with and excluding the null windows). -->

<!-- ```{r} -->
<!-- load(paste(orig_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->
<!-- p2 <- rstan::traceplot(fit_cohort_mmhp, -->
<!--                  pars = c("f")) +  -->
<!--   ggtitle("Original, no null windows") -->

<!-- fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_fit_",cohort_names[current_cohort], -->
<!--          ".RDS",sep='')) -->

<!-- p3 <- rstan::traceplot(fit_cohort_mmhp, -->
<!--                  pars = c("f")) +  -->
<!--   ggtitle("Original, null windows") -->

<!-- p2 -->
<!-- p3 -->

<!-- ``` -->

<!-- ```{r,fig.height=9} -->
<!-- ggpubr::ggarrange(p1,p2,p3,nrow=3) -->
<!-- ``` -->

<!-- These seem similar either way, and maybe tighter with these new priors. -->

# Lower bound on $\lambda_0$

We have modified the current priors so that 
$\lambda_1 = \lambda_0(1+w_{\lambda})$ with lognormal priors for
all parameters. Now $w_{\lambda}>0$ as opposed to $>1$ previously.

We ran into some issues with this where `w_lambda` can explode,
likely caused by extremely small draws for `lambda0`. To circumvent this
we place a lower bound on `lambda0` of 0.01 and refit this model.

The resulting posterior draws are shown below for a single cohort.

```{r}
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"

print(current_cohort)
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Priors")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Priors")

```

This avoids the previous issue with parameters blowing up however one chain
seems to be stuck. This doesn't occur for other cohorts, and is likely not
a problem.

```{r}
  print(current_cohort)
  load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
  rstan::check_hmc_diagnostics(fit_cohort_mmhp)
```

Similarly, we can look at the inferred latent ranks.

```{r}
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound on lambda")

p1_new
```

We can compare this to the previous fit (without this lower bound).

These fits appear to be slightly tighter overall for some cohorts, but slightly
larger for others.

```{r, fig.height=10}
print(current_cohort)
ggpubr::ggarrange(p1,p1_new,nrow=2)
```

We can repeat this for another cohort.

```{r, fig.height=10}
current_cohort <- 2

print(current_cohort)
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior")

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound lambda0")

ggpubr::ggarrange(p1,p1_new,nrow=2)

```

# Degree correction for C-MMHP

We instead fit a degree corrected C-MMHP. Here we have

$$
\lambda_{0}^{i,j} = \gamma_i + \zeta_j
$$

for each pair. We use Laplacian priors on these (used in DC-HP). We
can explore the resulting fits.

This seems to avoid the issue with $w_{\lambda}$ taking extremely large values
also.

```{r}
current_cohort <- 1

dc_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior/"

load(paste(dc_data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))

rstan::check_hmc_diagnostics(fit_cohort_mmhp)

rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Degree Corrected")

rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Degree Corrected")

#bayesplot::rhat(fit_cohort_mmhp)

rstan::traceplot(fit_cohort_mmhp,
               pars = c("f")) + 
  ggtitle("New Degree Corrected")

```

## Compare posterior ranks

Here we plot the latent ranks in the previous section against those
obtained with the degree correction.

```{r, fig.height=10}
current_cohort <- 1
print(current_cohort)
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Bound lambda0")

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Degree Corrected")

orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/"

load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
p2 <- rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) +
  ggtitle("Original, no null windows")

ggpubr::ggarrange(p1,p1_new,p2,nrow=3)
```

## Plot ranks against degree parameters

As before, we can also plot these inferred degree corrected rates against
the I&SI rankings.

```{r}
cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank <- apply(sim_cohort_mmhp$f,2,mean)

  mean_gamma <- apply(sim_cohort_mmhp$gamma,2,mean)
  mean_zeta <- apply(sim_cohort_mmhp$zeta, 2, mean)

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,value))  + geom_point()  +
    theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

}
```

This seems to indicate that the degree correction is somewhat useful for 
certain mice in providing a more expressive fit.

I'm running this through for the rest of the inference (predictions,
redisuals) at the moment.

# Plot ranks from DC model against I&SI

```{r}
cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank <- apply(sim_cohort_mmhp$f,2,mean)

  mean_gamma <- apply(sim_cohort_mmhp$gamma,2,mean)
  mean_zeta <- apply(sim_cohort_mmhp$zeta, 2, mean)

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    #rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    #pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    #mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,mean_rank))  + geom_point()  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

}

```

## Compare these to ranks from original fit

```{r}

cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank_dc <- apply(sim_cohort_mmhp$f,2,mean)
  load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank_orig <- apply(sim_cohort_mmhp$f, 2, mean)
  mean_values <- tibble(mean_rank_dc, mean_rank_orig,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    pivot_longer(cols = c(mean_rank_dc,mean_rank_orig),names_to = "model") %>%
    mutate(model = as.factor(model)) %>%  
    ggplot(aes(isi,value))  + 
    geom_jitter(aes(color=model), alpha = 0.5,width = 0.1)  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

}

```

# Pearson Residuals for DC C-MMHP

We wish to examine the effect the degree correction has on model fit in 
terms of Pearson residuals.

```{r}
save_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"
# the path for the result with fixes to account for the null windows

dc_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior/"

#cohorts

for(current_cohort in cohorts[-3]) {
  m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

matrix_lst_plot <- list(
                        indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                        m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]]) 
myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                  xLabels=expert_rank_10[[current_cohort]],
                  yLabels=rev(expert_rank_10[[current_cohort]]),
                  min=-150,max=150,axis_cex=2,title_cex = 1.8,
                  colorPalette="RdBu",if.emp=FALSE,
                  #legend.mar=c(0.5,.5,0.5,0.5),
                  title_lst=list("I-MMHP","C-HP","C-DCHP","DC C-MMHP"), 
                  #title_lst = list("I-MMHP"),
                  col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                  matrix.mar=c(2.5,2.5,2.5,1)
                  )
}

```

This does not appear any better than the standard C-MMHP.

```{r}
save_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"

for(current_cohort in cohorts[-3]) {
  m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  m3_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                       "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))
  indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  matrix_lst_plot <- list( indep_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                           m2_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
    m3_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                          m3_pr_dc[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]]
                          ) 
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-150,max=150,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("I-MMHP","DC-HP","C-MMHP","DC-C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                    matrix.mar=c(2.5,1.5,2.5,1)
                    )
}
```