### 2020-06-25 output_scripts/New_Priors.Rmd
exploring fits

---
title: "Priors for Latent Hawkes"
author: "Owen G. Ward"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior/"

func_dir <- "../lib/"

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

theme_set(theme_minimal())
```

```{r mice_setup,include=FALSE,echo=FALSE}
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <- expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
}

```

<!-- # First Prior -->

<!-- To remove some of the issues with the previously specified priors, -->
<!-- we reparameterise `lambda1` and choose more appropriate priors -->
<!-- for `w_lambda,lambda0` and `beta`. -->

<!-- ```{r eval=FALSE} -->

<!-- lambda1 = lambda0*(1+w_lambda); -->
<!-- // now w_lambda only has to be positive -->

<!-- //priors -->
<!-- lambda0 ~ gamma(0.1,1); // prior has mean 0.1, var 0.1 -->
<!-- w_lambda ~ lognormal(0,2); -->
<!-- eta_1 ~ lognormal(0,1); -->
<!-- eta_2 ~ lognormal(0,1); -->
<!-- eta_3 ~ lognormal(0,1); -->
<!-- beta ~ lognormal(0,2); -->

<!-- ``` -->

<!-- ## Examine `stan` diagnostics -->

<!-- ```{r} -->
<!-- fit_cohorts <- 1:10 -->
<!-- fit_cohorts <- fit_cohorts[-7] -->
<!-- for(current_cohort in fit_cohorts){ -->
<!--   print(current_cohort) -->
<!--   load(paste(data_path,cohort_names[current_cohort], -->
<!--            "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--            ".RData",sep='')) -->
<!--   rstan::check_hmc_diagnostics(fit_cohort_mmhp) -->
<!-- } -->

<!-- ``` -->

<!-- This leads to better results in terms of divergent transitions and -->
<!-- max tree depth, which were both issues previously. -->

<!-- ## Param Histograms -->

<!-- We can check the traceplots/posterior distributions -->
<!-- based on these prior choices. -->

<!-- ```{r} -->
<!-- current_cohort <- 1 -->
<!-- print(current_cohort) -->
<!-- load(paste(data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->
<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("New Priors") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("New Priors") -->

<!-- ``` -->

<!-- We can similarly look at the original fits (with and without considering the windows with  -->
<!-- no events). -->

<!-- ```{r} -->

<!-- orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/" -->

<!-- new_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/debug_copy/" -->

<!-- load(paste(orig_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->

<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("Original, no null windows") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("Original, no null windows") -->

<!-- fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_fit_",cohort_names[current_cohort], -->
<!--          ".RDS",sep='')) -->

<!-- rstan::stan_hist(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) + -->
<!--   ggtitle("Original, null windows") -->
<!-- rstan::traceplot(fit_cohort_mmhp, -->
<!--                pars = c("lambda0","lambda1","w_lambda", -->
<!--                         "eta_1","eta_2","eta_3","beta")) +  -->
<!--   ggtitle("Original, null windows") -->

<!-- ``` -->

<!-- ## Latent Rankings -->

```{r, include=FALSE}
current_cohort <- 1
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) +
  ggtitle("New Priors")

p1

```

<!-- We can compare this to the original priors (both with and excluding the null windows). -->

<!-- ```{r} -->
<!-- load(paste(orig_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_result_",cohort_names[current_cohort], -->
<!--          ".RData",sep='')) -->
<!-- p2 <- rstan::traceplot(fit_cohort_mmhp, -->
<!--                  pars = c("f")) +  -->
<!--   ggtitle("Original, no null windows") -->

<!-- fit_cohort_mmhp <- readRDS(paste(new_data_path,cohort_names[current_cohort], -->
<!--          "/cohort_mmhp_stan_fit_",cohort_names[current_cohort], -->
<!--          ".RDS",sep='')) -->

<!-- p3 <- rstan::traceplot(fit_cohort_mmhp, -->
<!--                  pars = c("f")) +  -->
<!--   ggtitle("Original, null windows") -->

<!-- p2 -->
<!-- p3 -->

<!-- ``` -->

<!-- ```{r,fig.height=9} -->
<!-- ggpubr::ggarrange(p1,p2,p3,nrow=3) -->
<!-- ``` -->

<!-- These seem similar either way, and maybe tighter with these new priors. -->

# Lower bound on $\lambda_0$

We have modified the current priors so that 
$\lambda_1 = \lambda_0(1+w_{\lambda})$ with lognormal priors for
all parameters. Now $w_{\lambda}>0$ as opposed to $>1$ previously.

We ran into some issues with this where `w_lambda` can explode,
likely caused by extremely small draws for `lambda0`. To circumvent this
we place a lower bound on `lambda0` of 0.01 and refit this model.

The resulting posterior draws are shown below for a single cohort.

```{r}
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"

print(current_cohort)
load(paste(data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Priors")
rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Priors")

```

This avoids the previous issue with parameters blowing up however one chain
seems to be stuck. This doesn't occur for other cohorts, and is likely not
a problem.

```{r}
  print(current_cohort)
  load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
  rstan::check_hmc_diagnostics(fit_cohort_mmhp)
```

Similarly, we can look at the inferred latent ranks.

```{r}
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound on lambda")

p1_new
```

We can compare this to the previous fit (without this lower bound).

These fits appear to be slightly tighter overall for some cohorts, but slightly
larger for others.

```{r, fig.height=10}
print(current_cohort)
ggpubr::ggarrange(p1,p1_new,nrow=2)
```

We can repeat this for another cohort.

```{r, fig.height=10}
current_cohort <- 2

print(current_cohort)
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior")

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1/"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("New Prior, lower bound lambda0")

ggpubr::ggarrange(p1,p1_new,nrow=2)

```

# Degree correction for C-MMHP

We instead fit a degree corrected C-MMHP. Here we have

$$
\lambda_{0}^{i,j} = \gamma_i + \zeta_j
$$

for each pair. We use Laplacian priors on these (used in DC-HP). We
can explore the resulting fits.

This seems to avoid the issue with $w_{\lambda}$ taking extremely large values
also.

```{r}
current_cohort <- 1

dc_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior/"

load(paste(dc_data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))

rstan::check_hmc_diagnostics(fit_cohort_mmhp)

rstan::stan_hist(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) +
  ggtitle("New Degree Corrected")

rstan::traceplot(fit_cohort_mmhp,
               pars = c("lambda0","lambda1","w_lambda",
                        "eta_1","eta_2","eta_3","beta")) + 
  ggtitle("New Degree Corrected")

#bayesplot::rhat(fit_cohort_mmhp)

rstan::traceplot(fit_cohort_mmhp,
               pars = c("f")) + 
  ggtitle("New Degree Corrected")

```

## Compare posterior ranks

Here we plot the latent ranks in the previous section against those
obtained with the degree correction.

```{r, fig.height=10}
current_cohort <- 1
print(current_cohort)
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_p1//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1 <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Bound lambda0")

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior//"
load(paste(data_path,cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
p1_new <-rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) + 
  ggtitle("Degree Corrected")

orig_data_path <- "C:/Users/owenw/Google Drive/Tian/Research+JingWu/thesis_data/part2/real_data/"

load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))
p2 <- rstan::traceplot(fit_cohort_mmhp,
                 pars = c("f")) +
  ggtitle("Original, no null windows")

ggpubr::ggarrange(p1,p1_new,p2,nrow=3)
```

## Plot ranks against degree parameters

As before, we can also plot these inferred degree corrected rates against
the I&SI rankings.

```{r}
cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank <- apply(sim_cohort_mmhp$f,2,mean)

  mean_gamma <- apply(sim_cohort_mmhp$gamma,2,mean)
  mean_zeta <- apply(sim_cohort_mmhp$zeta, 2, mean)

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,value))  + geom_point()  +
    theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

}
```

This seems to indicate that the degree correction is somewhat useful for 
certain mice in providing a more expressive fit.

I'm running this through for the rest of the inference (predictions,
redisuals) at the moment.

# Plot ranks from DC model against I&SI

```{r}
cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank <- apply(sim_cohort_mmhp$f,2,mean)

  mean_gamma <- apply(sim_cohort_mmhp$gamma,2,mean)
  mean_zeta <- apply(sim_cohort_mmhp$zeta, 2, mean)

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    #rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    #pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    #mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,mean_rank))  + geom_point()  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

}

```

## Compare these to ranks from original fit

```{r}

cohorts <- 1:10
#cohorts <- cohorts[c(-3,-6,-7)]
for(current_cohort in cohorts){
  #cat(paste("Cohort ",current_cohort,sep = ""))
  load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank_dc <- apply(sim_cohort_mmhp$f,2,mean)
  load(paste(orig_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

  mean_rank_orig <- apply(sim_cohort_mmhp$f, 2, mean)
  mean_values <- tibble(mean_rank_dc, mean_rank_orig,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    pivot_longer(cols = c(mean_rank_dc,mean_rank_orig),names_to = "model") %>%
    mutate(model = as.factor(model)) %>%  
    ggplot(aes(isi,value))  + 
    geom_jitter(aes(color=model), alpha = 0.5,width = 0.1)  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

}

```

# Pearson Residuals for DC C-MMHP

We wish to examine the effect the degree correction has on model fit in 
terms of Pearson residuals.

```{r}
save_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"
# the path for the result with fixes to account for the null windows

dc_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_dc_prior/"

#cohorts

for(current_cohort in cohorts[-3]) {
  m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                       "/dchp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  m3_pr <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                       "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                       "/immhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  matrix_lst_plot <- list(
                          indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                     expert_rank_10[[current_cohort]]],
                          m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                          m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                          m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]]) 
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-150,max=150,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("I-MMHP","C-HP","C-DCHP","DC C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                    matrix.mar=c(2.5,2.5,2.5,1)
                    )
}

```

This does not appear any better than the standard C-MMHP.

```{r}
save_data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"

for(current_cohort in cohorts[-3]) {
  m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  m3_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                       "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))
  m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  matrix_lst_plot <- list( m1_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                           m2_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
    m3_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                          m3_pr_dc[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]]
                          ) 
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-150,max=150,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("C-HP","DC-HP","C-MMHP","DC-C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                    matrix.mar=c(2.5,1.5,2.5,1)
                    )
}
```

# Investigate the Pearson Residuals further

```{r}
current_cohort <- 1

m2_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                   "/dchp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))

m3_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                   "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))
m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                   "/chp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))

max(m3_pr_dc)

which(m3_pr_dc == max(m3_pr_dc),arr.ind = T)

i <- 10
j <- 1

```
Let's examine this pair further.

```{r}

stan_input_lst <- prepareDataStan(current_cohort)
row_ind <- which(stan_input_lst$I_fit == i & stan_input_lst$J_fit == j)

pair_events <- stan_input_lst$event_matrix[row_ind, ,]
window_lengths <- stan_input_lst$finishing_time

# this is the events between this pair

load(paste(dc_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

fit_cohort_mmhp_dc <- fit_cohort_mmhp
sim_cohort_mmhp_dc <- sim_cohort_mmhp

load(paste(save_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

lambda0_pair <- mean(sim_cohort_mmhp_dc$gamma[,i]) +
  mean(sim_cohort_mmhp_dc$zeta[,j])

lambda0_pair

mean(sim_cohort_mmhp$lambda0)

mean(sim_cohort_mmhp_dc$f[,i]) - mean(sim_cohort_mmhp$f[,i])

mean(sim_cohort_mmhp_dc$f[,j]) - mean(sim_cohort_mmhp$f[,j])

# essentially no difference in the latent ranks, some differences in other parameters

mean(sim_cohort_mmhp_dc$beta) - mean(sim_cohort_mmhp$beta)
mean(sim_cohort_mmhp_dc$eta_1) - mean(sim_cohort_mmhp$eta_1)
mean(sim_cohort_mmhp_dc$eta_2) - mean(sim_cohort_mmhp$eta_2)
mean(sim_cohort_mmhp_dc$eta_3) - mean(sim_cohort_mmhp$eta_3)

```

```{r}
pair_events <- as.data.frame(t(pair_events))
colnames(pair_events) <- paste("wind_",c(1:ncol(pair_events)),sep = "")

tidy_events <- tibble(pair_events) %>% 
  pivot_longer(starts_with("wind"), names_to = "wind") %>%
  filter(value > 0) %>%
  group_by(wind) %>%
  summarise(times = list(value)) %>%
  mutate(wind_num = stringr::str_extract(wind, pattern = "[:digit:]+"),
         wind_num = as.numeric(wind_num)) %>%
  arrange(wind_num)

tidy_events # this is only windows with events now

windows_pair <- tibble(wind_num = 1:length(window_lengths),window_lengths)

events <- windows_pair %>% left_join(tidy_events) %>% select(-wind) %>%
  rowwise() %>%
  mutate(num_events = length(times))

```

Then we need to try get the latent states for each of these windows also.

```{r}
load(paste(dc_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[70]]
rm(interpolation_array_list) # too big to keep around

```

However, the only windows with any uncertainty in their latent state, for this pair, are those where

```{r}

mean_state <- matrix(nrow =  length(window_lengths), ncol = no_segments)

for(i in 1:length(window_lengths)) {
  mean_state[i, ] <- apply(latent_pair[[i]], 1,var)
}

as.data.frame(mean_state)

```

Did this happen when not using the degree correction? It seems likely.

```{r}
load(paste(save_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[row_ind]]
rm(interpolation_array_list) # too big to keep around

mean_state <- matrix(nrow =  length(window_lengths), ncol = no_segments)

for(i in 1:length(window_lengths)) {
  mean_state[i, ] <- apply(latent_pair[[i]], 1,mean)
}

as.data.frame(mean_state)
```

```{r}
tibble(send = stan_input_lst$I_fit, rec = stan_input_lst$J_fit,
       num_events = apply(stan_input_lst$Nm, 1, sum), id = row_number(stan_input_lst$I_fit)) %>%
  arrange(desc(num_events))

```

Let's check from before I updated to account for empty windows.

This is still to big to load locally.

```{r, eval=FALSE}
old_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/extend_stan/"

current_cohort <- 1

inter_array_list <- readRDS(paste(old_path,cohort_names[current_cohort],
                                  "/cohort_mmhp_state_array_list_",
                                  cohort_names[current_cohort],".RDS",sep = ""))

pair_array <- inter_array_list[[row_ind]]
rm(inter_array_list)

dim(pair_array[[27]])

```
Let's look at a single pair from Jing's original fit.

```{r}
single_pair <- readRDS("C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/single_inter_array.RDS")

dim(single_pair[[2]])
apply(single_pair[[25]],1,var)

for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(single_pair[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(single_pair[[i]],1,var)),"\n",sep = ""))
}

```
We can repeat this for the previous fit also and compare.

```{r}
for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(pair_array[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(pair_array[[i]],1,var)),"\n",sep = ""))
}

```
These numbers make no sense for sure.

Let's compare to the current model.

```{r}
load(paste(dc_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[69]]

for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(latent_pair[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(latent_pair[[i]],1,var)),"\n",sep = ""))
}

```
So a handful of windows give different results but overall there is little 
difference, and it seems likely all of these differences are due to different
parameters and also running less segments for each window.