### 2020-07-07 output_scripts/Diagnostics.Rmd
update diagnostics and paper plots

---
title: "Diagnostics Week of June 28th"
author: "Owen Ward"
date: "6/30/2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

<!-- The aim of this script is to further investigate model fit  -->
<!-- by better diagnostic -->
<!-- plots for mice pairs in a specific window. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_prior/"

data_path <- "../output/"

func_dir <- "../lib/"

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

source("C:/Users/owenw/Dropbox/with Owen/code/part3/lib/intensityPlot.R")
source("C:/Users/owenw/Dropbox/with Owen/code/part3/lib/uniHawkes.R")

theme_set(theme_minimal())

model1_fn <- list(alpha.fun = function(x,y,eta1,eta2,eta3){return(eta1*x*y*exp(-eta2*abs(x-y))/(1+exp(-eta3*(x-y))))})

# c_mmhp_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_folders/output_june30/"
# 
# other_path  <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"
# old_data_path  <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"

```

```{r mice_setup,include=FALSE,echo=FALSE, cache=TRUE}
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <- expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
}

```

# Correction to Pair level rates

A previous bug with C-MMHP was not correctly accounting for the degree 
corrected `lambda0` and corresponding
`lambda1` between each pair when interpolating the latent states and then 
identifying the residual matrices. This gives more reasonable results for
overall model fit in terms of the total Pearson residuals summed across all 
observation windows.

```{r}
cohorts <- 1:10

for(current_cohort in cohorts) {
  m1_pr <- readRDS(paste(data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  m2_pr <- readRDS(paste(data_path,cohort_names[current_cohort],
                       "/dchp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  m3_pr <- readRDS(paste(data_path,cohort_names[current_cohort],
                       "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  indep_pr <- readRDS(paste(data_path,cohort_names[current_cohort],
                       "/immhp_pr_matrix_",cohort_names[current_cohort],
                       ".RDS",sep=''))

  matrix_lst_plot <- list(
                          indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                     expert_rank_10[[current_cohort]]],
                          m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                          m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                          m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]]) 
  cat("Cohort ",current_cohort,"\n",sep="")
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-60,max=60,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("I-MMHP","C-HP","C-DCHP","DC C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-60,-30,0,30,60), fake_matrix=FALSE,
                    matrix.mar=c(2.5,2.5,2.5,1)
                    )
}

```

# Examine Intensity and QQ Plot

```{r test pairs, include=FALSE,eval=FALSE}
current_cohort <- 1

clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

```

```{r select cohort pair and window,eval=FALSE,include=FALSE}
no_segments <- 500

current_cohort <- 4#5
current_i <- 2 #8
current_j <- 8 #2
current_win <- 42
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

```

```{r plot_functions, include=FALSE, cache=TRUE}
drawRealIntensityPaper <- function(t,
                                   object=NULL,simulation=NULL,
                                   yupper=10,add=FALSE,color=1,line.width=1,
                                 title_name="",title.cex=9, y.ratio=0,
                                 min.y=0, min.x=0, max.x=NULL, 
                                 annotate.time=FALSE,
                                 event.cex=1, box.type="l",
                                 box.col="gray40", title.line=1){
  # input object: the parameter list used for generating mmhp
  #       simulation: simulation result from simulate.mmhp 
  if(t[1]>0){
    t <- c(0,t)
  }
  if(is.null(max.x)){
    max.x <- tail(t,1)
  }
  if(is.null(object)){
    termination <- tail(t,1)
    n <- length(t)
    plot(0,0,xlim=c(min.x,max.x),ylim=c(min.y,yupper*(1+y.ratio)),
         type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
    box(lty ='solid', col=box.col, lwd=0.7, bty=box.type)
    title(title_name,line=title.line,cex.main=title.cex)
    #left axes
    #axis(2,at=seq(0,yupper,5),labels=FALSE,cex.lab=2,lwd=0,lwd.ticks=1) 
    #text(-termination/50, y=seq(0,yupper,5), labels=seq(0,yupper,5), cex=1.6, srt=0, xpd=TRUE)
    #bottom axes
    # events
    points(t[-1],rep(min.y,n-1),cex=event.cex,pch=16,col="black")
  }else{
    state <- simulation$z
    state_time <- simulation$x
    if(tail(state_time,1)<tail(t,1)){
      state_time <- c(state_time,tail(t,1))
      state <- c(state,3-tail(state,1))
    }
    lambda0 <- object$lambda0
    lambda1 <- object$lambda1
    alpha <- object$alpha
    beta <- object$beta
    termination <- tail(t,1)
    n <- length(t)
    m <- length(state)
    if(is.null(min.y)){
      min.y <- -lambda0
    }
    if(is.null(max.x)){
      max.x <- termination
    }
    #plot(1,4,xlim=c(0.27,23),ylim=c(-0.5,9), type="n",xlab="Time",ylab="",xaxt="n",yaxt="n",cex.lab=1.8,bty="n")

    if(add==FALSE){
      plot(0,0,xlim=c(min.x,max.x),ylim=c(min.y,yupper*(1+y.ratio)),
           type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
      box(lty ='solid', col=box.col, lwd=0.7, bty=box.type)
      title(title_name,line=title.line,cex.main=title.cex)
      #left axes
      #axis(2,at=seq(0,yupper,5),labels=FALSE,cex.lab=2,lwd=0,lwd.ticks=1) 
      #text(-termination/50, y=seq(0,yupper,5), labels=seq(0,yupper,5), cex=1.6, srt=0, xpd=TRUE)
      #bottom axes
      if(annotate.time){
        axis(1,at=seq(0,ceiling(termination),length.out=5),labels=FALSE,cex.lab=2,lwd=0,lwd.ticks=1) #bottom axes
        text(seq(0,ceiling(termination),length.out=5), y=-y.ratio*6, cex=1.7, srt=0, xpd=TRUE,
             labels=round(seq(0,ceiling(termination),length.out=5),0))
      }
      # events
      points(t[-1],rep(min.y,n-1),cex=event.cex,pch=16,col="black")
    }

    if(m==2&state[1]==1&state_time[2]==tail(t,1)){
      drawHawkesIntensityPaper(lambda0,alpha,beta,t,color=color,
                               line.width=line.width)
    }else{
      for(i in 1:(m-1)){
        if(state[i]==1){
          hawkes_time <- t[t>=state_time[i]&t<=state_time[i+1]]
          if(i==1) hawkes_time <- hawkes_time[-1]
          history <- t[t<state_time[i]]
          # bug here
          drawHawkesIntensityPaper(lambda0 =lambda1,
                                   #i,
                                   alpha = alpha,
                                   beta = beta,
                                   #state_time[i],
                                   #state_time[i+1],
                                   #history[-1],
                                   events = hawkes_time,
                                   color=color,line.width=line.width)
        }else{
          segments(x0=state_time[i],x1=state_time[i+1], y0=lambda0,
                   lty=1,col=color,lwd=line.width)
        }
        if(i>1){
          segments(x0=state_time[i],y0=lambda0, y1=lambda1, lty=2, col=color,lwd=line.width)
        }
      }
    }
  }
}

drawHawkesIntensityPaper <- function(lambda0,alpha,beta,
                                     events,color=1,line.width=2, termination=NULL){
  # input object: parameters for Hawkes process, include lambda0, alpha, beta 
  #       events: vector of event happening time
  horizon <- tail(events,1)
  events <- events[events>0]
  N <- length(events)

  if(N==1){ # only one event condition
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)
    if(horizon > tail(events,1)){
      lambda.n <- function(s) lambda0+alpha*sum(exp(-beta*(s-events[1])))
      new.lambda.n<-Vectorize(lambda.n)
      curve(new.lambda.n, from=tail(events,1), to=horizon, add=TRUE, col=color,lwd=line.width)
    }
    #if( !is.null(title) ) title(main=title)
  }else{
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)

    for(i in 1:(N-1)){
      lambda.n <- function(s) lambda0+alpha*sum(exp(-beta*(rep(s,i)-events[1:i])))
      new.lambda.n<-Vectorize(lambda.n)
      curve(new.lambda.n, from=events[i], to=events[i+1], add=TRUE, col=color,lwd=line.width)
      segments(x0=events[i+1],y0=lambda.n(events[i+1]),y1=lambda.n(events[i+1])+alpha, col=color,lwd=line.width)
    }
  }
  if(!is.null(termination)){
    lambda.n <- function(s) lambda0+alpha*sum(exp(-beta*(rep(s,N)-events[1:N])))
    new.lambda.n<-Vectorize(lambda.n)
    curve(new.lambda.n, from=events[N], to=termination, add=TRUE, col=color,lwd=line.width)
  }
}

```

```{r extract hawkes fits, results='hide',eval=FALSE,include=FALSE}
print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

```

```{r extract_cmmhp_fit, results='hide',eval=FALSE,include=FALSE}
print(current_cohort)
## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

```

```{r extract indep mmhp fit, results='hide',eval=FALSE,include=FALSE}

print(current_cohort)
#### I-MMHP  ####
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

```

For each cohort, we extract the mouse pair with the most interactions, plotting
thes estimated intensity in the window with the most interactions along with
the QQ plot of all events (in all windows) between that pair.

```{r pair diagnostic plot, include=FALSE,eval=FALSE}
row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

We want to repeat this for the pair with the most interactions in a single
cohort, displaying the fitted intensity in the window where they have
the most interactions.

### Cohort 1

```{r cohort-1,cache=TRUE}
current_cohort <- 1
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]

unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

# row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
# row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
# current_event_time <- return_df[row_indicator,"event.times"][[1]]
# observe_period <- return_df[row_indicator,"observe.time"]
# no_segments <- 500
# time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 2

```{r cohort-2,cache=TRUE}
current_cohort <- 2
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]

unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 3

```{r cohort-3,cache=TRUE}
current_cohort <- 3
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]

unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 4

```{r cohort-4,cache=TRUE}
current_cohort <- 4
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 5

```{r cohort-5,cache=TRUE}
current_cohort <- 5
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 6

```{r cohort-6,cache=TRUE}
current_cohort <- 6
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 7

```{r cohort-7,cache=TRUE}
y.upper <- 80
current_cohort <- 7
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 8

```{r cohort-8,cache=TRUE}
current_cohort <- 8
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 9

```{r cohort-9,cache=TRUE}
y.upper <- 60
current_cohort <- 9
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win[1]#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
#y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```

### Cohort 10

```{r cohort-10,cache=TRUE}
current_cohort <- 10
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[1]
best_j <- top_pairs$recipient[1]
wind <- unlist(lapply(top_pairs$no.events,which.max))[1]

win_length <- top_pairs$observe.length[[1]][wind]
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])
actual_win <- which(unique_observe_win$observe.time == win_length)

no_segments <- 500

current_i <- best_i #8
current_j <- best_j #2
current_win <- actual_win#wind 
# this is the 18th window, not the 18th window with observations
### that this pair have events in, which it should be

# clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
# return_df <- cleanObservationPeriod(current_cohort, clean_data)
# unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
#   dplyr::summarize(count=n(),
#             observe=list(observe.id),
#             observe.length=list(observe.time),
#             no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
time_segment <- seq(0,observe_period,length.out=no_segments)

### load the corresponding stan fits ####
#print(current_cohort)

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model1_par_est <- list(lambda0=mean(sim_cohort_hp$lambda0),
                       eta_1=mean(sim_cohort_hp$eta_1),
                       eta_2=mean(sim_cohort_hp$eta_2),
                       eta_3=mean(sim_cohort_hp$eta_3),
                       beta=mean(sim_cohort_hp$beta),
                       f=apply(sim_cohort_hp$f,2,mean))
model1_par_matrix <- list(lambda0_matrix=matrix(model1_par_est$lambda0,
                                                nrow=mice_number,ncol=mice_number),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model1_par_est$eta_1,
                                                                                    model1_par_est$eta_2,
                                                                                    model1_par_est$eta_3),
                                                  model1_par_est$f),
                          beta_matrix=matrix(model1_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

#### Degree Corrected Hawkes ####
load(paste(data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

# create a list to store the hawkes parameters for this
model2_par_est <- list(gamma = apply(sim_cohort_dchp$gamma,2,mean),
                       zeta = apply(sim_cohort_dchp$zeta,2,mean),
                       #lambda0=mean(sim_cohort_dchp$lambda0),
                       eta_1=mean(sim_cohort_dchp$eta_1),
                       eta_2=mean(sim_cohort_dchp$eta_2),
                       eta_3=mean(sim_cohort_dchp$eta_3),
                       beta=mean(sim_cohort_dchp$beta),
                       f=apply(sim_cohort_dchp$f,2,mean))
model2_par_matrix <- list(lambda0_matrix= outer(model2_par_est$gamma,
                                                model2_par_est$zeta,FUN = "+"),
                          alpha_matrix=formMatrix(function(x,y) model1_fn$alpha.fun(x,y,
                                                                                    model2_par_est$eta_1,
                                                                                  model2_par_est$eta_2,
                                                                                    model2_par_est$eta_3),
                                                  model2_par_est$f),
                          beta_matrix=matrix(model2_par_est$beta,
                                             nrow=mice_number,ncol=mice_number))

hawkes.par_dc <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/mmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

c_mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

c_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
c_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=c_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

## Cohort MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/sep_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

sep_mmhp_par <- list(lambda1=mean(sim_mmhp_sep$lambda1[,pair]),
                 lambda0=mean(sim_mmhp_sep$lambda0[,pair]),
                 alpha=mean(sim_mmhp_sep$alpha[,pair]),
                 beta=mean(sim_mmhp_sep$beta),
                 q1=mean(sim_mmhp_sep$q1[,pair]),
                 q2=mean(sim_mmhp_sep$q2[,pair]))

sep_mmhp.est.latent <- ifelse(apply(state_array_list[[pair]][[current_win]],
                                1,
                                function(x) sum(x==1)) >500,rep(1,1000),rep(2,1000))
sep_mmhp.est.latent <- interpolateLatentTrajectory(params=c_mmhp_par,
                                               events=current_event_time,
                                                       zt=sep_mmhp.est.latent,                                         initial.state=mean(initial_state_list[[pair]][[current_win]]),
                     termination.time=observe_period,
                     termination.state = 2)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

row_indicator <- return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win
row_idx <- which(return_df$initiator==current_i&return_df$recipient==current_j&return_df$observe.id==current_win)
current_event_time <- return_df[row_indicator,"event.times"][[1]]
observe_period <- return_df[row_indicator,"observe.time"]
no_segments <- 500
time_segment <- seq(0,observe_period,length.out=no_segments)

par(mfrow=c(2,4), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20

#### Cohort Hawkes  ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="C-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)

### DC Cohort Hawkes ####

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="DC-HP",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par_dc[1],
                         alpha=hawkes.par_dc[2],
                         beta=hawkes.par_dc[3],
                         events=current_event_time, termination=observe_period)

#### DC C-MMHP ####
drawRealIntensityPaper(current_event_time, object=c_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,c_mmhp.est.latent$x.hat),
                                       z=c_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="DC C-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### Sep MMHP ####

drawRealIntensityPaper(current_event_time, object=sep_mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,sep_mmhp.est.latent$x.hat),
                                       z=sep_mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="I-MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")

#### then the corresponding QQ plots ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

#### Cohort Hawkes ####

Lambda.c_hawkes.test <- Lambda_c_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.c_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.c_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### Cohort DC Hawkes ####

Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### DC C-MMHP ####
Lambda.cmmhp.test <- Lambda_cmmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.cmmhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.cmmhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

#### I-MMHP ####
Lambda.immhp.test <- Lambda_mmhp_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.immhp.test))    # 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.immhp.test,p=p,na.rm=TRUE) # percentiles of the sample distribution
plot(qexp(p),q, 
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1,cex=1,col="Black",ylim=c(0,8),xlim=c(0,5))
mtext(side=1, text="TQ", line=1.6, cex=anno.cex)
mtext(side=2, text="EQ", line=1.2, cex=anno.cex)
Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:8),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

```