### 2020-08-26 output_scripts/aug_updates.Rmd
start adding updates for James

---
title: "Comments on Latent Ranking"
author: "Owen Ward"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

data_path <- "../output/"

func_dir <- "../lib/"

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
library(ggplot2)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

#source("C:/Users/owenw/Dropbox/with Owen/code/part3/lib/intensityPlot.R")
#source("C:/Users/owenw/Dropbox/with Owen/code/part3/lib/uniHawkes.R")

theme_set(theme_minimal())

model1_fn <- list(alpha.fun = function(x,y,eta1,eta2,eta3){
  return(eta1*x*y*exp(-eta2*abs(x-y))/(1+exp(-eta3*(x-y))))}
  )
```

```{r mice_setup,include=FALSE,echo=FALSE, cache=TRUE}
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
for(current_cohort in fit_cohorts){
  naive_rank_10[[current_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[current_cohort]]])
  expert_rank_10[[current_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[current_cohort]]])
}
```

# Exploring eta parameters

```{r}
for(current_cohort in 1:10) {
  #cat("Cohort, ", current_cohort,"\n",sep="")
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  eta <- tibble(eta_1 = sim_cohort_mmhp$eta_1,
                eta_2 = sim_cohort_mmhp$eta_2,
                eta_3 = sim_cohort_mmhp$eta_3)
  p <- GGally::ggpairs(eta) + 
    ggtitle(paste("Cohort ", current_cohort,sep=""))

  print(p)
}
```

To explore what these parameters correspond to, the model for
the conditional intensity is given by

$$
\lambda^{i,j}(t) = \gamma_i + \zeta_j +(\lambda_1-\gamma_i-\zeta_j)Z^{i,j}(t) +
\eta_1 f_i f_j 
\exp(-\eta_2|f_i-f_j|)Z^{i,j}(t)\sum_{k}\exp(-\beta(t-t_{k}^{i,j})),
$$
where 
$$
q_1^{i,j} = \exp(-\eta_3 f_i),\ q_0^{i,j} = \exp(-\eta_3 f_j).
$$
So here, these parameters correspond to:
- $\eta_1$

# Are gamma/zeta associated with rankings?

# Do closer ranked mice fight more?