### 2020-11-17 output_scripts/rank_recovery.Rmd
add frob norm and structure score plots

---
title: "Rank Recovery"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: TRUE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
library(here)
library(tidyverse)
library(rstan)
library(cmdstanr)
library(compete)
library(posterior)

theme_set(theme_minimal())
```

# Introduction

We want to consider two further concepts related to the models we have built
and used as comparisons in this paper. In particular:

1. Examine the performance of the comparison models in recovering data simulated under
the (degree corrected) C-MMHP model.
1. Based on the above, and using the I&SI objective, evaluate the 
performance of models when we estimate the latent ranking on a subset of the
data, such as the first half of the events, and then infer the I and SI scores
based on those rankings for the remaining events.

# Comparison models on simulated data

We first simulate a single dense network from the C-MMHP model,
with a latent rank vector given by

$$
f = (0.1, 0.2, 0.4, 0.7, 0.9).
$$

```{r load functions}
source(here("lib",'uniHawkes.R'))
source(here("lib","mmhp.R"))
source(here("lib",'simulatePrediction.R'))
source(here("lib",'plotUtil.R'))
source(here("lib",'inferLatentMmhp.R'))
source(here("lib",'drawIntensity.R'))
source(here("lib",'myGlicko.R'))
source(here("lib",'naiveRankHierarchy.R'))
source(here("lib",'expertRankHierarchy.R'))
cut_off <- 3

model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3){
  return(eta1 * x * y *exp(-eta2 * abs(x - y))/(1 + exp(-eta3 * (x - y) ) ) ) })

model3_fn <- list(alpha.fun = function(x, y, eta1, eta2){
  return(eta1 * x * y * exp(-eta2 * abs(x - y) ) )},
                  q1.fun = function(x, y, eta3){
                    return(exp(-eta3 * x) )},
                  q0.fun = function(x, y, eta3){
                    return(exp(-eta3* y) ) })

#### Save the simulation parameters ####

object_fn <- list(alpha.fun = function(x, y, eta1, eta2){
  return(eta1* x *y *exp(-eta2 * abs(x - y) ) )},
                  q1.fun = function(x, y, eta3){
                    return(exp(- eta3 * x) )},
                  q0.fun = function(x, y, eta3){
                    return(exp(- eta3 * y) )})

object_par <- list(sim_lambda_0 = 0.08,
                   sim_lambda_1 = 0.15,
                   sim_eta_1 = 2.5,
                   sim_eta_2 = 0.6,
                   sim_eta_3 = 5,
                   sim_beta = 1.5,
                   f_vec_1=c(0.1, 0.2, 0.4, 0.7, 0.9))

object_matrix <- list(lambda0_matrix = matrix(object_par$sim_lambda_0,
                                            nrow = length(object_par$f_vec_1),
                                            ncol = length(object_par$f_vec_1)),
                      lambda1_matrix = matrix(object_par$sim_lambda_1,
                                            nrow = length(object_par$f_vec_1),
                                            ncol = length(object_par$f_vec_1)),
                      alpha_matrix = formMatrix(function(x,y)
                        object_fn$alpha.fun(x, y,
                                            object_par$sim_eta_1, 
                                            object_par$sim_eta_2),
                        object_par$f_vec_1),
                      beta_matrix = matrix(object_par$sim_beta,
                                         nrow = length(object_par$f_vec_1),
                                         ncol = length(object_par$f_vec_1)),
                      q1_matrix = formMatrix(function(x, y) 
                        object_fn$q1.fun(x, y, object_par$sim_eta_3),
                                           object_par$f_vec_1),
                      q2_matrix = formMatrix(function(x, y) 
                        object_fn$q0.fun(x, y, object_par$sim_eta_3),
                                           object_par$f_vec_1))

```

## Simulated data from original model

```{r sim data from c-mmhp, eval=FALSE}
sim_model3_data <- simulateLatentMMHP(lambda0_matrix = object_matrix$lambda0_matrix,
                                      lambda1_matrix = object_matrix$lambda1_matrix,
                                      alpha_matrix = object_matrix$alpha_matrix,
                                      beta_matrix = object_matrix$beta_matrix,
                                      q1_matrix = object_matrix$q1_matrix,
                                      q2_matrix = object_matrix$q2_matrix,
                                      horizon = 200)
clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data, 
                                        cut_off = cut_off, 
                                      N = length(object_par$f_vec_1))

saveRDS(clean_sim_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS"))

```

```{r recover cmmhp model, eval=FALSE}
# model3 <- stan_model(here("lib","sim_model3.stan"))
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS"))
model3 <- cmdstan_model(here("lib","sim_model3.stan"))
data_list <- list(max_Nm=max(clean_sim_data$N_count),
                    N_til = length(clean_sim_data$I_fit),
                    M=sum(clean_sim_data$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data$I_fit,
                    J_fit = clean_sim_data$J_fit,
                    T = tail(clean_sim_data$day_hour,1),
                    Nm = as.vector(clean_sim_data$N_count[clean_sim_data$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data$event_matrix,
                    interevent_time_matrix = clean_sim_data$time_matrix,
                    max_interevent = clean_sim_data$max_interevent)

model3_stan_fit <- model3$sample(data = data_list,
                                seed = 123,
                                iter_warmup = 500,
                                iter_sampling = 500,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)
model3_stan_fit$save_object(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))

```

We can then extract the posterior mean rank from our fitted model.

```{r extract stan fit}
model3_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))

m3_draws <- as_draws_df(model3_stan_fit$draws())
m3_post_rank <- m3_draws %>% select(starts_with("f")) %>% apply(2,mean)
m3_post_rank
## truth is .1, .2, .4, .7, .9

# posterior mean
# posterior sd
# apply(stansims$f, 2, sd)

# most important is the order
m3_rank <- order(m3_post_rank)
#order(apply(stansims$f, 2, sd))
```

We can compare this to the ranking from I&SI on the count data alone, which
is given below. This incorrectly orders the two lowest ranked nodes.

```{r fit I&SI, results='hide'}
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS") )
count_data <- get_wl_matrix(df = cbind(clean_sim_data$start,clean_sim_data$end))
isi.out <- compete::isi98(m = count_data, random = TRUE)
isi_rank <- as.numeric(rev(isi.out$best_order)) # ranked from lowest to highest
```
```{r print isi result}
isi_rank
```

Similarly we can compare these and the true ranking pairwise.

```{r overall sim ranks}
true_rank <- c(1:5)

table(true_rank, m3_rank)
table(true_rank, isi_rank)
table(isi_rank, m3_rank)
```

## Simulated from degree corrected model

How does this ranking inference compare when we instead simulate
data from the proposed degree corrected model.

```{r sim dc mmhp, eval=FALSE}
gamma_var <- c(0.01, 0.02, 0.03, 0.06, 0.07)
zeta_var <- c(0.075, 0.02, 0.03, 0.05, 0.08 )
lambda_dc_matrix <- outer(gamma_var, zeta_var, "+")
# lambda_dc_matrix

object_dc_matrix <- object_matrix
object_dc_matrix$lambda0_matrix <- lambda_dc_matrix

sim_model3_data_dc <- 
  simulateLatentMMHP(lambda0_matrix = object_dc_matrix$lambda0_matrix,
                      lambda1_matrix = object_dc_matrix$lambda1_matrix,
                      alpha_matrix = object_dc_matrix$alpha_matrix,
                      beta_matrix = object_dc_matrix$beta_matrix,
                      q1_matrix = object_dc_matrix$q1_matrix,
                      q2_matrix = object_dc_matrix$q2_matrix,
                      horizon = 200)
clean_sim_data_dc <- cleanSimulationData(raw_data = sim_model3_data_dc, 
                                        cut_off = cut_off, 
                                      N = length(object_par$f_vec_1))

saveRDS(clean_sim_data_dc,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS"))

```

```{r fit dc mmhp, eval=FALSE}
clean_sim_data_dc <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS"))
model3_dc <- cmdstan_model(here("lib","sim_model3_dc.stan"))
data_list_dc <- list(max_Nm = max(clean_sim_data_dc$N_count),
                    N_til = length(clean_sim_data_dc$I_fit),
                    M=sum(clean_sim_data_dc$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data_dc$I_fit,
                    J_fit = clean_sim_data_dc$J_fit,
                    T = tail(clean_sim_data_dc$day_hour,1),
                    Nm = as.vector(clean_sim_data_dc$N_count[
                      clean_sim_data_dc$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data_dc$event_matrix,
                    interevent_time_matrix = clean_sim_data_dc$time_matrix,
                    max_interevent = clean_sim_data_dc$max_interevent)

model3_dc_stan_fit <- model3_dc$sample(data = data_list_dc,
                                seed = 123,
                                iter_warmup = 500,
                                iter_sampling = 500,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)

model3_dc_stan_fit$save_object(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_dc_stan_fit.RDS"))
```

```{r load dc stan fit}
model3_dc_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_dc_stan_fit.RDS"))
m3_dc_draws <- as_draws_df(model3_dc_stan_fit$draws())
m3_dc_rank <- m3_dc_draws %>% select(starts_with("f")) %>% apply(2,mean)
m3_dc_rank
## truth is .1, .2, .4, .7, .9
# apply(stansims_dc$f, 2, sd)

# most important is the order
m3_dc_rank <- order(m3_dc_rank)
```

Then we can compare the I&SI model here also. Here there is exact agreement.

```{r i and si dc, results='hide'}
clean_sim_data_dc <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS") ) 
count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data_dc$start, 
                                       clean_sim_data_dc$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
isi_rank_dc <- as.numeric(rev(isi_dc.out$best_order)) 
# ranked from lowest to highest
```

```{r tables of agreement}
isi_rank_dc
true_rank <- c(1:5)

table(true_rank, m3_rank)
table(true_rank, isi_rank_dc)
table(isi_rank_dc, m3_rank)
```

So when we consider this degree corrected model we
correctly recover the latent ranking order in this simulation from 
both the stan fit and the I&SI model.

## Multiple Simulations

To confirm this result we simulate and fit these models to data from 
the degree corrected model 20 times. We wish to see how many times 
they agree.

We see that the C-MMHP model somewhat better recovers the true
ranks in this simulation, with both performing well. We also
see that any discrepancies from the true ranking are small, with the
order of a pair of nodes being swapped.

```{r load in sims, options}
sim_files <- list.files(here("output", "rank_sims",
                             "sims_m3_dc_isi"))

all_sims <- tibble()
for(i in seq_along(sim_files)) {
 single_fit <- readRDS(here("output", "rank_sims", 
                            "sims_m3_dc_isi",
                            sim_files[i]))
 all_sims <- all_sims %>% bind_rows(single_fit)
}

```

We can look at the overall summary of the ranking of the two models and
the truth.

```{r summary ranks}
all_sims %>% 
  group_by(truth,m3_dc,isi) %>%
  count() %>% arrange(-n) 
```

We would like to identify how many times each model correctly recovers the
full ranking.
We see that the C-MMHP is fully correct 80% of the time while the 
I&SI ranking is fully correct in 60% of these simulations.

```{r prop full recover}
all_sims$sim <- rep(1:50, each = 5)

all_sims %>% select(sim, truth, m3_dc) %>%
  mutate(diff = truth - m3_dc) %>%
  group_by(sim) %>%
  summarise(prop_diff = mean(diff!=0)) %>%
  summarise(full_recovery = mean(prop_diff == 0))

all_sims %>% select(sim, truth, isi) %>%
  mutate(diff = truth - isi) %>%
  group_by(sim) %>%
  summarise(prop_diff = mean(diff!=0)) %>%
  summarise(full_recovery = mean(prop_diff == 0))
```

We see strong agreement between our model and the true ranking of
the simulated data for most nodes.

```{r truth_m3}
all_sims %>%
  group_by(truth, m3_dc) %>%
  count() %>%
  mutate(per = n/50) %>%
  arrange(-per)
```

Finally, we look at the agreement between the ISI ranking and the truth.

```{r truth isi}
all_sims %>%
  group_by(truth, isi) %>%
  count() %>%
  mutate(per = n/50) %>%
  arrange(-per)

```

Similarly, we can summarise this by
computing the Spearman rank correlation between each of these 
models and the truth. The C-MMHP model unsurprisingly best recovers the
true rankings, with the I&SI model performing similarly.

```{r analyse sims}
sim_files <- list.files(here("output", "rank_sims",
                             "sims_m3_dc_isi_sparse"))

all_sims <- list()
for(i in seq_along(sim_files)) {
 single_fit <- readRDS(here("output", "rank_sims", 
                            "sims_m3_dc_isi",
                            sim_files[i]))
 all_sims[[i]] <- single_fit
}

spear <- function(sim_data) {
  truth <- sim_data$truth
  sim_data %>%
    select(m1:isi) %>%
    map( ~cor.test(truth, .x,
                   method = "spearman")$estimate  ) %>%
    enframe() %>%
    unnest(cols = c(value))
}

sim_data <- map_dfr(all_sims, spear)

name <- sim_data$name

sim_data$name <- case_when(
          name == "isi" ~ "I&SI",
          name == "m1" ~ "C-HP",
          name == "m2" ~ "C-DCHP",
          name == "m3_dc" ~ "C-MMHP")

sim_data$name <- factor(sim_data$name, levels = c("I&SI","C-HP",
                                                  "C-DCHP", "C-MMHP"))

sim_data %>% 
  rename(Method = name, corr = value) %>%
  ggplot(aes(Method,corr)) + geom_boxplot() +
  labs(y = "Rank Correlation")
  # in a nice long format

```

# Extrapolation from inferred rank on subset

We also wish to consider how well we can infer the latent rank 
over time. To investigate this, we wish to split the data in a Mice cohort
(or simulation data) temporally, and examine the inferred model on the 
first half of the interactions compared to on the remaining data,
by looking at the inferred I&SI scores.

```{r compute isi}
# taking code from compete package
compute_isi <- function(win_loss_matrix, curr_ranking) {
  # this takes ranking from highest to lowest
  wl <- win_loss_matrix[curr_ranking,curr_ranking]
  n <- as.numeric(dim(wl)[1])
  m <- matrix(0, n, n)
  siind <- matrix(0, n, n)
  I = 0
  SI = 0
  for (j in 2:n) {
      for (i in 1:(j - 1)) {
          siind[j, i] = j - i
          siind[i, j] = j - i
          if (wl[j, i] == 0 & wl[i, j] == 0) {
              m[j, i] = 0
              m[i, j] = 0
          }
          else {
              if (wl[j, i] > wl[i, j]) {
                m[j, i] = 1
                m[i, j] = -1
              }
              if (wl[j, i] < wl[i, j]) {
                m[j, i] = -1
                m[i, j] = 1
              }
              if (wl[j, i] == wl[i, j]) {
                m[j, i] = 0.5
                m[i, j] = 0.5
              }
          }
      }
  }
  I = sum(lower.tri(m) * m == 1)
  SI = sum((lower.tri(m) * m == 1) * siind)

  list(I_score = I, SI_score = SI, wl_mat = wl)
}

```

```{r check isi function, results='hide'}
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS"))
clean_sim_data_dc$N_count

count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data_dc$start, 
                                       clean_sim_data_dc$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)

# these agree

compute_isi(count_data_dc, curr_ranking = c(2,5,4,1,3))

```

We want to fit our models on the first half of the data
and then compare the ranking estimated from it on 
the remaining data, in terms of I&SI score.

```{r sim data and split, eval=FALSE}
sim_model3_data_dc <- 
  simulateLatentMMHP(lambda0_matrix = object_dc_matrix$lambda0_matrix,
                      lambda1_matrix = object_dc_matrix$lambda1_matrix,
                      alpha_matrix = object_dc_matrix$alpha_matrix,
                      beta_matrix = object_dc_matrix$beta_matrix,
                      q1_matrix = object_dc_matrix$q1_matrix,
                      q2_matrix = object_dc_matrix$q2_matrix,
                      horizon = 200)

sim_data <- tibble(start = sim_model3_data_dc$start,
                   end = sim_model3_data_dc$end,
                   day_hour = sim_model3_data_dc$day_hour)

training_data <- sim_data %>% filter(day_hour <= 100)
test_data <- sim_data %>% filter(day_hour > 100 )

saveRDS(training_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))
saveRDS(test_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))

```

```{r fit m3 to first half, eval=FALSE}
training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

clean_training <- cleanSimulationData(training_data, cut_off = 3, N = 5)

data_list_training <- list(max_Nm = max(clean_training$N_count),
                    N_til = length(clean_training$I_fit),
                    M=sum(clean_training$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_training$I_fit,
                    J_fit = clean_training$J_fit,
                    T = tail(clean_training$day_hour,1),
                    Nm = as.vector(clean_training$N_count[
                      clean_training$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_training$event_matrix,
                    interevent_time_matrix = clean_training$time_matrix,
                    max_interevent = clean_training$max_interevent)

model3_dc <- cmdstan_model(here("lib","sim_model3_dc.stan"))

model3_dc_train <- model3_dc$sample(data = data_list_training,
                                seed = 123,
                                iter_warmup = 500,
                                iter_sampling = 500,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)

model3_dc_train$save_object(file = here("output", "rank_sims",
                                        "rmd_stan",
                                        "model_3_dc_train_fit.RDS"))

```

```{r get rank from this,results='hide'}
model3_dc_train <- readRDS(here("output", "rank_sims",
                                        "rmd_stan",
                                        "model_3_dc_train_fit.RDS"))

m3_draws_train <- as_draws_df(model3_dc_train$draws())
m3_train_rank <- m3_draws_train %>% select(starts_with("f")) %>% apply(2,mean)

m3_train_rank

m3_train_rank <- rev(order(m3_train_rank))
# then feed this ranking into the test data

test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))

```

We can compare the I&SI score from the posterior ranks from the
first half of the data with that obtained from using the I&SI
ranks from the first half of the data, which agree here.

```{r fit isi train, results='hide'}

test_wl <- get_wl_matrix(cbind(test_data$start,test_data$end))
m3_test_score <- compute_isi(test_wl,curr_ranking = m3_train_rank)

# then compare this to isi for test data

training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

train_wl <- get_wl_matrix(cbind(training_data$start,training_data$end))

isi_train <- isi98(train_wl)
isi_train_rank <- isi_train$best_order

isi_test_score <- compute_isi(test_wl,curr_ranking = isi_train_rank)

```

```{r}
m3_test_score
isi_test_score
```

How does this compare to the optimal score using only the test data.
They agree in this instance.
```{r fit isi to test}
isi98(test_wl)
```

## Aggregate Ranking Model

We can similarly do this for some of the other models we use
as comparison, such as the Aggregate ranking model.

```{r fit agg rank training, results='hide'}
training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

clean_training <- cleanSimulationData(training_data, cut_off = 3, N = 5)

agg_rank_training <- clean_training$N_count

agg_rank_model <- cmdstan_model(here("lib","latent_rank_agg_sim.stan"))

agg_rank_fit <- agg_rank_model$sample(data = list(n = 5,
                                                  n_matrix = agg_rank_training),
                                      seed = 123,
                                      iter_warmup = 500,
                                      iter_sampling = 500,
                                      chains = 4, 
                                      parallel_chains = 4,
                                      refresh = 100)

agg_rank_draws_train <- as_draws_df(agg_rank_fit$draws())
agg_train_mean_rank <- agg_rank_draws_train %>%
  select(starts_with("x")) %>% apply(2,mean)

# for highest to lowest
agg_rank_train <- rev(order(agg_train_mean_rank))
```

```{r show agg rank mean}
agg_train_mean_rank
```

The aggregate ranking model struggles to correctly identify the correct
ranking, resulting in a large I&SI score.

```{r compute_isi for agg rank}
test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))
test_wl <- get_wl_matrix(cbind(test_data$start, test_data$end))

agg_test_score <- compute_isi(test_wl,curr_ranking = agg_rank_train)
agg_test_score
```

## Glicko Model

We can also perform this comparison with the Glicko 
model on the training data.

```{r fit glicko}
library(PlayerRatings)
source(here("lib","myGlicko.R"))

training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

# training_data
### need score (1) for all of these, along with the row number

glick_df <- training_data %>% 
  mutate(id = row_number(), score = 1) %>%
  select(id, start, end , score)

gl_train <- my_glicko(glick_df, history=TRUE, cval=2)

gl_train

gl_ranks <- rev(gl_train$ratings$Player)
# highest to lowest

```

This correctly recovers the true ranking and will therefore result in 
the same optimal I&SI score.

```{r glicko test}
test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))
test_wl <- get_wl_matrix(cbind(test_data$start, test_data$end))

glicko_test_score <- compute_isi(test_wl,curr_ranking = gl_ranks)
glicko_test_score

```

## Real Data

Here we estimate the rankings from each model using data from the
first 15 days of events, and then compute the estimate I&SI scores
from using those ranks on the remaining held out data.

```{r plot isi scores}
sim_files <- list.files(here("output","rank_sims","predict_isi"))

all_sims <- list()
for(i in seq_along(sim_files)) {
  single_fit <- readRDS(here("output", "rank_sims", 
                             "predict_isi",
                             sim_files[i]))
  all_sims[[i]] <- single_fit
}

cohort_scores <- all_sims %>%
  map(~as_tibble(.)) %>%
  bind_rows(.id="index")

cohort_scores %>% 
  mutate(ISI = I_score + SI_score) %>%
  ggplot(aes(model,ISI)) + geom_boxplot()
```

We can similarly repeat this for fitting each model to all 
data and look at the estimated I&SI scores based on those
inferred rankings.

```{r compute isi total, results='hide'}
score_fits_all <- list()
source(here("lib","cleanData.R"))

data_path <- "../output/"

full_data <- readRDS(here("data","mice.RData"))
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

mini_wl_matrix <- function(df, num_mice = 12) {
  # to avoid issues with reordering the data
  df <- as.data.frame(df)
  mylevs <- 1:num_mice
  df[, 1] <- factor(df[, 1], levels = mylevs)
  df[, 2] <- factor(df[, 2], levels = mylevs)
  df$result <- 1
  m1 = stats::xtabs(result ~ ., data = df)
  m1
}

compute_isi_score <- function(win_loss_matrix, curr_ranking) {
  # this takes ranking from highest to lowest
  wl <- win_loss_matrix[curr_ranking,curr_ranking]
  n <- as.numeric(dim(wl)[1])
  m <- matrix(0, n, n)
  siind <- matrix(0, n, n)
  I = 0
  SI = 0
  for (j in 2:n) {
    for (i in 1:(j - 1)) {
      siind[j, i] = j - i
      siind[i, j] = j - i
      if (wl[j, i] == 0 & wl[i, j] == 0) {
        m[j, i] = 0
        m[i, j] = 0
      }
      else {
        if (wl[j, i] > wl[i, j]) {
          m[j, i] = 1
          m[i, j] = -1
        }
        if (wl[j, i] < wl[i, j]) {
          m[j, i] = -1
          m[i, j] = 1
        }
        if (wl[j, i] == wl[i, j]) {
          m[j, i] = 0.5
          m[i, j] = 0.5
        }
      }
    }
  }
  I = sum(lower.tri(m) * m == 1)
  SI = sum((lower.tri(m) * m == 1) * siind)

  list(I_score = I, SI_score = SI)
}

for(current_cohort in 1:10) {
  # current_cohort <- cohort_id
  print(paste("Cohort",current_cohort))

  training_day <- 15
  ## if I change this then I need to refit

  print(current_cohort)
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(current_cohort, clean_data)

  #### Fit M1 ####
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_hp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  # highest to lowest

  rank_m1 <- rev(order(apply(sim_cohort_hp$f, 2, mean)))

  #### Predict M2 ####
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_dchp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
  # highest to lowest
  rank_m2 <- rev(order(apply(sim_cohort_dchp$f, 2, mean)))

  #### Predict M3 ####
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  # highest to lowest
  rank_m3 <- rev(order(apply(sim_cohort_mmhp$f, 2, mean)))

  #### Predict Agg Ranking ####
  load(paste(data_path,cohort_names[current_cohort],
             "/agg_rank_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  # highest to lowest
  rank_agg <- rev(order(apply(sim_agg_rank$x, 2, mean)))

  #### ISI Predict ####
  full_wl <- mini_wl_matrix(df = cbind(return_df$initiator,
                                        return_df$recipient))

  isi_train <- isi98(full_wl)
  isi_rank <- as.numeric(isi_train$best_order)

  #### glicko predict
  glick_df <- return_df %>% 
    mutate(id = row_number(), score = 1) %>%
    select(id, initiator, recipient , score)

  gl_train <- my_glicko(glick_df, history=TRUE, cval=2)

  gl_ranks <- rev(order(gl_train$ratings$Rating))

  #### save this output ####

  ### need to make sure the isi ordering is correct here

  est_ranks <- tibble(isi_full = isi_rank,
                      agg_full = rank_agg,
                      glicko_full = gl_ranks,
                      m1_full = rank_m1,
                      m2_full = rank_m2,
                      m3_full = rank_m3)

  # full_wl <- mini_wl_matrix(cbind(return_df$initiator, return_df$recipient))
  # best_isi <- isi98(full_wl)
  # this is giving me the ranks correctly but am I mapping them
  # correctly for reordering the wl matrix?

  ###
  ## do I want order here?
  ###
  scores <- est_ranks %>% map_dfr(~ compute_isi_score(full_wl, .x))

  scores$model <- c("isi","agg", "glicko", "m1","m2","m3")
  # scores <- scores %>% add_row(tibble_row(I_score = best_isi$I, 
  #                              SI_score = best_isi$SI,
  #                              model = "test_isi"))

  score_fits_all[[current_cohort]] <- scores

}

```

```{r plot these isi scores}
score_fits_all %>%
  map(~as_tibble(.)) %>%
  bind_rows(.id="index") %>%
  mutate(ISI = I_score + SI_score) %>%
  ggplot(aes(model,ISI)) + geom_boxplot()
```