### 2020-12-01 output_scripts/cmmhp_comparison/sim_comparison.Rmd

---
title: "Simulation Comparison"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: TRUE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(rstan)
library(cmdstanr)
library(compete)
library(posterior)
library(RColorBrewer)

source(here("lib", 'plotUtil.R'))
source(here("lib", 'simulatePrediction.R'))
source(here("lib", 'mmhp.R'))
source(here("lib", 'uniHawkes.R'))
source(here("lib", 'inferLatentMmhp.R'))
source(here("lib",'cleanData.R')) 
source(here("lib",'prepareDataStan.R')) 
source(here("lib",'drawIntensity.R'))
# source(here("lib",'myGlicko.R',sep = ""))
# source(here("lib",'matrixPlotParameter.R',sep = ""))
# source(here("lib",'residualStructureScore.R',sep = ""))
theme_set(theme_minimal())

dc_sim_path <- here::here("output", "sims", "sims_m3_dc")
comm_sim_path <- here::here("output", "common_rate", "sims_m3")
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')

cut_off <- 3
n_sim <- 50

model1_fn <- list(alpha.fun = function(x,y,eta1,eta2,eta3){return(eta1*x*y*exp(-eta2*abs(x-y))/(1 +
                                                        exp(-eta3*(x-y))))})

model3_fn <- list(alpha.fun = function(x,y,eta1,eta2){return(eta1*x*y*exp(-eta2*abs(x-y)))},
                  q1.fun = function(x,y,eta3){return(exp(-eta3*x))},
                  q0.fun = function(x,y,eta3){return(exp(-eta3*y))})

```

```{r get isi rank, cache = TRUE}
source(here("lib",'naiveRankHierarchy.R'))
source(here("lib",'expertRankHierarchy.R'))

full_data <- readRDS(here("data","mice.RData"))
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45
cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- utils::capture.output(for(curr_cohort in fit_cohorts){
  naive_rank_10[[curr_cohort]] <- naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <- expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})
```

Here we compare the results of two versions of our C-MMHP model, namely

- The degree corrected model, where the base rate for a given pair is
$$
\lambda_0^{i,j} = \gamma_i + \zeta_j,
$$

- The common base model, with a common base rate $\lambda_0$ across all pairs.

## Compare rank recovery

We first compare rank recovery. Here we simulate data under each of these
two models and wish to examine the rank recovery in each, relative to
the competing C-HP and C-DCHP models.

### Degree Corrected

```{r rank-recov-dc}
m3_files <- list.files(dc_sim_path, pattern = "sim_model3_fit*")

# load(paste0(m1_path,m1_files[1]))
sim_model3_stan_sim1_total <- list()
sim_model3_stan_sim2_total <- list()
sim_model3_stan_sim3_total <- list()

for(i in seq_along(m3_files)) {
  load(here(dc_sim_path, m3_files[i]))
  sim_model3_stan_sim1_total[[i]] <- sim_model3_stan_sim1
  sim_model3_stan_sim2_total[[i]] <- sim_model3_stan_sim2
  sim_model3_stan_sim3_total[[i]] <- sim_model3_stan_sim3

}

n_sim <- 50

par(mfrow=c(3, length(object_par$f_vec_1)),
    mar=c(0,0,1,0),
    oma=c(6.5,11,3.5,1))

y.ub <- c(12,12,12,12,12)
text.y <- c(rep(8,4),8)
text.x <- c(0.1,0.2,0.4,0.7,0.9)
text.x.adj <- c(0.3,0.4,0.6,0.5,0.8)
model_colors <- viridis::viridis(3)

for(m in c(1:3)){
  for(j in c(1:length(object_par$f_vec_1))){
    plot(1,4,xlim=c(0,1),ylim=c(0,y.ub[j]),
         type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)

    ## Posterior density
    # current_result <- eval(parse(text=paste("sim_model3_stan_sim",m,sep="")))
    current_result <- eval(parse(text=paste("sim_model3_stan_sim",
                                            m,
                                            "_total",
                                            sep="")))
    for(i in c(1:n_sim)){
      lines(density(current_result[[i]]$f[,j]),
            col=add.alpha(model_colors[m], alpha=0.8),cex=0.6,lwd=0.8)
    }

    ## plot true value
    abline(v=object_par$f_vec_1[j],lwd=1.8,lty=2,col="gray30")

    ## plot box
    box.wdt <- 1.5
    box.col <- "black"
    box(lty = 'solid',col=box.col,lwd=0.7)

    if(m==3){
      text(text.x.adj[j],text.y[j],text.x[j],cex=2.4)
    }
  }
  ## bottom
  mtext(text=expression(f[1],f[2],f[3],f[4],f[5]), side=1, line=2.4, outer=TRUE, 
        at=c(1:length(object_par$f_vec_1))/length(object_par$f_vec_1)-0.1,
        cex=2)
  mtext(text="Latent rank variables", side=1, line=5.3, outer=TRUE, cex=2.25,
        at=0.5)
  ## top
  # mtext(text="(a)", side=3, line=0, outer=TRUE, cex=2.5,
  #       at=0.5, font=2)
  ## left
  mtext(text=rev(c("C-HP  ","C-DCHP","C-MMHP")), side=2, line=0.5, outer=TRUE, 
        at=c(0:2)/3+0.15,
        cex=1.75, las=2)
}

```

### Common rate

We can also do this for the common rate model.
```{r rank-recov-common}
m3_files <- list.files(comm_sim_path, pattern = "sim_model3_fit*")

# load(paste0(m1_path,m1_files[1]))
sim_model3_stan_sim1_total <- list()
sim_model3_stan_sim2_total <- list()
sim_model3_stan_sim3_total <- list()

for(i in seq_along(m3_files)) {
  load(here(comm_sim_path, m3_files[i]))
  sim_model3_stan_sim1_total[[i]] <- sim_model3_stan_sim1
  sim_model3_stan_sim2_total[[i]] <- sim_model3_stan_sim2
  sim_model3_stan_sim3_total[[i]] <- sim_model3_stan_sim3

}

n_sim <- 50

par(mfrow=c(3, length(object_par$f_vec_1)),
    mar=c(0,0,1,0),
    oma=c(6.5,11,3.5,1))

y.ub <- c(12,12,12,12,12)
text.y <- c(rep(8,4),8)
text.x <- c(0.1,0.2,0.4,0.7,0.9)
text.x.adj <- c(0.3,0.4,0.6,0.5,0.8)
model_colors <- viridis::viridis(3)

for(m in c(1:3)){
  for(j in c(1:length(object_par$f_vec_1))){
    plot(1,4,xlim=c(0,1),ylim=c(0,y.ub[j]),
         type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)

    ## Posterior density
    # current_result <- eval(parse(text=paste("sim_model3_stan_sim",m,sep="")))
    current_result <- eval(parse(text=paste("sim_model3_stan_sim",
                                            m,
                                            "_total",
                                            sep="")))
    for(i in c(1:n_sim)){
      lines(density(current_result[[i]]$f[,j]),
            col=add.alpha(model_colors[m], alpha=0.8),cex=0.6,lwd=0.8)
    }

    ## plot true value
    abline(v=object_par$f_vec_1[j],lwd=1.8,lty=2,col="gray30")

    ## plot box
    box.wdt <- 1.5
    box.col <- "black"
    box(lty = 'solid',col=box.col,lwd=0.7)

    if(m==3){
      text(text.x.adj[j],text.y[j],text.x[j],cex=2.4)
    }
  }
  ## bottom
  mtext(text=expression(f[1],f[2],f[3],f[4],f[5]), side=1, line=2.4,
        outer=TRUE, 
        at=c(1:length(object_par$f_vec_1))/length(object_par$f_vec_1)-0.1,
        cex=2)
  mtext(text="Latent rank variables", side=1, line=5.3, outer=TRUE, cex=2.25,
        at=0.5)
  ## top
  # mtext(text="(a)", side=3, line=0, outer=TRUE, cex=2.5,
  #       at=0.5, font=2)
  ## left
  mtext(text=rev(c("C-HP  ","C-DCHP","C-MMHP")), side=2, line=0.5,
        outer=TRUE, 
        at=c(0:2)/3+0.15,
        cex=1.75, las=2)
}
```

## KS Statistics and Pearson Residuals

### Degree Corrected Model

```{r ks-pr-computation-dc}
object_par <- list(sim_lambda_0 = 0.08,
                   sim_lambda_1 = 0.16,
                   sim_eta_1 = 2.5,
                   gamma_var = c(0.01, 0.02, 0.03, 0.06, 0.07),
                   zeta_var = c(0.075, 0.02, 0.03, 0.05, 0.08),
                   sim_eta_2 = 0.6,
                   sim_eta_3 = 5,
                   sim_beta = 1.5,
                   f_vec_1 = c(0.1, 0.2, 0.4, 0.7, 0.9))

N_array <- array(0, dim = c(n_sim, length(object_par$f_vec_1),
                         length(object_par$f_vec_1))) 
Lambda_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1))) 
Lambda_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1)))
Lambda_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1))) 
Lambda_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1))) 
PR_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
PR_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1))) 
PR_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
PR_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                               length(object_par$f_vec_1))) 
KS_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
KS_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
KS_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))
KS_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                               length(object_par$f_vec_1)))
KS_p_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1)))
KS_p_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1)))
KS_p_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))
KS_p_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))

## Iteration on 50 simulations
for(s in c(1:n_sim)){
  load(here(dc_sim_path,paste0("sim_model3_fit123_",s,".RData")))
  load(here(dc_sim_path,paste0("fit123_state_est_",s,".RData",sep='')))
  clean_sim_data <- cleanSimulationData(raw_data=sim_model3_data, 
                                        cut_off = cut_off,
                                        N = length(object_par$f_vec_1))

  N_array[s,,] <- clean_sim_data$N_count

  ## model 1
  #########
  model1_par_est <- list(lambda0 = mean(sim_model3_stan_sim1$lambda0),
                         eta_1 = mean(sim_model3_stan_sim1$eta_1),
                         eta_2 = mean(sim_model3_stan_sim1$eta_2),
                         eta_3 = mean(sim_model3_stan_sim1$eta_3),
                         beta = mean(sim_model3_stan_sim1$beta),
                         f = apply(sim_model3_stan_sim1$f, 2, mean))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      current_t <- sim_model3_data$mmhp_matrix[i,j][[1]]$tau
      m1_object <- list(lambda0 = model1_par_est$lambda0,
                        alpha = model1_fn$alpha.fun(model1_par_est$f[i],
                                                  model1_par_est$f[j],
                                                  model1_par_est$eta_1,
                                                  model1_par_est$eta_2,
                                                  model1_par_est$eta_3),
                        beta = model1_par_est$beta)
      Lambda_model1_array[s,i,j] <- 
        uniHawkesIntegralIntensity(object = m1_object, 
                                   events = current_t,
                                   termination = tail(current_t,1))
      Lambda_test <- uniHawkesCompensator(object = m1_object, 
                                          events = current_t)
      KS_model1_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$statistic
      KS_p_model1_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      PR_model1_array[s,i,j] <- uniHawkesPearsonResidual(object = m1_object, 
                                                         events = current_t,
                                                         termination = 
                                                           tail(current_t,1))
    }
  }

  ## model2
  #############
  model2_par_est <- list(gamma = apply(sim_model3_stan_sim2$gamma, 2, mean),
                         zeta = apply(sim_model3_stan_sim2$zeta,2,mean),
                         eta_1 = mean(sim_model3_stan_sim2$eta_1),
                         eta_2 = mean(sim_model3_stan_sim2$eta_2),
                         eta_3 = mean(sim_model3_stan_sim2$eta_3),
                         beta = mean(sim_model3_stan_sim2$beta),
                         f = apply(sim_model3_stan_sim2$f, 2, mean))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      current_t <- sim_model3_data$mmhp_matrix[i,j][[1]]$tau
      m2_object <- list(lambda0 = model2_par_est$gamma[i] + 
                          model2_par_est$zeta[j],
                        alpha = model1_fn$alpha.fun(model2_par_est$f[i],
                                                    model2_par_est$f[j],
                                                    model2_par_est$eta_1,
                                                    model2_par_est$eta_2,
                                                    model2_par_est$eta_3),
                        beta = model2_par_est$beta)
      Lambda_model2_array[s,i,j] <-
        uniHawkesIntegralIntensity(object = m1_object, 
                                   events = current_t,
                                   termination = tail(current_t,1))
      Lambda_test <- uniHawkesCompensator(object = m2_object,
                                          events = current_t)
      KS_model2_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$statistic
      KS_p_model2_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      PR_model2_array[s,i,j] <- 
        uniHawkesPearsonResidual(object = m2_object, 
                                 events = current_t,
                                 termination = tail(current_t,1))
    }
  }

  ## Model3 & true model
  ######
  lambda_0_est <- apply(sim_model3_stan_sim3$lambda0, 2, mean)
  lambda_1_est <- apply(sim_model3_stan_sim3$lambda1, 2, mean)
  ## then put these into a matrix
  lam0_par_est <- matrix(0, nrow = length(object_par$f_vec_1),
                         ncol = length(object_par$f_vec_1))
  lam1_par_est <- matrix(0, nrow = length(object_par$f_vec_1),
                         ncol = length(object_par$f_vec_1))
  for(i in seq_along(lambda_0_est)) {
    row_id <- clean_sim_data$I_fit[i]
    col_id <- clean_sim_data$J_fit[i]
    lam0_par_est[row_id, col_id] <- lambda_0_est[i]
    lam1_par_est[row_id, col_id] <- lambda_1_est[i]
  }
  mmhp_par_est <- list(lambda0 = lam0_par_est,
                       lambda1 = lam1_par_est,
                       eta_1 = mean(sim_model3_stan_sim3$eta_1),
                       eta_2 = mean(sim_model3_stan_sim3$eta_2),
                       eta_3 = mean(sim_model3_stan_sim3$eta_3),
                       beta = mean(sim_model3_stan_sim3$beta),
                       f = apply(sim_model3_stan_sim3$f,2,mean))
  clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data, 
                                        cut_off = cut_off,
                                        N = length(object_par$f_vec_1))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      test.mmhp <- sim_model3_data$mmhp_matrix[i,j][[1]]
      temp.t <- test.mmhp$tau
      current.n <- length(temp.t)-1
      time.segment <- seq(0,tail(temp.t,1),length.out=10000)

      object_hat <- list(lambda0=mmhp_par_est$lambda0[i,j],
                         lambda1=mmhp_par_est$lambda1[i,j],
                         alpha=model3_fn$alpha.fun(mmhp_par_est$f[i],
                                                   mmhp_par_est$f[j],
                                                   mmhp_par_est$eta_1,
                                                   mmhp_par_est$eta_2),
                         beta=mmhp_par_est$beta,
                         q1=model3_fn$q1.fun(mmhp_par_est$f[i],
                                             mmhp_par_est$f[j],
                                             mmhp_par_est$eta_3),
                         q2=model3_fn$q0.fun(mmhp_par_est$f[i],
                                             mmhp_par_est$f[j],
                                             mmhp_par_est$eta_3))

      object_true <- lapply(object_matrix,function(x) x[i,j])
      names(object_true) <- c("lambda0","lambda1","alpha","beta","q1","q2")

      ## est Lmabda and [R]
      Lambda.test <- mmhpCompensator(params = object_hat, 
                                     t = temp.t,
                                     pzt=2 - 
                                       event_state_est_lst[i,j][[1]]$zt_v,
                                     if.pzt = FALSE)
      KS_model3_array[s,i,j] <-  ks.test(Lambda.test,"pexp")$statistic
      KS_p_model3_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      state.est.latent.mmhp <- interpolate_state_est_lst[i,j][[1]]
      step.fun.est <- stepfun(state.est.latent.mmhp$x.hat,
                              2 - state.est.latent.mmhp$z.hat)
      est.intensity <- mmhpIntensityNumeric(params = object_hat,
                                            t = temp.t[-1],
                                            time.vec = time.segment,
                                            latent.vec = 
                                              step.fun.est(time.segment))
      est.intensity.events <- mmhpIntensityAtEvents(params=object_hat,
                                                    t = temp.t,
                                  latent_z = 
                                    event_state_est_lst[i,j][[1]]$zt_v)
      Lambda_model3_array[s,i,j] <-
        sum(est.intensity)*(time.segment[2]-time.segment[1])
      PR_model3_array[s,i,j] <- sum(1/sqrt(est.intensity.events))-
        sum(sqrt(est.intensity))*(time.segment[2]-time.segment[1])

      ## Under the true model
      Lambda.test.true <- mmhpCompensator(params = object_true,
                                          t = temp.t,
                                          pzt = 2 - 
                                            test.mmhp$zt[2:length(test.mmhp$zt)],
                                          if.pzt=FALSE)
      KS_true_array[s,i,j] <-  ks.test(Lambda.test.true,"pexp")$statistic
      KS_p_true_array[s,i,j] <-  ks.test(Lambda.test.true,"pexp")$p.value
      true.intensity <- 
        mmhpTrueIntensityNumeric(params = object_true,
                                                 t = temp.t,
                                 latent = list(x = 
                                                fixStateTransition(test.mmhp)$x,
                                               z = 
                                                fixStateTransition(test.mmhp)$z),
                                     time.vec = time.segment)
      true.intensity.events <- mmhpIntensityAtEvents(params = object_true,
                                                     t = temp.t, 
                                                     latent_z = test.mmhp$zt)
      Lambda_true_array[s,i,j] <- sum(true.intensity$lambda.t)*(time.segment[2]-time.segment[1])
      PR_true_array[s,i,j] <- sum(1/sqrt(true.intensity.events))-
        sum(sqrt(true.intensity$lambda.t))*(time.segment[2]-time.segment[1])
    }
  }
}

cols_plot <- viridis::viridis(4)
#load("../output/sim_model3_fit123_residuals.RData")
par(mfrow=c(5,5),mar=c(0.2,0.2,0.2,0.2),oma=c(2.4,2.4,2.9,0.5))
for(i in c(5:1)){
  for(j in c(5:1)){
    if(i==j){
      plot(1,1,type="n",xlim=c(0,1),ylim=c(0,1),
           bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
      if(i==5){
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04, 
                col = cols_plot[1], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.22,
                col = cols_plot[2], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.44,
                col = cols_plot[3], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.66,
                col = cols_plot[4], border = NA)
        text(c(0.43,0.58,0.61,0.4), seq(0.8,0.14,-0.22),
             c("C-HP", "C-DCHP", "C-MMHP", "True"),
             cex=1.7)
      }
    }else{
      # PR <- c((N_array-Lambda_model1_array)[,i,j],(N_array-Lambda_model2_array)[,i,j],
      #         (N_array-Lambda_model3_array)[,i,j],(N_array-Lambda_true_array)[,i,j])
      PR <- c(KS_model1_array[,i,j], KS_model2_array[,i,j],
              KS_model3_array[,i,j], KS_true_array[,i,j])
      ### try pr scores instead
      # PR <- c(PR_model1_array[,i,j],PR_model2_array[,i,j],
      #         PR_model3_array[,i,j],PR_true_array[,i,j])
      Type <- c(rep(1,n_sim), rep(2,n_sim), rep(3,n_sim), rep(4,n_sim))
      boxplot(PR ~ Type,  
              col = cols_plot,
              #col = c("steelblue", "goldenrod2", "firebrick2", "chartreuse3"),   
              xlab="",ylab="",xaxt="n",yaxt="n")
      #names = c("C-HP", "C-HP-DC", "C-MMHP", "True")
    }
  }
}
mtext(text=c(1:5), side=2, line=0.5, outer=TRUE, cex=2, las=2,
      at=c(0:4)/5+0.1)
mtext(text=c(5:1), side=1, line=1, outer=TRUE, cex=2,
      at=c(0:4)/5+0.1)
mtext(text="Degree Corrected KS",side=3,line=0.5,outer=TRUE,cex=2.5,font=2)

#####
par(mfrow=c(5,5),mar=c(0.2,0.2,0.2,0.2),oma=c(2.4,2.4,2.9,0.5))
for(i in c(5:1)){
  for(j in c(5:1)){
    if(i==j){
      plot(1,1,type="n",xlim=c(0,1),ylim=c(0,1),
           bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
      if(i==5){
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04, 
                col = cols_plot[1], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.22,
                col = cols_plot[2], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.44,
                col = cols_plot[3], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.66,
                col = cols_plot[4], border = NA)
        text(c(0.43,0.58,0.61,0.4), seq(0.8,0.14,-0.22),
             c("C-HP", "C-DCHP", "C-MMHP", "True"),
             cex=1.7)
      }
    }else{
      # PR <- c((N_array-Lambda_model1_array)[,i,j],(N_array-Lambda_model2_array)[,i,j],
      #         (N_array-Lambda_model3_array)[,i,j],(N_array-Lambda_true_array)[,i,j])
      # PR <- c(KS_model1_array[,i,j], KS_model2_array[,i,j],
      #         KS_model3_array[,i,j], KS_true_array[,i,j])
      ### try pr scores instead
      PR <- c(PR_model1_array[,i,j],PR_model2_array[,i,j],
              PR_model3_array[,i,j],PR_true_array[,i,j])
      Type <- c(rep(1,n_sim), rep(2,n_sim), rep(3,n_sim), rep(4,n_sim))
      boxplot(PR ~ Type,  
              col = cols_plot,
              #col = c("steelblue", "goldenrod2", "firebrick2", "chartreuse3"),   
              xlab="",ylab="",xaxt="n",yaxt="n")
      #names = c("C-HP", "C-HP-DC", "C-MMHP", "True")
    }
  }
}
mtext(text=c(1:5), side=2, line=0.5, outer=TRUE, cex=2, las=2,
      at=c(0:4)/5+0.1)
mtext(text=c(5:1), side=1, line=1, outer=TRUE, cex=2,
      at=c(0:4)/5+0.1)
mtext(text="Degree Corrected PR",side=3,line=0.5,outer=TRUE,cex=2.5,font=2)

```

### Common rate model

```{r ks-pr-common-rate model}
complete_sims <- 1:n_sim
complete_sims <- complete_sims[-29]

N_array <- array(0, dim = c(n_sim, length(object_par$f_vec_1),
                         length(object_par$f_vec_1))) 
Lambda_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1))) 
Lambda_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1)))
Lambda_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                     length(object_par$f_vec_1))) 
Lambda_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1))) 
PR_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
PR_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1))) 
PR_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
PR_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                               length(object_par$f_vec_1))) 
KS_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
KS_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                 length(object_par$f_vec_1)))
KS_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))
KS_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                               length(object_par$f_vec_1)))
KS_p_model1_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1)))
KS_p_model2_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                   length(object_par$f_vec_1)))
KS_p_model3_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))
KS_p_true_array <- array(0, dim = c(n_sim,length(object_par$f_vec_1),
                                    length(object_par$f_vec_1)))

## Iteration on 50 simulations
for(s in 30:n_sim){
  load(here(comm_sim_path, paste0("sim_model3_fit123_",s,".RData")))
  load(here(comm_sim_path, paste0("fit123_state_est_",s,".RData",sep='')))
  clean_sim_data <- cleanSimulationData(raw_data=sim_model3_data, 
                                        cut_off = cut_off,
                                        N = length(object_par$f_vec_1))

  N_array[s,,] <- clean_sim_data$N_count

  ## model 1
  #########
  model1_par_est <- list(lambda0 = mean(sim_model3_stan_sim1$lambda0),
                         eta_1 = mean(sim_model3_stan_sim1$eta_1),
                         eta_2 = mean(sim_model3_stan_sim1$eta_2),
                         eta_3 = mean(sim_model3_stan_sim1$eta_3),
                         beta = mean(sim_model3_stan_sim1$beta),
                         f = apply(sim_model3_stan_sim1$f, 2, mean))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      current_t <- sim_model3_data$mmhp_matrix[i,j][[1]]$tau
      m1_object <- list(lambda0 = model1_par_est$lambda0,
                        alpha = model1_fn$alpha.fun(model1_par_est$f[i],
                                                  model1_par_est$f[j],
                                                  model1_par_est$eta_1,
                                                  model1_par_est$eta_2,
                                                  model1_par_est$eta_3),
                        beta = model1_par_est$beta)
      Lambda_model1_array[s,i,j] <- 
        uniHawkesIntegralIntensity(object = m1_object, 
                                   events = current_t,
                                   termination = tail(current_t,1))
      Lambda_test <- uniHawkesCompensator(object = m1_object, 
                                          events = current_t)
      KS_model1_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$statistic
      KS_p_model1_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      PR_model1_array[s,i,j] <- uniHawkesPearsonResidual(object = m1_object, 
                                                         events = current_t,
                                                         termination = 
                                                           tail(current_t,1))
    }
  }

  ## model2
  #############
  model2_par_est <- list(gamma = apply(sim_model3_stan_sim2$gamma, 2, mean),
                         zeta = apply(sim_model3_stan_sim2$zeta,2,mean),
                         eta_1 = mean(sim_model3_stan_sim2$eta_1),
                         eta_2 = mean(sim_model3_stan_sim2$eta_2),
                         eta_3 = mean(sim_model3_stan_sim2$eta_3),
                         beta = mean(sim_model3_stan_sim2$beta),
                         f = apply(sim_model3_stan_sim2$f, 2, mean))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      current_t <- sim_model3_data$mmhp_matrix[i,j][[1]]$tau
      m2_object <- list(lambda0 = model2_par_est$gamma[i] + 
                          model2_par_est$zeta[j],
                        alpha = model1_fn$alpha.fun(model2_par_est$f[i],
                                                    model2_par_est$f[j],
                                                    model2_par_est$eta_1,
                                                    model2_par_est$eta_2,
                                                    model2_par_est$eta_3),
                        beta = model2_par_est$beta)
      Lambda_model2_array[s,i,j] <-
        uniHawkesIntegralIntensity(object = m1_object, 
                                   events = current_t,
                                   termination = tail(current_t,1))
      Lambda_test <- uniHawkesCompensator(object = m2_object,
                                          events = current_t)
      KS_model2_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$statistic
      KS_p_model2_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      PR_model2_array[s,i,j] <- 
        uniHawkesPearsonResidual(object = m2_object, 
                                 events = current_t,
                                 termination = tail(current_t,1))
    }
  }

  ## Model3 & true model
  ######
  # lambda_0_est <- apply(sim_model3_stan_sim3$lambda0, 2, mean)
  # lambda_1_est <- apply(sim_model3_stan_sim3$lambda1, 2, mean)
  # ## then put these into a matrix
  # lam0_par_est <- matrix(0, nrow = length(object_par$f_vec_1),
  #                        ncol = length(object_par$f_vec_1))
  # lam1_par_est <- matrix(0, nrow = length(object_par$f_vec_1),
  #                        ncol = length(object_par$f_vec_1))
  # for(i in seq_along(lambda_0_est)) {
  #   row_id <- clean_sim_data$I_fit[i]
  #   col_id <- clean_sim_data$J_fit[i]
  #   lam0_par_est[row_id, col_id] <- lambda_0_est[i]
  #   lam1_par_est[row_id, col_id] <- lambda_1_est[i]
  # }
  mmhp_par_est <- list(lambda0 = mean(sim_model3_stan_sim3$lambda0),
                       lambda1 = mean(sim_model3_stan_sim3$lambda1),
                       eta_1 = mean(sim_model3_stan_sim3$eta_1),
                       eta_2 = mean(sim_model3_stan_sim3$eta_2),
                       eta_3 = mean(sim_model3_stan_sim3$eta_3),
                       beta = mean(sim_model3_stan_sim3$beta),
                       f = apply(sim_model3_stan_sim3$f,2,mean))
  clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data, 
                                        cut_off = cut_off,
                                        N = length(object_par$f_vec_1))

  for(i in c(1:5)){
    for(j in c(1:5)[-i]){
      test.mmhp <- sim_model3_data$mmhp_matrix[i,j][[1]]
      temp.t <- test.mmhp$tau
      current.n <- length(temp.t)-1
      time.segment <- seq(0,tail(temp.t,1),length.out=10000)

      object_hat <- list(lambda0=mmhp_par_est$lambda0,
                         lambda1=mmhp_par_est$lambda1,
                         alpha=model3_fn$alpha.fun(mmhp_par_est$f[i],
                                                   mmhp_par_est$f[j],
                                                   mmhp_par_est$eta_1,
                                                   mmhp_par_est$eta_2),
                         beta=mmhp_par_est$beta,
                         q1=model3_fn$q1.fun(mmhp_par_est$f[i],
                                             mmhp_par_est$f[j],
                                             mmhp_par_est$eta_3),
                         q2=model3_fn$q0.fun(mmhp_par_est$f[i],
                                             mmhp_par_est$f[j],
                                             mmhp_par_est$eta_3))

      object_true <- lapply(object_matrix,function(x) x[i,j])
      names(object_true) <- c("lambda0","lambda1","alpha","beta","q1","q2")

      ## est Lmabda and [R]
      Lambda.test <- mmhpCompensator(params = object_hat, 
                                     t = temp.t,
                                     pzt=2 - 
                                       event_state_est_lst[i,j][[1]]$zt_v,
                                     if.pzt = FALSE)
      KS_model3_array[s,i,j] <-  ks.test(Lambda.test,"pexp")$statistic
      KS_p_model3_array[s,i,j] <-  ks.test(Lambda_test,"pexp")$p.value
      state.est.latent.mmhp <- interpolate_state_est_lst[i,j][[1]]
      step.fun.est <- stepfun(state.est.latent.mmhp$x.hat,
                              2 - state.est.latent.mmhp$z.hat)
      est.intensity <- mmhpIntensityNumeric(params = object_hat,
                                            t = temp.t[-1],
                                            time.vec = time.segment,
                                            latent.vec = 
                                              step.fun.est(time.segment))
      est.intensity.events <- mmhpIntensityAtEvents(params=object_hat,
                                                    t = temp.t,
                                  latent_z = 
                                    event_state_est_lst[i,j][[1]]$zt_v)
      Lambda_model3_array[s,i,j] <-
        sum(est.intensity)*(time.segment[2]-time.segment[1])
      PR_model3_array[s,i,j] <- sum(1/sqrt(est.intensity.events))-
        sum(sqrt(est.intensity))*(time.segment[2]-time.segment[1])

      ## Under the true model
      Lambda.test.true <- mmhpCompensator(params = object_true,
                                          t = temp.t,
                                          pzt = 2 - 
                                            test.mmhp$zt[2:length(test.mmhp$zt)],
                                          if.pzt=FALSE)
      KS_true_array[s,i,j] <-  ks.test(Lambda.test.true,"pexp")$statistic
      KS_p_true_array[s,i,j] <-  ks.test(Lambda.test.true,"pexp")$p.value
      true.intensity <- 
        mmhpTrueIntensityNumeric(params = object_true,
                                                 t = temp.t,
                                 latent = list(x = 
                                                fixStateTransition(test.mmhp)$x,
                                               z = 
                                                fixStateTransition(test.mmhp)$z),
                                     time.vec = time.segment)
      true.intensity.events <- mmhpIntensityAtEvents(params = object_true,
                                                     t = temp.t, 
                                                     latent_z = test.mmhp$zt)
      Lambda_true_array[s,i,j] <- sum(true.intensity$lambda.t)*(time.segment[2]-time.segment[1])
      PR_true_array[s,i,j] <- sum(1/sqrt(true.intensity.events))-
        sum(sqrt(true.intensity$lambda.t))*(time.segment[2]-time.segment[1])
    }
  }
}

cols_plot <- viridis::viridis(4)
#load("../output/sim_model3_fit123_residuals.RData")
par(mfrow=c(5,5),mar=c(0.2,0.2,0.2,0.2),oma=c(2.4,2.4,2.9,0.5))
for(i in c(5:1)){
  for(j in c(5:1)){
    if(i==j){
      plot(1,1,type="n",xlim=c(0,1),ylim=c(0,1),
           bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
      if(i==5){
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04, 
                col = cols_plot[1], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.22,
                col = cols_plot[2], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.44,
                col = cols_plot[3], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.66,
                col = cols_plot[4], border = NA)
        text(c(0.43,0.58,0.61,0.4), seq(0.8,0.14,-0.22),
             c("C-HP", "C-DCHP", "C-MMHP", "True"),
             cex=1.7)
      }
    }else{
      # PR <- c((N_array-Lambda_model1_array)[,i,j],(N_array-Lambda_model2_array)[,i,j],
      #         (N_array-Lambda_model3_array)[,i,j],(N_array-Lambda_true_array)[,i,j])
      PR <- c(KS_model1_array[,i,j], KS_model2_array[,i,j],
              KS_model3_array[,i,j], KS_true_array[,i,j])
      ### try pr scores instead
      # PR <- c(PR_model1_array[,i,j],PR_model2_array[,i,j],
      #         PR_model3_array[,i,j],PR_true_array[,i,j])
      Type <- c(rep(1,n_sim), rep(2,n_sim), rep(3,n_sim), rep(4,n_sim))
      boxplot(PR ~ Type,  
              col = cols_plot,
              #col = c("steelblue", "goldenrod2", "firebrick2", "chartreuse3"),   
              xlab="",ylab="",xaxt="n",yaxt="n")
      #names = c("C-HP", "C-HP-DC", "C-MMHP", "True")
    }
  }
}
mtext(text=c(1:5), side=2, line=0.5, outer=TRUE, cex=2, las=2,
      at=c(0:4)/5+0.1)
mtext(text=c(5:1), side=1, line=1, outer=TRUE, cex=2,
      at=c(0:4)/5+0.1)
mtext(text="Common Rate KS",side=3,line=0.5,outer=TRUE,cex=2.5,font=2)

####
cols_plot <- viridis::viridis(4)
#load("../output/sim_model3_fit123_residuals.RData")
par(mfrow=c(5,5),mar=c(0.2,0.2,0.2,0.2),oma=c(2.4,2.4,2.9,0.5))
for(i in c(5:1)){
  for(j in c(5:1)){
    if(i==j){
      plot(1,1,type="n",xlim=c(0,1),ylim=c(0,1),
           bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
      if(i==5){
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04, 
                col = cols_plot[1], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.22,
                col = cols_plot[2], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.44,
                col = cols_plot[3], border = NA)
        polygon(c(0,0,0.15,0.15),c(0.72,0.8,0.8,0.72)+0.04-0.66,
                col = cols_plot[4], border = NA)
        text(c(0.43,0.58,0.61,0.4), seq(0.8,0.14,-0.22),
             c("C-HP", "C-DCHP", "C-MMHP", "True"),
             cex=1.7)
      }
    }else{
      # PR <- c((N_array-Lambda_model1_array)[,i,j],(N_array-Lambda_model2_array)[,i,j],
      #         (N_array-Lambda_model3_array)[,i,j],(N_array-Lambda_true_array)[,i,j])
      # PR <- c(KS_model1_array[,i,j], KS_model2_array[,i,j],
      #         KS_model3_array[,i,j], KS_true_array[,i,j])
      ### try pr scores instead
      PR <- c(PR_model1_array[,i,j],PR_model2_array[,i,j],
              PR_model3_array[,i,j],PR_true_array[,i,j])
      Type <- c(rep(1,n_sim), rep(2,n_sim), rep(3,n_sim), rep(4,n_sim))
      boxplot(PR ~ Type,  
              col = cols_plot,
              #col = c("steelblue", "goldenrod2", "firebrick2", "chartreuse3"),   
              xlab="",ylab="",xaxt="n",yaxt="n")
      #names = c("C-HP", "C-HP-DC", "C-MMHP", "True")
    }
  }
}
mtext(text=c(1:5), side=2, line=0.5, outer=TRUE, cex=2, las=2,
      at=c(0:4)/5+0.1)
mtext(text=c(5:1), side=1, line=1, outer=TRUE, cex=2,
      at=c(0:4)/5+0.1)
mtext(text="Common Rate PR",side=3,line=0.5,outer=TRUE,cex=2.5,font=2)

```

This indicates a similar predictive performance for both these models,
when correctly implemented.

## Rank Single Cohort

### Degree Corrected Model

```{r single cohort rank dc}
rank_cohort <- 5
current_cohort <- rank_cohort

load(paste(here("output", "/"),
           cohort_names[current_cohort],
           "/agg_rank_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "/"),
           cohort_names[current_cohort],
           "/cohort_hp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "/"),
           cohort_names[current_cohort],
           "/cohort_dchp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "/"),
           cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))

no_method <- 4
plot_latent_df <- data.frame(Methods = rep(1:no_method,
                                           each = 1000*mice_number),
                             Mice = rep(rep(1:mice_number, each = 1000),
                                        no_method),
                             force = rep(0, 1000*mice_number*no_method))
## agg
plot_latent_df$force[(1:(1000*mice_number))] <-
  as.vector(sim_agg_rank$x[1001:2000,])

## M1
plot_latent_df$force[(1:(1000*mice_number)) + 
                       1000*mice_number] <- as.vector(sim_cohort_hp$f)

## M2
plot_latent_df$force[(1:(1000*mice_number)) + 
                       2000*mice_number] <- as.vector(sim_cohort_dchp$f)

## M3
plot_latent_df$force[(1:(1000*mice_number)) + 
                       3000*mice_number] <- as.vector(sim_cohort_mmhp$f)

plot_latent_df$Mice <- factor(plot_latent_df$Mice,
                              levels = expert_rank_10[[current_cohort]])
plot_latent_df$Method <- factor(plot_latent_df$Method,
                                labels =
                                  c("AggRanking","C-HP","C-DCHP","C-MMHP"))

p1 <- ggplot(plot_latent_df, 
             aes(x = Mice, y = force, fill = Method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(4)) +
  ggtitle("Degree Corrected") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_text(size = 10),
        legend.position="none") +
  facet_grid(cols=vars(Method)) +
  xlab("Mice (ranked by I&SI method)") +
  ylab("Latent rank")
p1

```

### Common Rate Model

```{r single cohort rank common rate}
current_cohort <- rank_cohort

load(paste(here("output", "common_rate", "/"),
           cohort_names[current_cohort],
           "/agg_rank_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "common_rate", "/"),
           cohort_names[current_cohort],
           "/cohort_hp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "common_rate", "/"),
           cohort_names[current_cohort],
           "/cohort_dchp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
load(paste(here("output", "common_rate", "/"),
           cohort_names[current_cohort],
           "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))

no_method <- 4
plot_latent_df <- data.frame(Methods = rep(1:no_method,
                                           each = 1000*mice_number),
                             Mice = rep(rep(1:mice_number, each = 1000),
                                        no_method),
                             force = rep(0, 1000*mice_number*no_method))
## agg
plot_latent_df$force[(1:(1000*mice_number))] <-
  as.vector(sim_agg_rank$x[1001:2000,])

## M1
plot_latent_df$force[(1:(1000*mice_number)) + 
                       1000*mice_number] <- as.vector(sim_cohort_hp$f)

## M2
plot_latent_df$force[(1:(1000*mice_number)) + 
                       2000*mice_number] <- as.vector(sim_cohort_dchp$f)

## M3
plot_latent_df$force[(1:(1000*mice_number)) + 
                       3000*mice_number] <- as.vector(sim_cohort_mmhp$f)

plot_latent_df$Mice <- factor(plot_latent_df$Mice,
                              levels = expert_rank_10[[current_cohort]])
plot_latent_df$Method <- factor(plot_latent_df$Method,
                                labels =
                                  c("AggRanking","C-HP","C-DCHP","C-MMHP"))

p1 <- ggplot(plot_latent_df, 
             aes(x = Mice, y = force, fill = Method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(4)) +
  ggtitle("Degree Corrected") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_text(size = 10),
        legend.position="none") +
  facet_grid(cols=vars(Method)) +
  xlab("Mice (ranked by I&SI method)") +
  ylab("Latent rank")
p1
```

## Overall Rank

### Degree Corrected Model

```{r all cohorts dc}
load(here("output",
          "weighted_rank_data.RData"))
no_method <- 4
plot_cor_df <- data.frame(Method=rep(c(1:no_method),each=10),
                          cor=as.vector(spearman_df))
plot_cor_df$Method <- factor(plot_cor_df$Method,
                             labels = c("Agg-rank","C-HP","C-DCHP","C-MMHP"))
p2 <- ggplot(plot_cor_df, aes(x="", y = cor, fill = Method)) +
        geom_boxplot(alpha=0.7, width=0.5) +
        scale_fill_manual(values = viridis::viridis(4)) + 
        ggtitle("Degree Corrected Model") +
        ylim(0,1) +
        theme_bw() +
        theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
              text = element_text(size = 20),
              axis.text.y =element_text(size = 11),
              legend.position="none",
              axis.title.x=element_blank()) +
        facet_grid(cols=vars(Method)) +
        ylab("Spearman rank correlation")
p2

```

### Common Rate Model

```{r all cohorts comm rate}
load(here("output", "common_rate",
          "weighted_rank_data.RData"))
no_method <- 4
plot_cor_df <- data.frame(Method=rep(c(1:no_method),each=10),
                          cor=as.vector(spearman_df))
plot_cor_df$Method <- factor(plot_cor_df$Method,
                             labels = c("Agg-rank","C-HP","C-DCHP","C-MMHP"))
p2 <- ggplot(plot_cor_df, aes(x="", y = cor, fill = Method)) +
        geom_boxplot(alpha=0.7, width=0.5) +
        scale_fill_manual(values = viridis::viridis(4)) + 
        ggtitle("Common Rate Model") +
        ylim(0,1) +
        theme_bw() +
        theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
              text = element_text(size = 20),
              axis.text.y =element_text(size = 11),
              legend.position="none",
              axis.title.x=element_blank()) +
        facet_grid(cols=vars(Method)) +
        ylab("Spearman rank correlation")
p2
```

## Predicted Events Single Cohort

### Degree Corrected Model

```{r predict single cohort dc}
load(here("output","plot_N_predict.RData"))
current_cohort <- 5# cohort 1 looks worse than rest

temp_plot_df <- predict_day_norm_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP","C-DCHP","C-MMHP"))
temp_plot_df$day <- factor(temp_plot_df$day)
p3 <- temp_plot_df %>%
  rowwise() %>%
  #mutate(norm = min(norm,60)) %>%
  # rename(Method = method) %>%
  ggplot( aes(x = method, y = norm, fill = method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  ggtitle("(a)") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  xlab("Prediction horizon (day)") + 
  ylab("Frobenius norm of error") +
  ylim(0, 150) + # aren't infinite but >150 and <200
  NULL
p3 + labs(fill = "Method")
```

### Common Rate Model

```{r predict single cohort comm rate}
load(here("output", "common_rate", "plot_N_predict.RData"))
current_cohort <- 5# cohort 1 looks worse than rest

temp_plot_df <- predict_day_norm_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP","C-DCHP","C-MMHP"))
temp_plot_df$day <- factor(temp_plot_df$day)
#png(paste(plot_path,"real_predict_N_one_cohort.png",sep=""),
# width=600,height=300)
p3 <- temp_plot_df %>%
  rowwise() %>%
  #mutate(norm = min(norm,60)) %>%
  # rename(Method = method) %>%
  ggplot( aes(x = method, y = norm, fill = method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  ggtitle("Common Rate") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  xlab("Prediction horizon (day)") + 
  ylab("Frobenius norm of error") +
  ylim(0, 100) + # aren't infinite but >150 and <200
  NULL
p3 + labs(fill = "Method")
```

## Predicted Events All Cohorts

### Degree Corrected Model

```{r all cohorts predict dc}
load(here("output",
          "plot_N_predict.RData"))

# median prediction for each cohort here

all_cohort_box <- all_cohort_norm_df

all_cohort_box$method <- factor(all_cohort_box$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL",
                                         "C-HP","C-DCHP",
                                         "C-MMHP"))
all_cohort_box$day <- factor(all_cohort_box$day)

all_cohort_box %>% 
  ggplot(aes(x=method,y=norm, fill = method)) +
  geom_boxplot() + 
  facet_grid(cols = vars(day)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  ggtitle("Degree Corrected") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  # guides(color=guide_legend(title="Method")) +
  xlab("Prediction horizon (day)") + 
  ylab("Frobenius norm of error") +
  labs(fill = "Method") +
  # ylim(0, 80) +
  NULL
```

### Common Rate Model

```{r all cohort predict comm rate}
load(here("output",
          "common_rate",
          "plot_N_predict.RData"))

# median prediction for each cohort here

all_cohort_box <- all_cohort_norm_df

all_cohort_box$method <- factor(all_cohort_box$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL",
                                         "C-HP","C-DCHP",
                                         "C-MMHP"))
all_cohort_box$day <- factor(all_cohort_box$day)

all_cohort_box %>% 
  ggplot(aes(x=method,y=norm, fill = method)) +
  geom_boxplot() + 
  facet_grid(cols = vars(day)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  ggtitle("Common Rate") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  # guides(color=guide_legend(title="Method")) +
  xlab("Prediction horizon (day)") + 
  ylab("Frobenius norm of error") +
  labs(fill = "Method") +
  # ylim(0, 80) +
  NULL

# all_sims <- do.call(rbind,predict_day_norm_df_lst)
# all_sims <- all_cohort_norm_df
# 
# all_sims$method <- factor(all_sims$method,
#                               levels = c("mmhp","dsnl","m1","m2","m3"),
#                               labels = c("I-MMHP","DSNL",
#                                          "C-HP","C-DCHP",
#                                          "C-MMHP"))
# 
# all_sims %>%
#   ggplot(aes(x = method, y = norm, fill = method)) +
#   geom_boxplot() +
#   facet_grid(cols = vars(day)) +
#   scale_fill_manual(values = viridis::viridis(5)) +
#    ggtitle("Common Rate") +
#   theme_bw() +
#   theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
#         text = element_text(size = 20),
#         axis.text.x=element_blank(),
#         axis.ticks.x = element_blank(),
#         legend.position="right") +
#   facet_grid(cols=vars(day)) +
#   # guides(color=guide_legend(title="Method")) +
#   xlab("Prediction horizon (day)") + 
#   ylab("Frobenius norm of error") +
#   labs(fill = "Method") +
#   ylim(0, 80) +
#   NULL

```

## Predicted Rank Single Cohort

### Degree Corrected Model

```{r pred rank single cohort degree corrected}
load(here("output", "plot_rank_predict.RData"))
current_cohort <- 5
temp_plot_df <- predict_day_rank_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP",
                                         "C-DCHP","C-MMHP"))
temp_plot_df$day <- factor(temp_plot_df$day)
# png(paste(plot_path,"real_predict_rank_one_cohort.png",sep=""),
# width=600,height=300)
p5 <- ggplot(temp_plot_df, aes(x = method, y = spearman, fill = method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  #scale_fill_manual(values =
  # c("gray","yellowgreen","steelblue","goldenrod2","firebrick2")) + 
  ggtitle("Degree Corrected") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(), 
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  xlab("Prediction horizon (day)") + 
  ylab("Spearman rank correlation") +
  # ylim(0, 1) +
  NULL
p5
```

### Common Rate Model

```{r rank predict single cohort common rate}
load(here("output", "common_rate", "plot_rank_predict.RData"))
current_cohort <- 5
temp_plot_df <- predict_day_rank_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP",
                                         "C-DCHP","C-MMHP"))
temp_plot_df$day <- factor(temp_plot_df$day)
# png(paste(plot_path,"real_predict_rank_one_cohort.png",sep=""),
# width=600,height=300)
p5 <- ggplot(temp_plot_df, aes(x = method, y = spearman, fill = method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values = viridis::viridis(5)) +
  #scale_fill_manual(values =
  # c("gray","yellowgreen","steelblue","goldenrod2","firebrick2")) + 
  ggtitle("Common Rate") +
  theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(), 
        legend.position="right") +
  facet_grid(cols=vars(day)) +
  xlab("Prediction horizon (day)") + 
  ylab("Spearman rank correlation") 
  # ylim(0, 1)
p5 + labs(fill = "Method")

```

## Predicted Rank All Cohorts

### Degree Corrected Model

```{r pred rank all cohorts degree corrected}
load(here("output", "plot_rank_predict.RData"))
# temp_plot_df <- all_cohort_rank_df %>% 
#   group_by(day, method) %>% 
#   summarize(avg=mean(spearman), n= n(), sd=sd(spearman))

all_cohort_rank_df$method <- factor(all_cohort_rank_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP","C-DCHP","C-MMHP"))
all_cohort_rank_df$day <- factor(all_cohort_rank_df$day)

p6 <- ggplot(all_cohort_rank_df,
             aes(x=method, y=spearman, fill = method)) + 
  geom_boxplot(alpha = 0.7) +
  # geom_line(size=1.2, alpha=0.8, position=position_dodge(0.2)) +
  # geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=0.8, size=0.8, alpha=0.8,
  #               position=position_dodge(0.2)) +
  scale_color_manual(values = viridis::viridis(5)) +
  #scale_color_manual(values = c("gray","yellowgreen","steelblue","goldenrod2","firebrick2")) +
  ggtitle("Degree Corrected") +
  theme_bw() +
  facet_grid(cols = vars(day)) +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(), 
        legend.position="right") + 
  xlab("Prediction horizon (day)") + 
  ylab("Spearman rank correlation") + 
  coord_cartesian(ylim=c(0,1)) + 
  guides(color=guide_legend(title="Model")) 
p6

```

### Common Rate Model

```{r pred rank all cohorts comm rate}
load(here("output", "common_rate", "plot_rank_predict.RData"))
# temp_plot_df <- all_cohort_rank_df %>% 
#   group_by(day, method) %>% 
#   summarize(avg=mean(spearman), n= n(), sd=sd(spearman))

all_cohort_rank_df$method <- factor(all_cohort_rank_df$method,
                              levels = c("mmhp","dsnl","m1","m2","m3"),
                              labels = c("I-MMHP","DSNL","C-HP","C-DCHP","C-MMHP"))
all_cohort_rank_df$day <- factor(all_cohort_rank_df$day)

p6 <- ggplot(all_cohort_rank_df,
             aes(x=method, y=spearman, fill = method)) + 
  geom_boxplot(alpha = 0.7) +
  # geom_line(size=1.2, alpha=0.8, position=position_dodge(0.2)) +
  # geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=0.8, size=0.8, alpha=0.8,
  #               position=position_dodge(0.2)) +
  scale_color_manual(values = viridis::viridis(5)) +
  #scale_color_manual(values = c("gray","yellowgreen","steelblue","goldenrod2","firebrick2")) +
  ggtitle("Common Rate") +
  theme_bw() +
  facet_grid(cols = vars(day)) +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(), 
        legend.position="right") + 
  xlab("Prediction horizon (day)") + 
  ylab("Spearman rank correlation") + 
  coord_cartesian(ylim=c(0,1)) + 
  guides(color=guide_legend(title="Model")) 
p6

```

## Pearson Residuals

### Degree Corrected Model

```{r dc pr matrix}
dc_cohort <- 5
current_cohort <- dc_cohort

source(here("lib", "matrixPlotParameter.R"))

m1_pr <- readRDS(paste(here("output", "/"),
                       cohort_names[current_cohort],
                   "/chp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))

m2_pr <- readRDS(paste(here("output", "/"),
                       cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr <- readRDS(paste(here("output", "/"),
                       cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

indep_pr <- readRDS(paste(here("output", "/"),
                          cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

matrix_lst_plot <- list(
                        indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                        m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]]) 
#cat("Cohort ",current_cohort,"\n",sep="")
myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                  xLabels=expert_rank_10[[current_cohort]],
                  yLabels=rev(expert_rank_10[[current_cohort]]),
                  min=-80,max=80,axis_cex=2,title_cex = 1.8,
                  colorPalette="RdBu",if.emp=FALSE,
                  #legend.mar=c(0.5,.5,0.5,0.5),
                  title_lst=list("I-MMHP","C-HP","C-DCHP","C-MMHP"), 
                  #title_lst = list("I-MMHP"),
                  col_axis=c(-80,-40,0,40,80), fake_matrix=FALSE,
                  matrix.mar=c(2.5,2.5,2.5,1)
                  )

```

### Common Rate Model

```{r PR matrix common rate}
comm_data_path <- "../output/common_rate/"
current_cohort <- dc_cohort

source(here("lib", "matrixPlotParameter.R"))

m1_pr <- readRDS(paste(here("output","common_rate","/"),
                       cohort_names[current_cohort],
                   "/chp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))

m2_pr <- readRDS(paste(here("output","common_rate", "/"),
                       cohort_names[current_cohort],
                     "/dchp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

m3_pr <- readRDS(paste(here("output","common_rate", "/"),
                       cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

indep_pr <- readRDS(paste(here("output", "common_rate", "/"),
                          cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

matrix_lst_plot <- list(
                        indep_pr[rev(expert_rank_10[[current_cohort]]),
                                                                   expert_rank_10[[current_cohort]]],
                        m1_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m2_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]],
                        m3_pr[rev(expert_rank_10[[current_cohort]]),
                                                                 expert_rank_10[[current_cohort]]]) 
#cat("Cohort ",current_cohort,"\n",sep="")
myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                  xLabels=expert_rank_10[[current_cohort]],
                  yLabels=rev(expert_rank_10[[current_cohort]]),
                  min=-80,max=80,axis_cex=2,title_cex = 1.8,
                  colorPalette="RdBu",if.emp=FALSE,
                  #legend.mar=c(0.5,.5,0.5,0.5),
                  title_lst=list("I-MMHP","C-HP","C-DCHP","C-MMHP"), 
                  #title_lst = list("I-MMHP"),
                  col_axis=c(-80,-40,0,40,80), fake_matrix=FALSE,
                  matrix.mar=c(2.5,2.5,2.5,1)
                  )

```