### 2021-06-01 output_scripts/revisions_may.Rmd
working on plots

---
title: "Comments on Latent Ranking"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  # fig.path = "paper_figure/",
  # dev = 'png',
  # fig.height = 4.5,
  # fig.width = 7,
  warning = FALSE
)

current_cohort <- 5

# don't write to this path
data_path <- "../output/"

func_dir <- "../lib/"

library(here)
library(rstan)
library(ppdiag)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(bayesplot)
library(viridis)
library(here)
library(igraph)
library(compete)
library(circlize)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source(paste(func_dir, "circularPlot.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

cohort_names <- paste("cohort",c(9,10,12,15,16,17,18,37,38,45),sep='')
cohort_short_names <- paste("C",c(9,10,12,15,16,17,18,37,38,45),sep='')
cut_off <- 3
mice_number <- 12

full_data <- readRDS("../data/mice.RData")

fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for(curr_cohort in fit_cohorts){
  naive_rank_10[[curr_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})

theme_set(theme_minimal())

col_df <- tibble(method = c("I&SI",
                            "AggRank",
                            "DSNL",
                            "Glicko",
                            "I-MMHP",
                            "C-HP",
                            "C-DCHP",
                            "True",
                            "C-MMHP"),
            cols = viridis(9))

model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 + exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)

```

## Run some longer/larger simulations

We can replace the simulation example with 5 nodes to one with
20 nodes.

```{r simulate_20_nodes}
num_nodes <- 20
cut_off <- 3
obs_time <- 200

#### Save the simulation parameters ####

object_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)

object_par <- list(
  sim_lambda_1 = 0.2,
  sim_eta_1 = 1.5, # this has to be < beta
  gamma_var = runif(n = num_nodes, min = 0.01, max = 0.05),
  zeta_var = runif(n = num_nodes, min = 0.01, max = 0.05),
  sim_eta_2 = 0.6,
  sim_eta_3 = 3,
  sim_beta = 2,
  f_vec_1 = seq(
    from = 0.05, to = 0.95,
    length.out = num_nodes
  )
)

object_matrix <- list(
  lambda0_matrix = outer(
    object_par$gamma_var,
    object_par$zeta_var, "+"
  ),
  lambda1_matrix = matrix(object_par$sim_lambda_1,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  ),
  alpha_matrix = formMatrix(
    function(x, y) {
      object_fn$alpha.fun(
        x, y, object_par$sim_eta_1,
        object_par$sim_eta_2
      )
    },
    object_par$f_vec_1
  ),
  beta_matrix = matrix(object_par$sim_beta,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  ),
  q1_matrix = formMatrix(
    function(x, y) {
      object_fn$q1.fun(
        x, y,
        object_par$sim_eta_3
      )
    },
    object_par$f_vec_1
  ),
  q2_matrix = formMatrix(
    function(x, y) {
      object_fn$q0.fun(
        x, y,
        object_par$sim_eta_3
      )
    },
    object_par$f_vec_1
  )
)

## Simulate
sim_model3_data <- list()
N_array <- array(0, c(1, num_nodes, num_nodes))
# for(i in c(1:n_sim)){
sim_model3_data <- simulateLatentMMHP(
  lambda0_matrix =
    object_matrix$lambda0_matrix,
  lambda1_matrix =
    object_matrix$lambda1_matrix,
  alpha_matrix = object_matrix$alpha_matrix,
  beta_matrix = object_matrix$beta_matrix,
  q1_matrix = object_matrix$q1_matrix,
  q2_matrix = object_matrix$q2_matrix,
  horizon = obs_time
)
clean_sim_data <- cleanSimulationData(
  raw_data = sim_model3_data,
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)
N_array <- clean_sim_data$N_count
```

We wish to confirm that this simulation study actually results
in sensible, such as whether the I&SI is consistent with
the true latent rankings.

We can also plot the MMHP fit for a single pair using `ppdiag`.

```{r plot_pair}
start <- 20
end <- 10

poss_start <- which(clean_sim_data$I_fit == start)
poss_end <- which(clean_sim_data$J_fit == end)

pair_id <- intersect(poss_start, poss_end)

events <- clean_sim_data$event_matrix[pair_id, ]
events <- events[events > 0]

q1 <- object_matrix$q1_matrix[start, end]
q2 <- object_matrix$q2_matrix[start, end]

obj <- pp_mmhp(
  lambda0 = object_matrix$lambda0_matrix[start, end],
  lambda1 = object_matrix$lambda1_matrix[start, end],
  alpha = object_matrix$alpha_matrix[start, end],
  beta = object_matrix$beta_matrix[start, end],
  Q = matrix(c(-q1, q1, q2, -q2), nrow = 2, ncol = 2, byrow = TRUE)
)

pp_residual(obj, events, end = obs_time)

# can't plot the mmhp unless you infer the latent process
### can use viterbi and such for now

obj$q1 <- q1
obj$q2 <- q2

viterbi_result <- myViterbi(
  events = events,
  param = obj,
  termination = obs_time
)

latent_inter <- interpolateLatentTrajectory(obj,
  events,
  viterbi_result$zt_v,
  initial.state =
    viterbi_result$initial_state,
  termination.time = obs_time,
  termination.state =
    viterbi_result$termination_state
)

event_state <- ppdiag:::mmhp_event_state(params = obj, events = events)

latent_state <- ppdiag:::interpolate_mmhp_latent(
  params = obj, events,
  zt = event_state$zt
)

mmhp_latent <- list()
mmhp_latent$start <- 0
mmhp_latent$end <- obs_time
mmhp_latent$x <- c(0, latent_state$x.hat, obs_time)
mmhp_latent$z <- c(viterbi_result$initial_state, latent_state$z.hat)
mmhp_latent$events <- c(0, events)
mmhp_latent$zt <- c(1, event_state$zt)
ppdiag::drawUniMMHPIntensity(obj, mmhp_latent)

# this plot is not correct for later Hawkes states...
## don't think the bug is with ppdiag
```

```{r ranking_sim_data, include=FALSE, eval=FALSE}
### i&si method
count_data_dc <- get_wl_matrix(df = cbind(
  clean_sim_data$start,
  clean_sim_data$end
))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
as.numeric(isi_dc.out$best_order) # from highest to lowest

### agg rank method
agg_rank_data <- clean_sim_data$N_count
agg_rank_model <- stan_model(here("lib", "latent_rank_agg_sim.stan"))

agg_rank_fit <- rstan::sampling(agg_rank_model,
  data = list(
    n = num_nodes,
    n_matrix = agg_rank_data
  ),
  iter = 1000,
  chains = 4
)
agg_sims <- rstan::extract(agg_rank_fit)

agg_rank <- order(apply(agg_sims$x, 2, mean))
rev(agg_rank) # from highest to lowest

### then glicko
glicko_data <- tibble(
  start = clean_sim_data$start,
  end = clean_sim_data$end
)

glicko_data <- glicko_data %>%
  mutate(id = row_number(), score = 1) %>%
  select(id, start, end, score)

library(PlayerRatings)
gl_train <- my_glicko(glicko_data, history = TRUE, cval = 2)

gl_train

gl_ranks <- order(gl_train$ratings$Rating)
gl_ranks # highest to lowest
```

So there is some structure in this data here that is perhaps 
best picked up by the I&SI method.

Similarly, we can confirm we get sensible fits by fitting the individual
MMHP model to each pair and showing the residuals. This
is consistent with what we see for real data, where the
residual matrices show overall small positive values, indicating
underestimation when we fit individual models.

```{r residuals-immhp-sim}
pr_matrix <- readRDS(here("output", "revisions", "pr_matrix_immhp.RDS"))
rr_matrix <- readRDS(here("output", "revisions", "rr_matrix_immhp.RDS"))
# image(pr_matrix)

matrix_lst_plot <- list()
matrix_lst_plot[[1]] <- rr_matrix
matrix_lst_plot[[2]] <- pr_matrix

myMultiMatrixPlot(
  X = matrix_lst_plot,
  no_matrix = 2,
  n_row = 1,
  xLabels = 1:20,
  yLabels = 20:1,
  min = -30,
  max = 30,
  axis_cex = 2,
  title_cex = 1.8,
  colorPalette = "RdBu",
  if.emp = FALSE,
  # legend.mar=c(0.5,.5,0.5,0.5),
  title_lst = list("Raw", "Pearson"),
  # title_lst = list("I-MMHP"),
  col_axis = c(-30, -15, 0, 15, 30),
  fake_matrix = TRUE,
  matrix.mar = c(2.5, 2.5, 2.5, 1)
)
```

## Explaining the g function

```{r plot_g}
alpha.fun <- function(x, y, eta1, eta2) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)))
}

f_vec <- c(0.1, 0.2, 0.4, 0.8, 0.9)

eta1 <- 6.6
eta2 <- 0.37

tidyr::crossing(send = f_vec, rec = f_vec) %>%
  filter(send != rec) %>%
  mutate(alpha = alpha.fun(send, rec, eta1, eta2))
```

## Larger Simulation Study

```{r load_in_large_sim}
load(here("output", "revisions", "sim_model3_immhp_.RData"))
load(here("output", "revisions", "sim_model3_fit_cmmhp_.RData"))

post_rank <- post_draws %>% 
  select(starts_with("f")) %>% 
  ## rename the f cols first
  rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
  pivot_longer(everything(),
               names_to = "node",
               values_to = "rank") 
  # group_by(node) %>% 
  # summarise(ave_rank = mean(rank), sd_rank = sd(rank)) %>%

## order nodes
# post_rank$node <- as.factor(post_rank$node)
true_rank <- tibble(names = paste0("", 1:20),
                    f_vec = seq(from = 0.05,
                                  to = 0.95,
                                  length.out = 20),
                    value = "True Value")

post_rank$node <- factor(post_rank$node, levels = paste0("", 1:20))

post_mean <- post_rank %>%
  group_by(node) %>% 
  summarise(ave_rank = mean(rank))

post_rank %>% 
  ggplot(aes(node, rank)) + 
  geom_boxplot() +
  geom_point(data = true_rank,
             mapping = aes(names, f_vec, colour = value)) +
  # geom_point(data = post_mean,
  #            mapping = aes(node, ave_rank, shape = 16),
  #            colour = "blue") +
  # scale_shape_identity() +
  labs(title = "Inferred Latent Ranking with Truth",
       x = "Node", y = "Latent Ranking",
       colour = "Summary") +
  theme(legend.title = element_blank(),
        legend.position = "bottom")
## blue is posterior mean, red is true value
## can add legend to fix this
```

What else should we infer with a larger simulation example? How do we visualise 
the results in this case?

## Stability of Rankings over Time

```{r load_cohort_rankings}
for(cohort_id in 1:10){
  start_rank <- readRDS(file = paste(data_path,
                                 "/revisions/rank_stab/", "cohort_",
                                 cohort_id, "start.RDS",
                                 sep = ""))

  end_rank <- readRDS(file = paste(data_path,
                                   "/revisions/rank_stab/", "cohort_",
                                   cohort_id, "end.RDS",
                                   sep = ""))
  ### plot both of these
  init_ranks <- start_rank %>%
    select(starts_with("f")) %>%
    rename_with(~gsub("[", "_", .x, fixed = TRUE)) %>%
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  final_ranks <- end_rank %>%
    select(starts_with("f")) %>%
    rename_with(~gsub("[", "_", .x, fixed = TRUE)) %>%
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  init_ranks$node <- factor(init_ranks$node, levels = paste0("f_", 1:12))
  final_ranks$node <- factor(final_ranks$node, levels = paste0("f_", 1:12))

  init_ranks$when <- "first14"
  final_ranks$when <- "last14"

  init_ranks %>% 
    ggplot(aes(node, rank)) + 
    geom_boxplot() +
    labs(title = "Ranking from first 14 Days")

  final_ranks %>% 
    ggplot(aes(node, rank)) + 
    geom_boxplot() +
    labs(title = "Ranking from final 14 Days")

  all_ranks <- bind_rows(init_ranks, final_ranks)

  coh_plot <- all_ranks %>% 
    ggplot(aes(node, rank, fill = when)) +
    geom_boxplot() +
    labs(x = "Mouse",
         y = "Latent Rank", 
         fill = "Time Period",
         title = "Agreement between initial and later ranking",
         subtitle = "One Cohort")
  print(coh_plot)
}

```

## Community Structure Plot

```{r construct_plot, include=FALSE, eval=FALSE}
iAndSI <- function(M) {
  n <- ncol(M)
  result <- rep(0, 2)
  k <- (M - t(M)) / 2
  k[upper.tri(k)] <- 0
  result[1] <- 0
  a <- length(which(k > 0))
  y <- which(k > 0) %% n
  x <- (which(k > 0) - 1) %/% n + 1
  y[y == 0] <- y[y == 0] + n
  if (a > 0) {
    result[1] <- a
    result[2] <- sum(y - x)
  }
  return(result)
}

state_types <- c("total", "utility", "social")
metric_types <- c("dc", "ttri", "i")
state_separation_result <- data.frame(array(0,
                                            c(10 * 
                                                length(state_types) *
                                                length(metric_types),
                                              4)))
colnames(state_separation_result) <- c("cohort", "state", "metric", "value")

total_event_array_lst <- list()
active_event_array_lst <- list()
for (current_cohort in c(1:10)) {
  print(current_cohort)
  load(paste(data_path, 
             cohort_names[current_cohort],
    "/cmmhp_est_zt_",
    cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(
    current_cohort, 
    raw_df =  full_data[[cohort_names[current_cohort]]],
    clean_data)
  unique_pairs_df <- return_df %>%
    group_by(initiator, recipient) %>%
    dplyr::summarize(
      count = n(),
      observe = list(observe.id),
      observe.length = list(observe.time),
      no.events = list(no.events)
    )
  unique_observe_win <- unique(return_df[, c("observe.id", "observe.time")])

  expert_rank <- expert_rank_10[[current_cohort]]

  # -------- Load the saved stan result, 
  # -------- plot the inferred parameters and save
  total_event_array_lst[[current_cohort]] <- 
    array(0, dim = c(mice_number, mice_number, max(return_df$observe.id)))
  active_event_array_lst[[current_cohort]] <-
    array(0, dim = c(mice_number, mice_number, max(return_df$observe.id)))

  for (i in 1:mice_number) {
    for (j in 1:mice_number) {
      pair <- which(unique_pairs_df$initiator == i & 
                      unique_pairs_df$recipient == j)
      if (length(pair) > 0 & i!=j) {
        current_window_vec <- unique_pairs_df$observe[[pair]]
        for (cur_win in current_window_vec) {
          row_indicator <- (return_df$initiator == i) & 
            (return_df$recipient == j) &
            (return_df$observe.id == cur_win)
          total_event_array_lst[[current_cohort]][i, j, cur_win] <-
            length(return_df[row_indicator, "event.times"][[1]])
          active_event_array_lst[[current_cohort]][i, j, cur_win] <- 
            sum(apply(2 - state_array_list[[pair]][[cur_win]], 1, mean) > 0.5)
        }
      }
    }
  }

  start_win <- floor(max(return_df$observe.id) / 2)
  end_win <- max(return_df$observe.id)
  matrix1 <- apply(
    total_event_array_lst[[current_cohort]][, , start_win:end_win],
    c(1, 2),
    sum
    )[expert_rank, expert_rank]
  matrix2 <- apply(
    active_event_array_lst[[current_cohort]][, , start_win:end_win],
    c(1, 2),
    sum
    )[expert_rank, expert_rank]
  matrix3 <- apply(
    total_event_array_lst[[current_cohort]][, , start_win:end_win] -
      active_event_array_lst[[current_cohort]][, , start_win:end_win],
    c(1, 2),
    sum
    )[expert_rank, expert_rank]

  # ------ compute metrics ------
  state_separation_result$cohort[
    ((current_cohort - 1) * 
      length(metric_types) *
      length(state_types) + 1):(current_cohort * length(metric_types) *
                                  length(state_types))] <-
    cohort_names[current_cohort]
  state_separation_result$metric[
    ((current_cohort - 1) *
       length(metric_types) *
       length(state_types) + 1):(current_cohort * length(metric_types) *
                                   length(state_types))] <- 
    rep(c("dc", "ttri", "i"), length(state_types))

  state_separation_result$state[
    ((current_cohort - 1) *
       length(metric_types) *
       length(state_types) + 1):(current_cohort * length(metric_types) *
                                   length(state_types) -
                                   length(metric_types) * 2)] <- "total"

  ## bug below here need to find iAndSI
  state_separation_result$value[
    ((current_cohort - 1) *
       length(metric_types) *
       length(state_types) + 1):(current_cohort * length(metric_types) *
                                   length(state_types) - 
                                   length(metric_types) * 2)] <- 
    c(dci(matrix1), ttri(matrix1)[[2]], iAndSI(matrix1)[1])

  state_separation_result$state[
    ((current_cohort - 1) *
       length(metric_types) *
       length(state_types) + 
       length(metric_types) + 1):(current_cohort * 
                                    length(metric_types) *
                                    length(state_types) -
                                    length(metric_types))] <-
    "utility"
  state_separation_result$value[
    ((current_cohort - 1) *
       length(metric_types) * 
       length(state_types) + 
       length(metric_types) + 1):(current_cohort * length(metric_types)
                                  * length(state_types)
                                  - length(metric_types))] <-
    c(dci(matrix2), ttri(matrix2)[[2]], iAndSI(matrix2)[1])

  state_separation_result$state[(current_cohort * length(metric_types) *
                                   length(state_types) - 
                                   length(metric_types) + 1):(
                                     current_cohort * length(metric_types) *
                                       length(state_types))] <- "social"
  state_separation_result$value[(current_cohort * length(metric_types) *
                                   length(state_types) -
                                   length(metric_types) + 1):(
                                     current_cohort * length(metric_types) *
                                       length(state_types))] <- 
    c(dci(matrix3), ttri(matrix3)[[2]], iAndSI(matrix3)[1])
  rm(interpolation_array_list)
  rm(state_array_list)
}

save(state_types, metric_types, state_separation_result,
  total_event_array_lst, active_event_array_lst,
  file = "../output/state_separation_plot.RData"
)

```

```{r recreate_jing_figure, message=FALSE}
load("../output/state_separation_plot.RData")
# for (current_cohort in c(1)) {
  current_cohort <- 9
  total_event_array <- total_event_array_lst[[current_cohort]]
  active_event_array <- active_event_array_lst[[current_cohort]]
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(current_cohort, 
                                      full_data[[cohort_names[current_cohort]]],
                                      clean_data)

  start_win <- floor(max(return_df$observe.id) / 3) * (0:2) + 1
  end_win <- floor(max(return_df$observe.id) / 3) * (1:3)
  matrix1 <- apply(
    total_event_array[, , start_win[1]:end_win[1]] -
      active_event_array[, , start_win[1]:end_win[1]],
    c(1, 2), sum
  )
  matrix2 <- apply(
    total_event_array[, , start_win[2]:end_win[2]] -
      active_event_array[, , start_win[2]:end_win[2]],
    c(1, 2), sum
  )
  matrix3 <- apply(
    total_event_array[, , start_win[3]:end_win[3]] -
      active_event_array[, , start_win[3]:end_win[3]],
    c(1, 2), sum
  )
  ### work to here
  g_community1 <- graph_from_adjacency_matrix(matrix1, mode = "undirected")
  lec1 <- cluster_leading_eigen(g_community1)

  g_community2 <- graph_from_adjacency_matrix(matrix2, mode = "undirected")
  lec2 <- cluster_leading_eigen(g_community2)

  g_community3 <- graph_from_adjacency_matrix(matrix3, mode = "undirected")
  lec3 <- cluster_leading_eigen(g_community3)

  ### Make plot for t=1
  # png(paste(plot_path, "real_clustering_1.png", sep = ""), 
  #     height = 600,
  #     width = 600)
  myCircularPlot(
    cur.matrix = matrix1,
    ## is the ordering here for the nodes? need to make sure I get this right
    ## working off this being the ordering from the clustering above
    # cur.order = c(1, 4, 10, 5, 11, 6, 7, 2, 8, 12, 9, 3),
    cur.order = order(lec1$membership),
    cur.color = c(
      colorRampPalette(brewer.pal(
        name = "YlOrBr",
        n = 8))(14)[c(13, 11, 9, 7)],
      colorRampPalette(brewer.pal(
        name = "PuRd",
        n = 8))(13)[c(4, 6, 8, 10, 12)],
      colorRampPalette(brewer.pal(
        name = "Blues",
        n = 8))(15)[c(4, 8, 12)]
    ),
    cur.gap = c(rep(5, 3), 30, rep(5, 4), 30, rep(5, 2), 30)
  )
  # dev.off()

  ### Make plot for t=2
 #  png(paste(plot_path, "real_clustering_2.png", sep = ""),
 #   height = 600,
 #   width = 600
 # )
 # problems with these 2, possibly if no events from 11?
 circos.par$cell.padding <- c(0,0,0,0)
 myCircularPlot(
   cur.matrix = matrix2[-11,-11],
   cur.order = order(lec2$membership)[-6],
   cur.color = c(
     colorRampPalette(brewer.pal(
       name = "YlOrBr",
       n = 8))(16)[c(15, 13, 11, 9, 7, 5)],
     colorRampPalette(brewer.pal(
       name = "Blues",
       n = 8))(15)[c(6, 8, 10, 12, 14)]
   ),
   cur.gap = c(rep(5, 5), 30, rep(5, 4), 30),
   cur.ID = c(1:12)[-11]
 )
 # dev.off()

  ### Make plot for t=3
  # png(paste(plot_path, "real_clustering_3.png", sep = ""),
  #     height = 600,
  #     width = 600)
  myCircularPlot(
    cur.matrix = matrix3[-11, -11],
    cur.order = order(lec3$membership)[-4],
    cur.color = c(
      colorRampPalette(
        brewer.pal(name = "YlOrBr",
                   n = 8))(14)[c(13, 9, 5, 3)],
      colorRampPalette(brewer.pal(
        name = "Blues",
        n = 8))(16)[c(9, 11, 13, 15)],
      colorRampPalette(brewer.pal(
        name = "PuRd",
        n = 8))(15)[c(4, 8, 12)]
    ),
    cur.gap = c(rep(5, 3), 30, rep(5, 3), 30, rep(5, 2), 30),
    cur.ID = c(1:12)[-11]
  )
  # dev.off()
# }
```

# Repeat Rank Simulations

```{r rank-with-20-nodes}
sim_files <- list.files(here("output", "revisions", "rank_sims"), 
                        pattern = "rank_sim")

all_sims <- list()
for(i in seq_along(sim_files)) {
 single_fit <- readRDS(here("output",
                            "revisions",
                            "rank_sims",
                            sim_files[i]))
 all_sims[[i]] <- single_fit
}

spear <- function(sim_data) {
  truth <- sim_data$truth
  sim_data %>%
    select(m1:glicko) %>%
    purrr::map( ~cor.test(truth, .x,
                   method = "spearman")$estimate  ) %>%
    enframe() %>%
    unnest(cols = c(value))
}

sim_data <- map_dfr(all_sims, spear)

name <- sim_data$name

sim_data$name <- case_when(
          name == "isi" ~ "I&SI",
          name == "m1" ~ "C-HP",
          name == "m2" ~ "C-DCHP",
          name == "m3" ~ "C-MMHP",
          name == "agg" ~ "AggRank",
          name == "glicko" ~ "Glicko")

sim_data$name <- factor(sim_data$name, levels = c("I&SI",
                                                  "AggRank",
                                                  "Glicko",
                                                  "C-HP",
                                                  "C-DCHP",
                                                  "C-MMHP"))

plot_cols <- col_df %>% filter(method %in% levels(sim_data$name)) %>%
  pull(cols)

sim_data %>%
  rename(Method = name, corr = value) %>%
  ggplot(aes(Method,corr)) + 
  geom_boxplot(aes(fill = Method), alpha = 0.7) +
  scale_fill_manual(values = plot_cols) +
  labs(y = "Rank Correlation") + theme_bw() +
  theme(plot.title = element_text(size = 22, hjust = 0.5, face="bold"),
        text = element_text(size = 20),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        legend.position="right") + ggtitle("(b)")

```

# Improve Fig 4b

Trying to write a better version of Figure 4b using these new larger
simulated networks.

```{r process-new-4b}
plot.s <- 1 # 7 is pretty good
plot.i <- 20
plot.j <- 15

cur_i <- plot.i
cur_j <- plot.j

load(paste0(here("output", "revisions", "sim_m3"),
  "/", "sim_model3_fit123_",
  plot.s,
  ".RData"
))

test.mmhp <- sim_model3_data$mmhp_matrix[plot.i, plot.j][[1]]
temp.t <- test.mmhp$tau
current.n <- length(temp.t) - 1
time.segment <- seq(0, tail(temp.t, 1), length.out = 5000)

### Preprocess the model inference result
## model 1
#########

## get f from cmdstan

f_draws <- sim_model3_stan_sim1 %>% 
  select(starts_with("f"))

model1_par_est <- list(
  lambda0 = mean(sim_model3_stan_sim1$lambda0),
  eta_1 =
    mean(sim_model3_stan_sim1$eta_1),
  eta_2 =
    mean(sim_model3_stan_sim1$eta_2),
  eta_3 =
    mean(sim_model3_stan_sim1$eta_3),
  beta = mean(sim_model3_stan_sim1$beta),
  f = apply(f_draws, 2, mean)
)
lambda.m1 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 =
      model1_par_est$lambda0,
    alpha =
      model1_fn$alpha.fun(
        model1_par_est$f[plot.i],
        model1_par_est$f[plot.j],
        model1_par_est$eta_1,
        model1_par_est$eta_2,
        model1_par_est$eta_3
      ),
    beta = model1_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## model2
#############
gamma_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("gamma"))

zeta_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("zeta"))

f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

model2_par_est <- list(
  gamma = apply(gamma_draws, 2, mean),
  zeta = apply(zeta_draws, 2, mean),
  eta_1 = mean(sim_model3_stan_sim2$eta_1),
  eta_2 = mean(sim_model3_stan_sim2$eta_2),
  eta_3 = mean(sim_model3_stan_sim2$eta_3),
  beta = mean(sim_model3_stan_sim2$beta),
  f = apply(f_draws, 2, mean)
)

lambda.m2 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 = model2_par_est$gamma[plot.i] +
      model2_par_est$zeta[plot.j],
    alpha = model1_fn$alpha.fun(
      model2_par_est$f[plot.i],
      model2_par_est$f[plot.j],
      model2_par_est$eta_1,
      model2_par_est$eta_2,
      model2_par_est$eta_3
    ),
    beta = model2_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## Model3
###

lam0_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("gamma"))

lam1_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("zeta"))

gamma_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("gamma"))

zeta_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("zeta"))

f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

lam0_vec <- apply(lam0_draws, 2, mean)
lam1_vec <- apply(lam1_draws, 2, mean)
lam0_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
lam1_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)

clean_sim_data <- cleanSimulationData(
  raw_data = sim_model3_data,
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)

for (i in seq_along(lam0_vec)) {
  row_id <- clean_sim_data$I_fit[i]
  col_id <- clean_sim_data$J_fit[i]
  lam0_par_est[row_id, col_id] <- lam0_vec[i]
  lam1_par_est[row_id, col_id] <- lam1_vec[i]
}

###

#############
mmhp_par_est <- list(
  lambda0 = lam0_par_est,
  lambda1 = lam1_par_est,
  eta_1 = mean(sim_model3_stan_sim3$eta_1),
  eta_2 = mean(sim_model3_stan_sim3$eta_2),
  eta_3 = mean(sim_model3_stan_sim3$eta_3),
  beta = mean(sim_model3_stan_sim3$beta),
  f = apply(f_draws, 2, mean)
)

object_hat <- list(
  lambda0 = mmhp_par_est$lambda0[cur_i, cur_j],
  lambda1 = mmhp_par_est$lambda1[cur_i, cur_j],
  alpha = model3_fn$alpha.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_1,
    mmhp_par_est$eta_2
  ),
  beta = mmhp_par_est$beta,
  q1 = model3_fn$q1.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  ),
  q2 = model3_fn$q0.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  )
)

## then need to load in the state estimate here
load(paste0(here("output", "revisions", "sim_m3"),
  "/", "fit123_state_est_",
  plot.s,
  ".RData"
))

state.est.latent.mmhp <- interpolate_state_est_lst[plot.i, plot.j][[1]]
state.est.latent.mmhp.new <- fixInterpolateState(state.est.latent.mmhp,
  termination = 200
)

step.fun.est <- stepfun(
  state.est.latent.mmhp$x.hat,
  2 - state.est.latent.mmhp$z.hat
)

lambda.m3 <- mmhpIntensityNumeric(
  params = object_hat,
  t = temp.t[-1],
  time.vec = time.segment,
  latent.vec = step.fun.est(time.segment)
)

object_true <- lapply(object_matrix, function(x) x[plot.i, plot.j])
names(object_true) <- c("lambda0", "lambda1", "alpha", "beta", "q1", "q2")

lambda.true <- mmhpTrueIntensityNumeric(
  params = object_true,
  t = temp.t,
  latent = list(
    x = fixStateTransition(test.mmhp)$x,
    z = fixStateTransition(test.mmhp)$z
  ),
  time.vec = time.segment
)

```

```{r fig4b-redone}
y.ub <- c(35,35,35)#c(130,72,31)
x_events <- 190#135#150
layout(matrix(c(1:6), 6, 1), heights=rep(c(7.5,1),3))
my.xlim <- 200#160#200#130
my.long.x <- 205#205#134.5
legend.x <- c(60,60,60)#c(135,135,135)
legend.y <- #c(34,32,32)
legend.y <- c(23,23,23)#c(15,12,15)#c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 #3
line.alpha <- c(0.9,0.6)
line.wdth <- c(2,2.5)
negative.col <- add.alpha('lightskyblue')
positive.col <- add.alpha('tomato')
#negative.col <- viridis::viridis(5,alpha = 0.6)[1]
#positive.col <- viridis::viridis(5,alpha = 0.6)[4]
par(mar = c(0,0.3,0,0.1), oma = c(3,3.5,4,0),
    tcl=0.2,mgp=c(0.5,0,0), xpd=TRUE)

model_colors <- col_df %>% filter(method %in% c("C-HP",
                                                "C-DCHP",
                                                "C-MMHP"))

model_colors <- model_colors %>% pull(cols)

# model1
############
true.object <- object_true

drawUniMMHPIntensityPaper(true.object,
                          simulation = fixStateTransition(test.mmhp),
                          yupper=y.ub[1], 
                          color=add.alpha("black", alpha=line.alpha[1]),
                          line.width=line.wdth[1],
                          y.ratio=-0.05, min.y=-2,
                          min.x=0, max.x=my.xlim,
                          title_name=paste(plot.i,"->",plot.j),
                          title.cex = 2,
                          box.type="n")
drawHawkesIntensityPaper(lambda0=model1_par_est$lambda0,
                         alpha=model1_fn$alpha.fun(model1_par_est$f[plot.i],
                                                   model1_par_est$f[plot.j],
                                                   model1_par_est$eta_1,
                                                   model1_par_est$eta_2,
                                                   model1_par_est$eta_3),
                         beta=model1_par_est$beta,
                         events=test.mmhp$tau,
                         color=add.alpha(model_colors[1],
                                         alpha=line.alpha[2]),
                         line.width=line.wdth[2])

## draw box
axis(1,at=c(-2,my.long.x),col="gray35",
     line=0.5,tick=T,labels=rep("",2),lwd=0.5,lwd.ticks=0,lty=3)

## add legend
# changed 34 to 23, 25 to 18, -20 to -10
# legend(10,legend.y[1],c("State 1/0 events","State change point"),
#        col = c(NA,"red"),
#        y.intersp=0.85,x.intersp=-0.1, bty = "n",
#        pch = c(NA,4), pt.cex = c(NA,2), cex=legend.cex,
#        lty = c(NA,NA), lwd=c(NA,4))
# points(29,84,pch=1,cex=2,col="blue")
# points(26,84,pch=16,cex=2,col="blue")
legend(legend.x[1],legend.y[1],c("True"),
       col = c("black"),
       y.intersp=0.88,x.intersp=0.15,bty = "n",
       lty = c(1), lwd = c(5), cex=legend.cex,seg.len=1)
legend(legend.x[1]+22,legend.y[1],c("C-HP"),
       col = c(model_colors[1]),
       y.intersp=0.88,x.intersp=0.15,bty = "n",
       lty = c(1), lwd = c(5), cex=legend.cex,seg.len=1)

## delta lambda
par(mar = c(0,0.3,0,0.1))
plot(0,0,xlim=c(0,my.xlim), ylim=c(-4,4), type="n",
     bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
delta.lambda <- lambda.m1$lambda.t-lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive<0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative>0] <- 0
### hiding these for now ###
polygon(c(0,lambda.true$time.vec,x_events),#200
        c(0,delta.negative,0),col=negative.col,border=NA)
polygon(c(0,lambda.true$time.vec,x_events),#200
        c(0,delta.positive,0),col=positive.col,border=NA)

axis(1,at=c(-2,my.long.x),col="gray35",
     line=0,tick=T,labels=rep("",2),lwd=0.7,lwd.ticks=0)

# model2
############
drawUniMMHPIntensityPaper(true.object,simulation=fixStateTransition(test.mmhp),
                          yupper=y.ub[2], 
                          color=add.alpha("black", alpha=line.alpha[1]),
                          line.width=line.wdth[1],
                          y.ratio=-0.05, min.y=-2, min.x=0, max.x=my.xlim, 
                          box.type="n")
drawHawkesIntensityPaper(lambda0=model2_par_est$gamma[plot.i]+
                           model2_par_est$zeta[plot.j],
                         alpha=model1_fn$alpha.fun(model2_par_est$f[plot.i],
                                                   model2_par_est$f[plot.j],
                                                   model2_par_est$eta_1,
                                                   model2_par_est$eta_2,
                                                   model2_par_est$eta_3),
                         beta=model2_par_est$beta,
                         events=test.mmhp$tau,
                         color=add.alpha(model_colors[2],
                                         alpha=line.alpha[2]),
                         line.width=line.wdth[2])
legend(legend.x[2],legend.y[2],"C-DCHP",
       col = model_colors[2],
       y.intersp=0.88,x.intersp=0.1,bty = "n",
       lty = 1, lwd = 5, cex= legend.cex,seg.len=1)

axis(1,at=c(-2,my.long.x),col="gray35",
     line=0.5,tick=T,labels=rep("",2),lwd=0.5,lwd.ticks=0,lty=3)

legend(legend.x[2]-50,
       legend.y[2]+2,c("State 0/1 events","State change point"),
       col = c(NA,"red"),
       y.intersp=0.85,x.intersp=-0.5, bty = "n",
       pch = c(NA,4), pt.cex = c(NA,2), cex=legend.cex,
       lty = c(NA,NA), lwd=c(NA,2))
points(legend.x[2]-46,legend.y[2] - 4.5, pch = 1, cex = 2, col = "blue")
points(legend.x[2]-44,legend.y[2] - 4.5, pch = 16, cex = 2, col = "blue")

## delta lambda
par(mar = c(0,0.3,0,0.1))
delta.lambda <- lambda.m2$lambda.t-lambda.true$lambda.t

plot(lambda.true$time.vec,delta.lambda,xlim=c(0,my.xlim), 
     ylim=c(-4,4), type="n",
     bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
delta.lambda <- lambda.m2$lambda.t-lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive<0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative>0] <- 0
polygon(c(0,lambda.true$time.vec,x_events), #200
        c(0,delta.negative,0),col=negative.col,border=NA)
polygon(c(0,lambda.true$time.vec,x_events), #200
        c(0,delta.positive,0),col=positive.col,border=NA)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,at=c(-2,my.long.x),col="gray35",
     line=0,tick=T,labels=rep("",2),lwd=0.7,lwd.ticks=0)

# model3
########
par(mar = c(0,0.3,1,0.1))
drawUniMMHPIntensityPaper(true.object,
                          simulation=fixStateTransition(test.mmhp),
                          yupper=y.ub[3],
                          color=add.alpha("black", alpha=line.alpha[1]),
                          line.width=line.wdth[1],
                          y.ratio=-0.05, min.y=-2, min.x=0, max.x=my.xlim, 
                          box.type="n")
drawUniMMHPIntensityPaper(object_hat,
                          simulation = list(
                            x=state.est.latent.mmhp.new$x.hat,                                          
                            z=state.est.latent.mmhp.new$z.hat,
                            tau=test.mmhp$tau,zt=test.mmhp$zt,
                            lambda.max=test.mmhp$lambda.max),
                          yupper=y.ub,add=TRUE,
                          color=add.alpha(model_colors[3],
                                          alpha=line.alpha[2]),
                          line.width=line.wdth[2])
legend(legend.x[3],legend.y[3],"C-MMHP",
       col = model_colors[3],
       y.intersp=0.6,x.intersp=0.1,bty = "n",
       lty =1, lwd = 5,cex=legend.cex,seg.len=1)
## draw box
axis(1,at=c(-2,my.long.x),col="gray35",
     line=0.5,tick=T,labels=rep("",2),lwd=0.5,lwd.ticks=0,lty=3)

legend(legend.x[3]-50,legend.y[3]+2, c("Overestimation","Underestimation"),
       col = c(positive.col,negative.col),
       y.intersp=0.75,x.intersp=0.5,bty = "n",
       lty = c(1,1), lwd = c(10,10), cex=legend.cex,seg.len=0.8)

## delta lambda
par(mar = c(0,0.3,0,0.1))
plot(0,0,xlim=c(0,my.xlim), ylim=c(-4,4), type="n",
     bty="n", xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
delta.lambda <- lambda.m3-lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive<0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative>0] <- 0
### hiding these for now
polygon(c(0,lambda.true$time.vec,x_events), # 200 
        c(0,delta.negative,0),col=negative.col,border=NA)
polygon(c(0,lambda.true$time.vec,x_events), # 200 
        c(0,delta.positive,0),col=positive.col,border=NA)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,at=c(-2,my.long.x),col="gray35",
     line=0,tick=T,labels=rep("",2),lwd=0.7,lwd.ticks=0)

## top
mtext(text="(b)",side=3,line=0.8,outer=TRUE,cex=2.5,font=2)
## bottom
mtext(text="Time",side=1,line=2,outer=TRUE,cex=2.25)
## left
mtext(text="Intensity",side=2,line=0.6,outer=TRUE,cex=2.25)

### just need to adjust the scales really!

```