### 2021-07-06 output_scripts/Paper_Plots.Rmd
just general tidying

---
title: "Plots for Latent Hawkes Paper"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.path = "paper_figure/",
  dev = "png", fig.height = 4.5, fig.width = 7,
  warning = FALSE, cache = TRUE
)

slides_plots <- FALSE
extra_plots <- FALSE

revision_plots <- TRUE
original_plots <- FALSE
### original plots for plots which are no longer needed, 
### revision plots are what required

current_cohort <- 5

# don't write to this path
data_path <- "../output/"

func_dir <- "../lib/"

library(here)
library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

cohort_names <- paste("cohort", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                      sep = "")
cohort_short_names <- paste("C", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                            sep = "")
cut_off <- 3
mice_number <- 12
```

```{r load rankings, echo=FALSE, include=FALSE, results='hide', cache=TRUE}
full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for (curr_cohort in fit_cohorts) {
  naive_rank_10[[curr_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})
```

```{r define_functions, eval=revision_plots, include=FALSE}
model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 +
    exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)
```

```{r creat colour df, eval=revision_plots, include=FALSE}
col_df <- tibble(
  method = c(
    "I&SI", "AggRank", "DSNL", "Glicko", "I-MMHP",
    "C-HP", "C-DCHP", "True", "C-MMHP"
  ),
  cols = viridis(9)
)
```

<!-- The aim of this notebook is a way to quickly and efficiently create all the plots -->
<!-- required for this paper. They will be built with the current fits and models. -->

# Figure 2: Contour Plot and C-HP Diagnostics

This is not effected by the final model.

## Figure 2a: Contour Plot

```{r fig-2a-contour-plot, eval=revision_plots}
current_cohort <- 5
# png(paste(plot_path,"contour.png",sep=""),width=480,height=480)
load(paste(data_path, cohort_names[current_cohort],
  "/cohort_hp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
eta1_mean <- mean(sim_cohort_hp$eta_1)
eta2_mean <- mean(sim_cohort_hp$eta_2)
eta3_mean <- mean(sim_cohort_hp$eta_3)
z.func <- function(x, y) {
  return(x * y * eta1_mean *
           exp(-eta2_mean * abs(x - y)) / (1 + exp(-eta3_mean * (y - x))))
  # this is because axes are flipped that we have y-x in last term
  # return(x*y*eta1_mean/(1+exp(eta2_mean*abs(x-y)-eta3_mean*(y-x))))
}
z_plot <- outer(
  seq(1, 0, -0.01),
  seq(0, 1, 0.01),
  z.func
)
my.image.plot(seq(0, 1, 0.01), seq(0, 1, 0.01), z_plot,
  xlab = expression(f[j]), ylab = expression(f[i]),
  my.main = "(a)", cex.main = 2.3,
  cex.lab = 2.5, xLabels = c(1, 0), yLabels = c(0, 1)
)
```

## Fig 2b: C-HP Diagnostics

<!-- This is using the correct form of the compensator for a Hawkes process. -->

```{r fig-2b-diag-plot,echo=FALSE,warning=FALSE, eval=revision_plots}
# current_cohort <- 5
current_cohort <- 5

load(paste(data_path, cohort_names[current_cohort],
  "/cohort_hp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
eta1_mean <- mean(sim_cohort_hp$eta_1)
eta2_mean <- mean(sim_cohort_hp$eta_2)
eta3_mean <- mean(sim_cohort_hp$eta_3)
f_est <- apply(sim_cohort_hp$f, 2, mean)
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort,
  raw_df =
    full_data[[cohort_names[current_cohort]]],
  clean_data
)
unique_pairs_df <- return_df %>%
  group_by(initiator, recipient) %>%
  dplyr::summarize(
    count = n(),
    observe = list(observe.id),
    observe.length = list(observe.time),
    no.events = list(no.events)
  )
unique_observe_win <- unique(return_df[, c("observe.id", "observe.time")])

ks.stat.matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
ks_pval_chp <- matrix(0, nrow = mice_number, ncol = mice_number)
n_matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
Lambda_matrix <- matrix(list(), nrow = mice_number, ncol = mice_number)
for (k in 1:mice_number) {
  for (l in c(1:mice_number)[-k]) {
    i <- expert_rank_10[[current_cohort]][k]
    j <- expert_rank_10[[current_cohort]][l]
    pair <- which(unique_pairs_df$initiator == i & 
                    unique_pairs_df$recipient == j)
    if (length(pair) > 0) {
      param_mean <- list(
        lambda0 = mean(sim_cohort_hp$lambda0),
        alpha = model1_fn$alpha.fun(
          f_est[i],
          f_est[j],
          eta1_mean,
          eta2_mean,
          eta3_mean
        ),
        beta = mean(sim_cohort_hp$beta)
      )
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_Lambda <- numeric(0)
      for (cur in c(1:length(current_window_vec))) { ## check length > 2
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator == i &
          return_df$recipient == j &
          return_df$observe.id == cur_win, "event.times"][[1]]
        current_obs_time <- return_df[return_df$initiator == i &
          return_df$recipient == j &
          return_df$observe.id == cur_win, "observe.time"]
        Lambda_vec <- uniHawkesCompensator(
          object = param_mean,
          events = c(
            current_event_time,
            current_obs_time
          )
        )
        all_Lambda <- c(all_Lambda, Lambda_vec)
        n_matrix[i, j] <- n_matrix[i, j] + length(current_event_time)
      }
      if (length(all_Lambda) > 0) {
        ks.stat.matrix[i, j] <- ks.test(all_Lambda, "pexp",
          rate = 1 / mean(all_Lambda)
        )$statistic
        ks_pval_chp[i, j] <- ks.test(all_Lambda, "pexp",
          rate = 1 / mean(all_Lambda)
        )$p.value
        Lambda_matrix[i, j][[1]] <- all_Lambda
      }
    }
  }
}
#
# png(paste(plot_path,
#           "/diagonsis_model1_",
#           cohort_names[current_cohort],
#           ".png",
#           sep = ''),
#     width=480,height=480)
# -------------- KS plot (stat)
col.vec <- cRamp(as.vector(ks.stat.matrix))
# need to do some sort of transformation between these here
vir_cols <- viridis::viridis(length((as.vector(ks.stat.matrix))))
# order(as.vector(ks.stat.matrix),method = "")

values <- as.vector(ks.stat.matrix)
sorted_vals <- sort(ks.stat.matrix)
# match(values,sorted_vals)
col.vec.vir <- vir_cols[match(values, sorted_vals)]

####
par(
  oma = c(6, 4, 4, 1), mar = rep(0, 4),
  mgp = c(0, .2, 0)
)
layout(cbind(matrix(
  data = c(1:(12 * 12)), nrow = 12, ncol = 12,
  byrow = TRUE
), rep(12 * 12 + 1, 12)),
widths = c(rep(1, 12), 2), heights = rep(1, 12)
)
for (i in expert_rank_10[[current_cohort]]) {
  for (j in expert_rank_10[[current_cohort]]) {
    if (ks.stat.matrix[i, j] == 0) {
      plot(0, 0,
        xlim = c(0, 5), ylim = c(0, 5), type = "n", xlab = "", ylab = "",
        axes = FALSE, frame = TRUE
      )
      rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4],
        border = 0, col = "white"
      )
    } else {
      Lambda.test <- Lambda_matrix[i, j][[1]]
      curve(pexp(x, 1 / mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
        ylab = "", xlab = "", axes = FALSE, frame = TRUE
      )
      rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4],
        border = 0, col = col.vec.vir[matrixToVecIdx(i, j, mice_number)]
      )
      curve(pexp(x, 1 / mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
        ylab = "", xlab = "", axes = FALSE, frame = TRUE, add = TRUE, lwd = 2
      )
      plot(ecdf(Lambda.test),
        add = TRUE, do.points = FALSE,
        verticals = TRUE, col = "red", lwd = 2
      )
    }
  }
}
par(mar = c(0, 2, 0, 3), mgp = c(0, 0.8, 0))
m1_ks <- ks.stat.matrix
col_obj <- list(
  x.vec = seq(0, max(as.vector(ks.stat.matrix)), length.out = 100),
  col.vec = viridis(100)
)

temp <- cLegend(as.vector(ks.stat.matrix))
# which(temp$x.vec == as.vector(ks.stat.matrix))
## need to adjust this for the legend also.
# image(1, cLegend(as.vector(ks.stat.matrix))$x.vec,
#       matrix(data=cLegend(as.vector(ks.stat.matrix))$x.vec,
#              ncol=length(cLegend(as.vector(ks.stat.matrix))$x.vec),nrow=1),
#       col=cLegend(as.vector(ks.stat.matrix))$col.vec,
#       xlab="",ylab="",axes=FALSE)
image(1, col_obj$x.vec,
  matrix(
    data = col_obj$x.vec,
    ncol = length(col_obj$x.vec), nrow = 1
  ),
  col = col_obj$col.vec,
  xlab = "", ylab = "", axes = FALSE
)
axis(4,
  at = c(0, 0.3, 0.6), labels = c(0, 0.3, 0.6), las = HORIZONTAL <- 1,
  cex.axis = 2.4
)

mtext(
  text = expert_rank_10[[current_cohort]],
  side = 1, line = 1.8, outer = TRUE, cex = 1.8,
  at = 1 / 14 * c(1:12) - 1 / 22
)
mtext(
  text = expert_rank_10[[current_cohort]],
  side = 2, line = 0.5, outer = TRUE, cex = 1.8,
  at = 1 / 12 * c(12:1) - 1 / 20, las = 2
)
mtext(
  text = "(b)", side = 3, line = 1.4, outer = TRUE,
  cex = 2.2,
  at = 0.45, font = 2
)
# dev.off()
ks_chp <- ks.stat.matrix

```

# Figure 3: Comparison of C-HP and C-DCHP

## Figure 3a: Baseline rates

```{r fig-3a-baseline-rate, eval=revision_plots}
current_cohort <- 5
# current_cohort <- 6 # 3 is maybe ok, 6/7 is interesting

load(paste(data_path, cohort_names[current_cohort],
  "/cohort_dchp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
gamma_mean <- apply(sim_cohort_dchp$gamma, 2, mean)
zeta_mean <- apply(sim_cohort_dchp$zeta, 2, mean)

# social matrix plot
latent_social_matrix <- gamma_mean %*%
  t(rep(1, mice_number)) + rep(1, mice_number) %*% t(zeta_mean)

# png(paste(plot_path,"/lambda0_",
# cohort_names[current_cohort],".png",sep=''),
#     width=400, height=300)
matrixPlotParameter(latent_social_matrix[
  rev(expert_rank_10[[current_cohort]]),
  expert_rank_10[[current_cohort]]
],
xLabels = expert_rank_10[[current_cohort]],
yLabels = rev(expert_rank_10[[current_cohort]])
)
# title = "(a)" )
# mtext(text=c(1,2,11,6,7,9), side=1, line=-0.35, outer=TRUE, cex=1.3,
#       at=1/17.8*seq(2,12,2)+0.09)
# dev.off()
```

## New Fig

KS Plot for C-DCHP

```{r ks-c-dchp, eval=FALSE, warning=FALSE}
ks.stat.matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
ks_pval_dchp <- matrix(0, ncol = mice_number, nrow = mice_number)
n_matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
Lambda_matrix <- matrix(list(), nrow = mice_number, ncol = mice_number)

gamma_mean <- apply(sim_cohort_dchp$gamma, 2, mean)
zeta_mean <- apply(sim_cohort_dchp$zeta, 2, mean)
eta1_mean <- mean(sim_cohort_dchp$eta_1)
eta2_mean <- mean(sim_cohort_dchp$eta_2)
eta3_mean <- mean(sim_cohort_dchp$eta_3)

for (k in 1:mice_number) {
  for (l in c(1:mice_number)[-k]) {
    i <- expert_rank_10[[current_cohort]][k]
    j <- expert_rank_10[[current_cohort]][l]
    pair <- which(unique_pairs_df$initiator == i & 
                    unique_pairs_df$recipient == j)
    if (length(pair) > 0) {
      param_mean <- list(
        lambda0 = gamma_mean[i] + zeta_mean[j],
        ## need to modify this
        # lambda0 <- mean(sim_cohort_dchp$gamma)
        alpha = model1_fn$alpha.fun(
          f_est[i],
          f_est[j],
          eta1_mean,
          eta2_mean,
          eta3_mean
        ),
        beta = mean(sim_cohort_dchp$beta)
      )
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_Lambda <- numeric(0)
      for (cur in c(1:length(current_window_vec))) { ## check length > 2
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator == i &
          return_df$recipient == j &
          return_df$observe.id == cur_win, "event.times"][[1]]
        current_obs_time <- return_df[return_df$initiator == i &
          return_df$recipient == j &
          return_df$observe.id == cur_win, "observe.time"]
        Lambda_vec <- uniHawkesCompensator(
          object = param_mean,
          events = c(
            current_event_time,
            current_obs_time
          )
        )
        all_Lambda <- c(all_Lambda, Lambda_vec)
        n_matrix[i, j] <- n_matrix[i, j] + length(current_event_time)
      }
      if (length(all_Lambda) > 0) {
        ks.stat.matrix[i, j] <- ks.test(all_Lambda, "pexp",
          rate = 1 / mean(all_Lambda)
        )$statistic
        ks_pval_dchp[i, j] <- ks.test(all_Lambda, "pexp",
          rate = 1 / mean(all_Lambda)
        )$p.value
        Lambda_matrix[i, j][[1]] <- all_Lambda
      }
    }
  }
}
col.vec <- cRamp(as.vector(ks.stat.matrix))
# need to do some sort of transformation between these here
vir_cols <- viridis::viridis(length((as.vector(ks.stat.matrix))))
# order(as.vector(ks.stat.matrix),method = "")

values <- as.vector(ks.stat.matrix)
sorted_vals <- sort(ks.stat.matrix)
# match(values,sorted_vals)
col.vec.vir <- vir_cols[match(values, sorted_vals)]

####
par(
  oma = c(6, 4, 4, 1), mar = rep(0, 4),
  mgp = c(0, .2, 0)
)
layout(cbind(matrix(
  data = c(1:(12 * 12)), nrow = 12, ncol = 12,
  byrow = TRUE
), rep(12 * 12 + 1, 12)),
widths = c(rep(1, 12), 2), heights = rep(1, 12)
)
for (i in expert_rank_10[[current_cohort]]) {
  for (j in expert_rank_10[[current_cohort]]) {
    if (ks.stat.matrix[i, j] == 0) {
      plot(0, 0,
        xlim = c(0, 5), ylim = c(0, 5), type = "n", xlab = "", ylab = "",
        axes = FALSE, frame = TRUE
      )
      rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4],
        border = 0, col = "white"
      )
    } else {
      Lambda.test <- Lambda_matrix[i, j][[1]]
      curve(pexp(x, 1 / mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
        ylab = "", xlab = "", axes = FALSE, frame = TRUE
      )
      rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4],
        border = 0, col = col.vec.vir[matrixToVecIdx(i, j, mice_number)]
      )
      curve(pexp(x, 1 / mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
        ylab = "", xlab = "", axes = FALSE, frame = TRUE, add = TRUE, lwd = 2
      )
      plot(ecdf(Lambda.test),
        add = TRUE, do.points = FALSE,
        verticals = TRUE, col = "red", lwd = 2
      )
    }
  }
}
par(mar = c(0, 2, 0, 3), mgp = c(0, 0.8, 0))
m1_ks <- ks.stat.matrix
col_obj <- list(
  x.vec = seq(0, max(as.vector(ks.stat.matrix)), length.out = 100),
  col.vec = viridis(100)
)

temp <- cLegend(as.vector(ks.stat.matrix))
image(1, col_obj$x.vec,
  matrix(
    data = col_obj$x.vec,
    ncol = length(col_obj$x.vec), nrow = 1
  ),
  col = col_obj$col.vec,
  xlab = "", ylab = "", axes = FALSE
)
axis(4,
  at = c(0, 0.3, 0.6), labels = c(0, 0.3, 0.6), las = HORIZONTAL <- 1,
  cex.axis = 2.4
)

mtext(
  text = expert_rank_10[[current_cohort]],
  side = 1, line = 1.8, outer = TRUE, cex = 1.8,
  at = 1 / 14 * c(1:12) - 1 / 22
)
mtext(
  text = expert_rank_10[[current_cohort]],
  side = 2, line = 0.5, outer = TRUE, cex = 1.8,
  at = 1 / 12 * c(12:1) - 1 / 20, las = 2
)
mtext(
  text = "C-DCHP", side = 3, line = 1.4, outer = TRUE,
  cex = 2.2,
  at = 0.45, font = 2
)
# dev.off()

ks_dchp <- ks.stat.matrix

```

Plot the negative log of the p values here.

```{r potential-fig3-diag, eval=FALSE}
M1_ks_p <- tibble(ks_p = as.vector(ks_chp), method = "C-HP")

M2_ks_p <- tibble(ks_p = as.vector(ks_dchp), 
                  method = "C-DCHP")

ks_p <- bind_rows(M1_ks_p, M2_ks_p) %>% 
  left_join(col_df, by = "method")

cols_plot <- col_df %>%
  filter(method %in% c("C-HP", "C-DCHP")) %>% 
  pull(cols)

exp_fcn <- function(x) exp(-x)

ks_p$method <- factor(ks_p$method, levels = c("C-HP", "C-DCHP"))

ks_p %>% 
  filter(ks_p != 0) %>% 
  mutate(log_p = -log(ks_p)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density.., fill = method), bins = 50) +
  scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value") +
  stat_function(fun = exp_fcn, colour = "red") +
  facet_wrap(~method) +
  labs(fill = "Method") +
  theme_minimal()

```

Could repeat this across all cohorts?

```{r above across all_cohorts, eval=FALSE}

ks_chp_array <- array(0, dim = c(10, mice_number, mice_number))
ks_dchp_array <- array(0, dim = c(10, mice_number, mice_number))

for(current_cohort in 1:10) {
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_hp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_dchp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  eta1_mean <- mean(sim_cohort_hp$eta_1)
  eta2_mean <- mean(sim_cohort_hp$eta_2)
  eta3_mean <- mean(sim_cohort_hp$eta_3)
  f_est <- apply(sim_cohort_hp$f, 2, mean)
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(current_cohort,
    raw_df =
      full_data[[cohort_names[current_cohort]]],
    clean_data
  )
  unique_pairs_df <- return_df %>%
    group_by(initiator, recipient) %>%
    dplyr::summarize(
      count = n(),
      observe = list(observe.id),
      observe.length = list(observe.time),
      no.events = list(no.events)
    )
  unique_observe_win <- unique(return_df[, c("observe.id", "observe.time")])

  ks.stat.matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
  ks_pval_chp <- matrix(0, nrow = mice_number, ncol = mice_number)
  n_matrix <- matrix(0, ncol = mice_number, nrow = mice_number)
  Lambda_matrix <- matrix(list(), nrow = mice_number, ncol = mice_number)
  for (k in 1:mice_number) {
    for (l in c(1:mice_number)[-k]) {
      i <- expert_rank_10[[current_cohort]][k]
      j <- expert_rank_10[[current_cohort]][l]
      pair <- which(unique_pairs_df$initiator == i & 
                      unique_pairs_df$recipient == j)
      if (length(pair) > 0) {
        param_mean <- list(
          lambda0 = mean(sim_cohort_hp$lambda0),
          alpha = model1_fn$alpha.fun(
            f_est[i],
            f_est[j],
            eta1_mean,
            eta2_mean,
            eta3_mean
          ),
          beta = mean(sim_cohort_hp$beta)
        )
        current_window_vec <- unique_pairs_df$observe[[pair]]
        all_Lambda <- numeric(0)
        for (cur in c(1:length(current_window_vec))) { ## check length > 2
          cur_win <- current_window_vec[cur]
          current_event_time <- return_df[return_df$initiator == i &
            return_df$recipient == j &
            return_df$observe.id == cur_win, "event.times"][[1]]
          current_obs_time <- return_df[return_df$initiator == i &
            return_df$recipient == j &
            return_df$observe.id == cur_win, "observe.time"]
          Lambda_vec <- uniHawkesCompensator(
            object = param_mean,
            events = c(
              current_event_time,
              current_obs_time
            )
          )
          all_Lambda <- c(all_Lambda, Lambda_vec)
          n_matrix[i, j] <- n_matrix[i, j] + length(current_event_time)
        }
        if (length(all_Lambda) > 0) {
          ks.stat.matrix[i, j] <- ks.test(all_Lambda, "pexp",
            rate = 1 / mean(all_Lambda)
          )$statistic
          ks_chp_array[current_cohort, i, j] <- ks.test(all_Lambda,
                                                        "pexp",
            rate = 1 / mean(all_Lambda)
          )$p.value
          Lambda_matrix[i, j][[1]] <- all_Lambda
        }
      }
    }
  }

  ### c-dchp
  gamma_mean <- apply(sim_cohort_dchp$gamma, 2, mean)
  zeta_mean <- apply(sim_cohort_dchp$zeta, 2, mean)
  eta1_mean <- mean(sim_cohort_dchp$eta_1)
  eta2_mean <- mean(sim_cohort_dchp$eta_2)
  eta3_mean <- mean(sim_cohort_dchp$eta_3)

  for (k in 1:mice_number) {
    for (l in c(1:mice_number)[-k]) {
      i <- expert_rank_10[[current_cohort]][k]
      j <- expert_rank_10[[current_cohort]][l]
      pair <- which(unique_pairs_df$initiator == i & 
                      unique_pairs_df$recipient == j)
      if (length(pair) > 0) {
        param_mean <- list(
          lambda0 = gamma_mean[i] + zeta_mean[j],
          ## need to modify this
          # lambda0 <- mean(sim_cohort_dchp$gamma)
          alpha = model1_fn$alpha.fun(
            f_est[i],
            f_est[j],
            eta1_mean,
            eta2_mean,
            eta3_mean
          ),
          beta = mean(sim_cohort_dchp$beta)
        )
        current_window_vec <- unique_pairs_df$observe[[pair]]
        all_Lambda <- numeric(0)
        for (cur in c(1:length(current_window_vec))) { ## check length > 2
          cur_win <- current_window_vec[cur]
          current_event_time <- return_df[return_df$initiator == i &
            return_df$recipient == j &
            return_df$observe.id == cur_win, "event.times"][[1]]
          current_obs_time <- return_df[return_df$initiator == i &
            return_df$recipient == j &
            return_df$observe.id == cur_win, "observe.time"]
          Lambda_vec <- uniHawkesCompensator(
            object = param_mean,
            events = c(
              current_event_time,
              current_obs_time
            )
          )
          all_Lambda <- c(all_Lambda, Lambda_vec)
          n_matrix[i, j] <- n_matrix[i, j] + length(current_event_time)
        }
        if (length(all_Lambda) > 0) {
          ks.stat.matrix[i, j] <- ks.test(all_Lambda, "pexp",
            rate = 1 / mean(all_Lambda)
          )$statistic
          ks_dchp_array[current_cohort, i, j] <- ks.test(all_Lambda, "pexp",
            rate = 1 / mean(all_Lambda)
          )$p.value
          Lambda_matrix[i, j][[1]] <- all_Lambda
        }
      }
    }
  }
}

```

```{r plot-all-cohorts, eval=FALSE}
M1_ks_p <- tibble(ks_p = as.vector(ks_chp_array), method = "C-HP")

M2_ks_p <- tibble(ks_p = as.vector(ks_dchp_array), 
                  method = "C-DCHP")

ks_p <- bind_rows(M1_ks_p, M2_ks_p) %>% 
  left_join(col_df, by = "method")

cols_plot <- col_df %>%
  filter(method %in% c("C-HP", "C-DCHP")) %>% 
  pull(cols)

exp_fcn <- function(x) exp(-x)

ks_p$method <- factor(ks_p$method, levels = c("C-HP", "C-DCHP"))

ks_p %>% 
  filter(ks_p != 0) %>% 
  mutate(log_p = -log(ks_p)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density.., fill = method), bins = 50) +
  scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value") +
  stat_function(fun = exp_fcn, colour = "red") +
  facet_wrap(~method) +
  labs(fill = "Method") +
  theme_minimal()

```

## Figure 3b: Ranking Comparison

```{r fig-3b-rank-comparison, include=FALSE,eval=original_plots}
current_cohort <- 5
colors_plot <- col_df %>%
  filter(method %in% c("C-HP", "C-DCHP")) %>%
  pull(cols)
load(paste(data_path,
  cohort_names[current_cohort],
  "/cohort_hp_stan_result_",
  cohort_names[current_cohort],
  ".RData",
  sep = ""
))
load(paste(data_path,
  cohort_names[current_cohort],
  "/cohort_dchp_stan_result_",
  cohort_names[current_cohort],
  ".RData",
  sep = ""
))
# png(paste(plot_path,"/rank_",cohort_names[current_cohort],".png",sep=''),
#     width=400, height=300)
par(oma = c(0, 0, 2, 0), mar = c(3, 4, 1, 1), mgp = c(0, 0.5, 0))
plot(order(expert_rank_10[[current_cohort]]),
  apply(sim_cohort_hp$f, 2, mean),
  pch = 16, cex = 1.2,
  # col= "steelblue",
  col = colors_plot[1],
  ylab = "", xlab = "",
  xlim = c(1, 12.5), ylim = c(0, 1),
  xaxt = "n", yaxt = "n"
)
segments(order(expert_rank_10[[current_cohort]]),
  apply(sim_cohort_hp$f, 2, mean) -
    apply(sim_cohort_hp$f, 2, sd),
  order(expert_rank_10[[current_cohort]]),
  apply(sim_cohort_hp$f, 2, mean) +
    apply(sim_cohort_hp$f, 2, sd),
  # col= add.alpha("steelblue",0.8),lwd=2)
  col = add.alpha(colors_plot[1], 0.8), lwd = 2
)
points(order(expert_rank_10[[current_cohort]]) + 0.2,
  apply(sim_cohort_dchp$f, 2, mean),
  pch = 16,
  cex = 1.2,
  col = colors_plot[2]
)
# col="goldenrod2")
segments(order(expert_rank_10[[current_cohort]]) + 0.2,
  apply(sim_cohort_dchp$f, 2, mean) - apply(sim_cohort_dchp$f, 2, sd),
  order(expert_rank_10[[current_cohort]]) + 0.2,
  apply(sim_cohort_dchp$f, 2, mean) + apply(sim_cohort_dchp$f, 2, sd),
  col = add.alpha(colors_plot[2], 0.8), lwd = 2
)
# col= add.alpha("goldenrod2",0.8),lwd=2)

axis(BELOW <- 1, at = 1:12, 
     labels = expert_rank_10[[current_cohort]],
     cex.axis = 1.3,
     tck = -0.015)
# mtext(text=11, side=1, line=0.5, cex=1.3, at=6)

axis(LEFT <- 2,
  at = seq(0, 1, 0.2), labels = seq(0, 1, 0.2), las = HORIZONTAL <- 1,
  cex.axis = 1.3, tck = -0.015
)
mtext(side = 1, text = "I&SI rank", line = 1.8, cex = 1.5)
mtext(side = 2, text = "Latent rank", line = 2.4, cex = 1.6)
mtext(side = 3, text = "(b)", line = 0.8, cex = 1.6, font = 2)
legend("bottomleft",
  legend = c("C-HP", "C-DCHP"),
  pch = c(16, 16), lty = c(1, 1), lwd = c(2, 2),
  # col=c("steelblue","goldenrod"),
  col = c(colors_plot[1], colors_plot[2]),
  cex = 1.5
)
# dev.off()
```

```{r, eval=FALSE, include=FALSE}
## need Hmisc for this
library(Hmisc)

par(mar = c(3, 3, 1, 1))
plot(as.vector(n_matrix[n_matrix > 0]),
     as.vector(ks.stat.matrix[n_matrix > 0]),
     xlab = "",
     ylab = "",
     pch = 16,
     col = "gray",
     cex = 0.8,
     xaxt = "n",
     yaxt = "n",
  ylim = c(0, 0.65)
)
mtext("Number of events", cex = 1.5, side = 1, line = 1.8)
mtext("K-S statistics", cex = 1.5, side = 2, line = 1.5)
mgp.axis(1, mgp = c(0, 0.3, 0),
         at = seq(0, 120, 10),
         cex.axis = 1,
         lwd = 0,
         lwd.ticks = 0.1)
mgp.axis(2, mgp = c(0, 0.1, 0),
         at = seq(-0.2, 0.6, 0.2),
         cex.axis = 1,
         lwd = 0,
         lwd.ticks = 0.8)
```

# Figure 4:

## Figure 4a Recover Latent Ranks
```{r fig-4a, fig.height=3.5, fig.width=8, eval=original_plots}
sim_data_path <- "../output/sims_m3_update/"

# load(paste(sim_data_path,"sim_model3_fit123.RData",sep=''))
# load(paste(sim_data_path,"fit123_state_est.RData",sep=''))
#### need to edit here to account for separate files now

m3_files <- list.files(sim_data_path, pattern = "sim_model3_fit*")
# m1_files

# load(paste0(m1_path,m1_files[1]))
sim_model3_stan_sim1_total <- list()
sim_model3_stan_sim2_total <- list()
sim_model3_stan_sim3_total <- list()

for (i in seq_along(m3_files)) {
  load(paste0(sim_data_path, m3_files[i]))
  sim_model3_stan_sim1_total[[i]] <- sim_model3_stan_sim1
  sim_model3_stan_sim2_total[[i]] <- sim_model3_stan_sim2
  sim_model3_stan_sim3_total[[i]] <- sim_model3_stan_sim3
}

n_sim <- 50

# png(paste(plot_path,"sim_latent_recover.png",sep=""), height=500, width=900)
par(
  mfrow = c(3, length(object_par$f_vec_1)),
  mar = c(0, 0, 1, 0),
  oma = c(6.5, 11, 3.5, 1)
)

y.ub <- c(12, 12, 12, 12, 12)
text.y <- c(rep(8, 4), 8)
text.x <- c(0.1, 0.2, 0.4, 0.7, 0.9)
text.x.adj <- c(0.3, 0.4, 0.6, 0.5, 0.8)

model_colors <- col_df %>%
  filter(method %in% c(
    "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

for (m in c(1:3)) {
  for (j in c(1:length(object_par$f_vec_1))) {
    plot(1, 4,
      xlim = c(0, 1), ylim = c(0, y.ub[j]),
      type = "n", xlab = "",
      ylab = "", xaxt = "n",
      yaxt = "n", axes = FALSE
    )

    ## Posterior density
    # current_result <- eval(parse(text=paste("sim_model3_stan_sim",m,sep="")))
    current_result <- eval(parse(text = paste("sim_model3_stan_sim",
      m,
      "_total",
      sep = ""
    )))
    for (i in c(1:n_sim)) {
      lines(density(current_result[[i]]$f[, j]),
        col = add.alpha(model_colors[m], alpha = 0.8), cex = 0.6, lwd = 0.8
      )
    }

    ## plot true value
    abline(v = object_par$f_vec_1[j], lwd = 1.8, lty = 2, col = "gray30")

    ## plot box
    box.wdt <- 1.5
    box.col <- "black"
    box(lty = "solid", col = box.col, lwd = 0.7)

    if (m == 3) {
      text(text.x.adj[j], text.y[j], text.x[j], cex = 2.4)
    }
  }
  ## bottom
  mtext(
    text = expression(f[1], f[2], f[3], f[4], f[5]),
    side = 1, line = 2.4, outer = TRUE,
    at = c(1:length(object_par$f_vec_1)) / length(object_par$f_vec_1) - 0.1,
    cex = 2
  )
  mtext(
    text = "Latent rank variables", side = 1,
    line = 5.3,
    outer = TRUE,
    cex = 2.25,
    at = 0.5
  )
  ## top
  mtext(
    text = "(a)", side = 3,
    line = 0,
    outer = TRUE,
    cex = 2.5,
    at = 0.5, font = 2
  )
  ## left
  mtext(
    text = rev(c("C-HP  ", "C-DCHP", "C-MMHP")),
    side = 2,
    line = 0.5,
    outer = TRUE,
    at = c(0:2) / 3 + 0.15,
    cex = 1.75, las = 2
  )
}
```

```{r fig4a-new-hist, fig.height=3.5, fig.width=8, eval=original_plots}
model_colors <- col_df %>% filter(method %in% c(
  "C-HP",
  "C-DCHP",
  "C-MMHP"
))

model_colors <- model_colors %>% pull(cols)

m1 <- sim_model3_stan_sim1_total %>%
  map(~ apply(.x$f, 2, mean)) %>%
  enframe() %>%
  unnest(cols = value) %>%
  mutate(
    node_id = rep(c("f1", "f2", "f3", "f4", "f5"), 50),
    method = "C-HP"
  )

m2 <- sim_model3_stan_sim2_total %>%
  map(~ apply(.x$f, 2, mean)) %>%
  enframe() %>%
  unnest(cols = value) %>%
  mutate(
    node_id = rep(c("f1", "f2", "f3", "f4", "f5"), 50),
    method = "C-DCHP"
  )

m3 <- sim_model3_stan_sim3_total %>%
  map(~ apply(.x$f, 2, mean)) %>%
  enframe() %>%
  unnest(cols = value) %>%
  mutate(
    node_id = rep(c("f1", "f2", "f3", "f4", "f5"), 50),
    method = "C-MMHP"
  )

rank_means <- bind_rows(m1, m2, m3)

# rank_means

true_values <- tibble(
  id = c("f1", "f2", "f3", "f4", "f5"),
  values = c(0.1, 0.2, 0.4, 0.7, 0.9),
  val_names = c("   0.1", "0.2", "0.4", "0.7", "0.9")
)

rank_means %>%
  left_join(true_values, by = c("node_id" = "id")) %>%
  ggplot(aes(value, fill = method)) +
  geom_histogram() +
  geom_vline(aes(xintercept = values),
    alpha = 0.6,
    linetype = "dashed"
  ) +
  ### add text here
  geom_text(aes(values, 18, label = val_names),
    size = 3, hjust = 1
  ) +
  scale_fill_manual(values = model_colors) +
  facet_grid(cols = vars(node_id), rows = vars(method)) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "", y = "")
```

## Modified Figure 4a

Reformatting this to boxplots for each of the 3 models, with the true
ranking values included for each.

Will need to update this with full final simulations but fully set up now.

```{r new-fig-4a, eval = revision_plots}
sim_data_path <- here("output", "revisions", "sim_m3", "sim_3", "/")

num_nodes <- 10

f_vec <- seq(from = 0.2,
               to = 0.9,
               length.out = num_nodes)

sim_ranks_mean <- tibble()
sim_ranks <- tibble()

model_colors <- col_df %>%
  filter(method %in% c(
    "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

sims <- 1:20
sims <- sims[c(-9, -13)]

for(curr_sim in sims) {
  load(paste0(sim_data_path,
              "sim_model3_fit123_",
              curr_sim,
              ".RData"))

  f_m1 <- sim_model3_stan_sim1 %>% select(starts_with("f"))
  m1_ranks <- apply(f_m1, 2, mean)

  f_m1_sims <- f_m1 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")
  f_m1_sims$model <- "C-HP"

  f_m2 <- sim_model3_stan_sim2 %>% select(starts_with("f"))
  m2_ranks <- apply(f_m2, 2, mean)

  f_m2_sims <- f_m2 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m2_sims$model <- "C-DCHP"

  f_m3 <- sim_model3_stan_sim3 %>% select(starts_with("f"))
  m3_ranks <- apply(f_m3, 2, mean)

  f_m3_sims <- f_m3 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m3_sims$model <- "C-MMHP"

  curr_mean_ranks <- tibble(model = c(rep("C-HP", num_nodes),
                                 rep("C-DCHP", num_nodes),
                                 rep("C-MMHP", num_nodes)),
                       ranks = c(m1_ranks,
                                 m2_ranks,
                                 m3_ranks),
                       node = rep(1:num_nodes, 3),
                       true_rank = rep(f_vec, 3),
                       sim_id = curr_sim)
  sim_ranks_mean <- sim_ranks_mean %>% 
    bind_rows(curr_mean_ranks)

  sim_ranks <- sim_ranks %>% 
    bind_rows(f_m1_sims, f_m2_sims, f_m3_sims)

}

sim_ranks$node <- factor(sim_ranks$node, levels = 1:num_nodes)
sim_ranks_mean$node <- factor(sim_ranks_mean$node, levels = 1:num_nodes)

# sim_ranks$node

sim_ranks_mean$model <- factor(sim_ranks_mean$model, 
                               levels = c( "C-HP", "C-DCHP", "C-MMHP"))

mean_sims <- sim_ranks_mean %>% 
  left_join(col_df, by = c("model" = "method"))

mean_sims$model <- factor(mean_sims$model, levels = c("C-HP", "C-DCHP",
                                                      "C-MMHP"))

true_ranks <- tibble(node = rep(1:num_nodes,3),
                     true_rank = rep(f_vec,3),
                     model = c(rep("C-HP", 10),rep("C-DCHP", 10),
                               rep("C-MMHP", 10)))
true_ranks$node <- factor(true_ranks$node, levels = 1:num_nodes)

true_ranks$model <- factor(true_ranks$model, 
                           levels = c("C-HP", "C-DCHP","C-MMHP"))

mean_sims %>% 
  # filter(model == "C-MMHP") %>% 
  ggplot(aes(node, ranks, fill = model)) + 
  geom_boxplot(show.legend = FALSE) +
  scale_fill_manual(values = model_colors, guide = FALSE) +
  # geom_point(aes(node, true_rank, colour = "red")) +
  ylim(c(0,1)) +
  labs(title = "(a)",
       subtitle = "Mean rank from each simulation", 
       y = "Posterior Mean Rank", x = "Node") +
  facet_wrap(~model) +
  geom_point(data = true_ranks,
             aes(node, true_rank, colour = "True Value")) +
  scale_shape_discrete(name = "True Value") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, size = 26),
        text = element_text(size = 18),
        legend.text = element_text(size = 18))

## this is just the mean from each simulation, not the posterior draws

# sim_ranks %>%
#   ggplot(aes(node, rank)) +
#   geom_boxplot() +
#   facet_wrap(~model) +
#   geom_point(data = true_ranks, aes(node, true_rank, colour = "red"),
#              show.legend = FALSE) +
#   labs(title = "All Posterior Draws")
```

## Figure 4b: Recover Latent States

Need to modify this after finally correctly formulating the 
simulations.

```{r prepare data for 4b,cache=TRUE, eval=FALSE}
plot.s <- 7 # 7 is pretty good
plot.i <- 4
plot.j <- 5

cur_i <- 4
cur_j <- 5

sim_data_path <- here("output", "revisions", "sim_m3", "june_18", "/")

load(paste0(
  sim_data_path, "/", "sim_model3_",
  plot.s,
  ".RData"
))
test.mmhp <- sim_model3_data$mmhp_matrix[plot.i, plot.j][[1]]
temp.t <- test.mmhp$tau
current.n <- length(temp.t) - 1
time.segment <- seq(0, tail(temp.t, 1), length.out = 5000)

### Preprocess the model inference result
## model 1
#########

## get f from cmdstan

f_draws <- sim_model3_stan_sim1 %>% 
  select(starts_with("f"))

model1_par_est <- list(
  lambda0 = mean(sim_model3_stan_sim1$lambda0),
  eta_1 = mean(sim_model3_stan_sim1$eta_1),
  eta_2 = mean(sim_model3_stan_sim1$eta_2),
  eta_3 = mean(sim_model3_stan_sim1$eta_3),
  beta = mean(sim_model3_stan_sim1$beta),
  f = apply(f_draws, 2, mean)
)
lambda.m1 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 =
      model1_par_est$lambda0,
    alpha =
      model1_fn$alpha.fun(
        model1_par_est$f[plot.i],
        model1_par_est$f[plot.j],
        model1_par_est$eta_1,
        model1_par_est$eta_2,
        model1_par_est$eta_3
      ),
    beta = model1_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## model2
#############
gamma_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("gamma"))

zeta_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("zeta"))

f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

model2_par_est <- list(
  gamma = apply(gamma_draws, 2, mean),
  zeta = apply(zeta_draws, 2, mean),
  eta_1 = mean(sim_model3_stan_sim2$eta_1),
  eta_2 = mean(sim_model3_stan_sim2$eta_2),
  eta_3 = mean(sim_model3_stan_sim2$eta_3),
  beta = mean(sim_model3_stan_sim2$beta),
  f = apply(f_draws, 2, mean)
)

lambda.m2 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 = model2_par_est$gamma[plot.i] +
      model2_par_est$zeta[plot.j],
    alpha = model1_fn$alpha.fun(
      model2_par_est$f[plot.i],
      model2_par_est$f[plot.j],
      model2_par_est$eta_1,
      model2_par_est$eta_2,
      model2_par_est$eta_3
    ),
    beta = model2_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## Model3
###

lam0_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("gamma"))

lam1_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("zeta"))

gamma_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("gamma"))

zeta_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("zeta"))

f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

lam0_vec <- apply(lam0_draws, 2, mean)
lam1_vec <- apply(lam1_draws, 2, mean)
lam0_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
lam1_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)

clean_sim_data <- cleanSimulationData(
  raw_data = sim_model3_data,
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)

### TO DO
### need to check this here for pairs with less than cut off events

for (i in seq_along(lam0_vec)) {
  row_id <- clean_sim_data$I_fit[i]
  col_id <- clean_sim_data$J_fit[i]
  lam0_par_est[row_id, col_id] <- lam0_vec[i]
  lam1_par_est[row_id, col_id] <- lam1_vec[i]
}

###

#############
mmhp_par_est <- list(
  lambda0 = lam0_par_est,
  lambda1 = lam1_par_est,
  eta_1 = mean(sim_model3_stan_sim3$eta_1),
  eta_2 = mean(sim_model3_stan_sim3$eta_2),
  eta_3 = mean(sim_model3_stan_sim3$eta_3),
  beta = mean(sim_model3_stan_sim3$beta),
  f = apply(f_draws, 2, mean)
)

object_hat <- list(
  lambda0 = mmhp_par_est$lambda0[cur_i, cur_j],
  lambda1 = mmhp_par_est$lambda1[cur_i, cur_j],
  alpha = model3_fn$alpha.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_1,
    mmhp_par_est$eta_2
  ),
  beta = mmhp_par_est$beta,
  q1 = model3_fn$q1.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  ),
  q2 = model3_fn$q0.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  )
)

## then need to load in the state estimate here
# load(paste0(here("output", "revisions", "sim_m3"),
#   "/", "fit123_state_est_",
#   plot.s,
#   ".RData"
# ))

load(paste0(
  sim_data_path, "/", "fit123_state_est_",
  plot.s,
  ".RData"
))

state.est.latent.mmhp <- interpolate_state_est_lst[plot.i, plot.j][[1]]
state.est.latent.mmhp.new <- fixInterpolateState(state.est.latent.mmhp,
  termination = 50 
  ### TO DO update for longer sims
)

step.fun.est <- stepfun(
  state.est.latent.mmhp$x.hat,
  2 - state.est.latent.mmhp$z.hat
)

lambda.m3 <- mmhpIntensityNumeric(
  params = object_hat,
  t = temp.t[-1],
  time.vec = time.segment,
  latent.vec = step.fun.est(time.segment)
)

object_true <- lapply(object_matrix, function(x) x[plot.i, plot.j])
names(object_true) <- c("lambda0", "lambda1", "alpha", "beta", "q1", "q2")

lambda.true <- mmhpTrueIntensityNumeric(
  params = object_true,
  t = temp.t,
  latent = list(
    x = fixStateTransition(test.mmhp)$x,
    z = fixStateTransition(test.mmhp)$z
  ),
  time.vec = time.segment
)

```

```{r fig-4b-recover-latent-states,fig.width=8,fig.height=4, eval=FALSE}
mmhp_lty <- 1

y.ub <- c(8, 8, 8) # c(130,72,31)
x_events <- 190 # 135#150
my.xlim <- 200 # 160#200#130
my.long.x <- 205 # 205#134.5
legend.x <- c(100, 100, 100) # c(135,135,135)
legend.y <- # c(34,32,32)
legend.y <- c(6, 6, 6) # c(15,12,15)#c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 # 3
line.alpha <- c(1, 0.5)
line.wdth <- c(1.5, 2)
negative.col <- add.alpha("lightskyblue")
positive.col <- add.alpha("tomato")

model_colors <- col_df %>% filter(method %in% c(
  "C-HP",
  "C-DCHP",
  "C-MMHP"
))

model_colors <- model_colors %>% pull(cols)

# model1
############
true.object <- object_true

```

```{r actual-plot, eval=FALSE}
layout(matrix(c(1:6), 6, 1), heights = rep(c(7.5, 2.5), 3))
par(
  mar = c(0, 0.3, 0, 0.1), oma = c(0, 3.5, 4, 0),
  tcl = 0.2, mgp = c(0.5, 0, 0), xpd = TRUE
)

mmhp_est <- fixStateTransition(test.mmhp)

drawUniMMHPIntensityPaper(true.object,
                          simulation = mmhp_est,
                          yupper = y.ub[1],
                          color = add.alpha("black", alpha = line.alpha[1]),
                          line.width = line.wdth[1],
                          y.ratio = -0.05, min.y = -0.5,
                          min.x = 0, max.x = my.xlim,
                          title_name = paste(plot.i, "->", plot.j),
                          title.cex = 2,
                          box.type = "n",
                          line_type = mmhp_lty
)
drawHawkesIntensityPaper(
  lambda0 = model1_par_est$lambda0,
  alpha = model1_fn$alpha.fun(
    model1_par_est$f[plot.i],
    model1_par_est$f[plot.j],
    model1_par_est$eta_1,
    model1_par_est$eta_2,
    model1_par_est$eta_3
  ),
  beta = model1_par_est$beta,
  events = test.mmhp$tau,
  color = add.alpha(model_colors[1],
                    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)

# draw box
axis(1,
     at = c(-2, my.long.x),
     col = "gray35",
     line = 0.5,
     tick = T,
     labels = rep("", 2),
     lwd = 0.5,
     lwd.ticks = 0,
     lty = 3
)

## add legend
# changed 34 to 23, 25 to 18, -20 to -10
# legend(10,legend.y[1],c("State 1/0 events","State change point"),
#        col = c(NA,"red"),
#        y.intersp=0.85,x.intersp=-0.1, bty = "n",
#        pch = c(NA,4), pt.cex = c(NA,2), cex=legend.cex,
#        lty = c(NA,NA), lwd=c(NA,4))
# points(29,84,pch=1,cex=2,col="blue")
# points(26,84,pch=16,cex=2,col="blue")

#### more legend
legend(legend.x[1] - 2, legend.y[1] + 1, c("True"),
       col = c("black"),
       y.intersp = 0.88, x.intersp = 0.65, bty = "n",
       lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)
legend(legend.x[1] + 22, legend.y[1] + 1, c("C-HP"),
       col = c(model_colors[1]),
       y.intersp = 0.88, x.intersp = 0.65, bty = "n",
       lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
     xlim = c(0, my.xlim), 
     ylim = c(-0.5, 4),
     type = "n",
     bty = "n",
     xlab = "",
     ylab = "",
     xaxt = "n",
     yaxt = "n",
     axes = FALSE
)
delta.lambda <- lambda.m1$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now ###
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.negative, 0),
        col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.positive, 0),
        col = positive.col, border = NA
)

axis(1,
     at = c(-1, my.long.x), 
     col = "gray35",
     line = 0, tick = T, 
     labels = rep("", 2),
     lwd = 0.7,
     lwd.ticks = 0
)

# model2
############
drawUniMMHPIntensityPaper(true.object,
                          simulation = mmhp_est,
                          ### these points aren't consistent?
                          yupper = y.ub[2],
                          color = add.alpha("black",
                                            alpha = line.alpha[1]),
                          line.width = line.wdth[1],
                          y.ratio = -0.05,
                          min.y = -0.5,
                          min.x = 0,
                          max.x = my.xlim,
                          box.type = "n",
                          line_type = mmhp_lty
)
drawHawkesIntensityPaper(
  lambda0 = model2_par_est$gamma[plot.i] +
    model2_par_est$zeta[plot.j],
  alpha = model1_fn$alpha.fun(
    model2_par_est$f[plot.i],
    model2_par_est$f[plot.j],
    model2_par_est$eta_1,
    model2_par_est$eta_2,
    model2_par_est$eta_3
  ),
  beta = model2_par_est$beta,
  events = test.mmhp$tau,
  color = add.alpha(model_colors[2],
                    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)

### used legend
legend(legend.x[2] - 2, legend.y[2] + 1, c("True"),
       col = c("black"),
       y.intersp = 0.88, x.intersp = 0.65, bty = "n",
       lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)
legend(legend.x[2] + 22, legend.y[2] + 1, "C-DCHP",
       col = model_colors[2],
       y.intersp = 0.88,
       x.intersp = 0.65,
       bty = "n",
       lty = 1,
       lwd = 5,
       cex = legend.cex,
       seg.len = 1
)

axis(1,
     at = c(-2, my.long.x),
     col = "gray35",
     line = 0.5,
     tick = T,
     labels = rep("", 2),
     lwd = 0.5,
     lwd.ticks = 0,
     lty = 3
)

## location of legend
legend(legend.x[2] - 95,
       legend.y[2] + 2, c("State 0/1 events", "State change point"),
       col = c(NA, "red"),
       y.intersp = 0.85,
       x.intersp = -0,
       bty = "n",
       pch = c(NA, 4), pt.cex = c(NA, 2), cex = legend.cex,
       lty = c(NA, NA), lwd = c(NA, 2)
)
points(legend.x[2] - 89, legend.y[2], 
       pch = 1, cex = 2, col = "blue")
points(legend.x[2] - 85, legend.y[2],
       pch = 16, cex = 2, col = "blue")

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t

plot(lambda.true$time.vec, delta.lambda,
     xlim = c(0, my.xlim),
     ylim = c(-0.5, 4), type = "n",
     bty = "n",
     xlab = "",
     ylab = "",
     xaxt = "n",
     yaxt = "n",
     axes = FALSE
)
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.negative, 0),
        col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.positive, 0),
        col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
     at = c(-2, my.long.x), col = "gray35",
     line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

# model3
########
par(mar = c(0, 0.3, 1, 0.1))
drawUniMMHPIntensityPaper(true.object,
                          simulation = mmhp_est,
                          yupper = y.ub[3],
                          color = add.alpha("black", alpha = line.alpha[1]),
                          line.width = line.wdth[1],
                          y.ratio = -0.05, min.y = -0.5,
                          min.x = 0, max.x = my.xlim,
                          box.type = "n", line_type = mmhp_lty
)
drawUniMMHPIntensityPaper(object_hat,
                          simulation = list(
                            x = state.est.latent.mmhp.new$x.hat,
                            z = state.est.latent.mmhp.new$z.hat,
                            tau = test.mmhp$tau, zt = test.mmhp$zt,
                            lambda.max = test.mmhp$lambda.max
                          ),
                          yupper = y.ub, add = TRUE,
                          color = add.alpha(model_colors[3],
                                            alpha = line.alpha[2]
                          ),
                          line.width = line.wdth[2]
)
legend(legend.x[3] - 2, legend.y[3] + 1, c("True"),
       col = c("black"),
       y.intersp = 0.88, x.intersp = 0.65, bty = "n",
       lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)

legend(legend.x[3] + 22, legend.y[3] + 1, "C-MMHP",
       col = model_colors[3],
       y.intersp = 0.88,
       x.intersp = .65,
       bty = "n",
       lty = 1,
       lwd = 5,
       cex = legend.cex,
       seg.len = 1
)
## draw box
axis(1,
     at = c(-2, my.long.x),
     col = "gray35",
     line = 0.5,
     tick = T,
     labels = rep("", 2), 
     lwd = 0.5,
     lwd.ticks = 0, lty = 3
)

legend(legend.x[3] - 95, legend.y[3] + 2,
       c("Overestimation", "Underestimation"),
       col = c(positive.col, negative.col),
       y.intersp = 0.75, x.intersp = 0.5, bty = "n",
       lty = c(1, 1), lwd = c(10, 10), cex = legend.cex, seg.len = 0.8
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
     xlim = c(0, my.xlim), 
     ylim = c(-0.5, 4), type = "n",
     bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
     axes = FALSE
)
delta.lambda <- lambda.m3 - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.negative, 0),
        col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
        c(0, delta.positive, 0),
        col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
     at = c(-2, my.long.x), col = "gray35",
     line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

## top
mtext(text = "(b)", side = 3, line = 0.8,
      outer = TRUE, cex = 2.5, font = 2)
## bottom
mtext(text = "Time", side = 1, line = 2,
      outer = TRUE, cex = 2.25)
## left
mtext(text = "Intensity", side = 2,
      line = 0.6, outer = TRUE, cex = 2.25)

```

## Fig 4 repeated for data generated from other models

For completeness, we also wish to repeat the
above two plots using data generated 
from M1 and M2.

For data simulated from M1.

```{r est ranks for m1 sim data, include=extra_plots, eval=extra_plots}
m1_path <- "../output/sims/M1/"

m1_files <- list.files(m1_path, pattern = "sim_model1_fit*")
# m1_files

# load(paste0(m1_path,m1_files[1]))
sim_model1_stan_sim1_total <- list()
sim_model1_stan_sim2_total <- list()
sim_model1_stan_sim3_total <- list()

for (i in seq_along(m1_files)) {
  load(paste0(m1_path, m1_files[i]))
  sim_model1_stan_sim1_total[[i]] <- sim_model1_stan_sim1
  sim_model1_stan_sim2_total[[i]] <- sim_model1_stan_sim2
  sim_model1_stan_sim3_total[[i]] <- sim_model1_stan_sim3
}

m2_path <- "../output/sims/M2/"

m2_files <- list.files(m2_path, pattern = "sim_model2_fit*")
# m1_files

# load(paste0(m1_path,m1_files[1]))
sim_model2_stan_sim1_total <- list()
sim_model2_stan_sim2_total <- list()
sim_model2_stan_sim3_total <- list()

for (i in seq_along(m2_files)) {
  load(paste0(m2_path, m2_files[i]))
  sim_model2_stan_sim1_total[[i]] <- sim_model2_stan_sim1
  sim_model2_stan_sim2_total[[i]] <- sim_model2_stan_sim2
  sim_model2_stan_sim3_total[[i]] <- sim_model2_stan_sim3
}
```

```{r plot est ranks M1 and M2, include=extra_plots, eval=extra_plots}
n_sim <- 50

# png(paste(plot_path,"sim_latent_recover.png",sep=""), height=500, width=900)
par(
  mfrow = c(3, length(object_par$f_vec_1)),
  mar = c(0, 0, 1, 0),
  oma = c(6.5, 11, 3.5, 1)
)

y.ub <- c(12, 12, 12, 12, 12)
text.y <- c(rep(8, 4), 8)
text.x <- c(0.1, 0.2, 0.4, 0.7, 0.9)
text.x.adj <- c(0.3, 0.4, 0.6, 0.5, 0.8)
model_colors <- c("steelblue", "goldenrod2", "firebrick2")
model_colors <- viridis::viridis(3)

for (m in c(1:3)) {
  for (j in c(1:length(object_par$f_vec_1))) {
    plot(1, 4,
      xlim = c(0, 1), ylim = c(0, y.ub[j]),
      type = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
    )

    ## Posterior density
    current_result <- eval(parse(text = paste("sim_model1_stan_sim",
      m,
      "_total",
      sep = ""
    )))
    for (i in c(1:n_sim)) {
      lines(density(current_result[[i]]$f[, j]),
        col = add.alpha(model_colors[m], alpha = 0.8), cex = 0.6, lwd = 0.8
      )
    }

    ## plot true value
    abline(v = object_par$f_vec_1[j], lwd = 1.8, lty = 2, col = "gray30")

    ## plot box
    box.wdt <- 1.5
    box.col <- "black"
    box(lty = "solid", col = box.col, lwd = 0.7)

    if (m == 3) {
      text(text.x.adj[j], text.y[j], text.x[j], cex = 2.4)
    }
  }
  ## bottom
  mtext(
    text = expression(f[1], f[2], f[3], f[4], f[5]), side = 1, line = 2.4, outer = TRUE,
    at = c(1:length(object_par$f_vec_1)) / length(object_par$f_vec_1) - 0.1,
    cex = 2.4
  )
  mtext(
    text = "Latent rank variables", side = 1, line = 5.3, outer = TRUE, cex = 2.7,
    at = 0.5
  )
  ## top
  mtext(
    text = "Simulated from M1", side = 3, line = 0, outer = TRUE, cex = 3,
    at = 0.5, font = 2
  )
  ## left
  mtext(
    text = rev(c("C-HP  ", "C-DCHP", "C-MMHP")), side = 2, line = 0.5, outer = TRUE,
    at = c(0:2) / 3 + 0.15,
    cex = 2, las = 2
  )
}

for (m in c(1:3)) {
  for (j in c(1:length(object_par$f_vec_1))) {
    plot(1, 4,
      xlim = c(0, 1), ylim = c(0, y.ub[j]),
      type = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
    )

    ## Posterior density
    current_result <- eval(parse(text = paste("sim_model2_stan_sim",
      m,
      "_total",
      sep = ""
    )))
    for (i in c(1:n_sim)) {
      lines(density(current_result[[i]]$f[, j]),
        col = add.alpha(model_colors[m], alpha = 0.8), cex = 0.6, lwd = 0.8
      )
    }

    ## plot true value
    abline(v = object_par$f_vec_1[j], lwd = 1.8, lty = 2, col = "gray30")

    ## plot box
    box.wdt <- 1.5
    box.col <- "black"
    box(lty = "solid", col = box.col, lwd = 0.7)

    if (m == 3) {
      text(text.x.adj[j], text.y[j], text.x[j], cex = 2.4)
    }
  }
  ## bottom
  mtext(
    text = expression(f[1], f[2], f[3], f[4], f[5]), side = 1, line = 2.4, outer = TRUE,
    at = c(1:length(object_par$f_vec_1)) / length(object_par$f_vec_1) - 0.1,
    cex = 2.4
  )
  mtext(
    text = "Latent rank variables", side = 1, line = 5.3, outer = TRUE, cex = 2.7,
    at = 0.5
  )
  ## top
  mtext(
    text = "Simulated from M2", side = 3, line = 0, outer = TRUE, cex = 3,
    at = 0.5, font = 2
  )
  ## left
  mtext(
    text = rev(c("C-HP  ", "C-DCHP", "C-MMHP")), side = 2, line = 0.5, outer = TRUE,
    at = c(0:2) / 3 + 0.15,
    cex = 2, las = 2
  )
}
```

Similarly, we can then repeat this for Fig 4b also for both simulated 
data examples, with slightly more difficulty.

```{r process m1 for fig4b, include=extra_plots, eval=extra_plots}
plot.s <- 1 # 7#9
plot.i <- 4
plot.j <- 5

cur_i <- plot.i
cur_j <- plot.j

load(paste0(m1_path, m1_files[plot.s]))
clean_sim_data_m1 <- cleanSimulationData(
  raw_data = sim_model1_data[[1]],
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)
# this loads in sim data too

temp.t <- sim_model1_data[[1]]$day_hour[which(sim_model1_data[[1]]$start == cur_i &
  sim_model1_data[[1]]$end == cur_j)]
# temp.t <- test.mmhp$tau
current.n <- length(temp.t) - 1
time.segment <- seq(0, tail(temp.t, 1), length.out = 5000)

### Preprocess the model inference result
## model 1
#########
model1_par_est <- list(
  lambda0 =
    mean(sim_model1_stan_sim1_total[[plot.s]]$lambda0),
  eta_1 = mean(sim_model1_stan_sim1_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model1_stan_sim1_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model1_stan_sim1_total[[plot.s]]$eta_3),
  beta = mean(sim_model1_stan_sim1_total[[plot.s]]$beta),
  f = apply(sim_model1_stan_sim1_total[[plot.s]]$f, 2, mean)
)
lambda.m1 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 =
      model1_par_est$lambda0,
    alpha =
      model1_fn$alpha.fun(
        model1_par_est$f[plot.i],
        model1_par_est$f[plot.j],
        model1_par_est$eta_1,
        model1_par_est$eta_2,
        model1_par_est$eta_3
      ),
    beta = model1_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## model2
#############
model2_par_est <- list(
  gamma = apply(
    sim_model1_stan_sim2_total[[plot.s]]$gamma,
    2, mean
  ),
  zeta = apply(
    sim_model1_stan_sim2_total[[plot.s]]$zeta,
    2, mean
  ),
  eta_1 = mean(sim_model1_stan_sim2_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model1_stan_sim2_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model1_stan_sim2_total[[plot.s]]$eta_3),
  beta = mean(sim_model1_stan_sim2_total[[plot.s]]$beta),
  f = apply(
    sim_model1_stan_sim2_total[[plot.s]]$f,
    2, mean
  )
)

lambda.m2 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 = model2_par_est$gamma[plot.i] +
      model2_par_est$zeta[plot.j],
    alpha = model1_fn$alpha.fun(
      model2_par_est$f[plot.i],
      model2_par_est$f[plot.j],
      model2_par_est$eta_1,
      model2_par_est$eta_2,
      model2_par_est$eta_3
    ),
    beta = model2_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## Model3
#############
####
lam0_vec <- apply(sim_model1_stan_sim3_total[[plot.s]]$lambda0, 2, mean)
lam1_vec <- apply(sim_model1_stan_sim3_total[[plot.s]]$lambda1, 2, mean)
lam0_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
lam1_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)

### loaded in sim data again above

for (i in seq_along(lam0_vec)) {
  row_id <- clean_sim_data_m1$I_fit[i]
  col_id <- clean_sim_data_m1$J_fit[i]
  lam0_par_est[row_id, col_id] <- lam0_vec[i]
  lam1_par_est[row_id, col_id] <- lam1_vec[i]
}
####
mmhp_par_est <- list(
  lambda0 = lam0_par_est,
  lambda1 = lam1_par_est,
  eta_1 = mean(sim_model1_stan_sim3_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model1_stan_sim3_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model1_stan_sim3_total[[plot.s]]$eta_3),
  beta = mean(sim_model1_stan_sim3_total[[plot.s]]$beta),
  f = apply(sim_model1_stan_sim3_total[[plot.s]]$f, 2, mean)
)

object_hat <- list(
  lambda0 = mmhp_par_est$lambda0[cur_i, cur_j],
  lambda1 = mmhp_par_est$lambda1[cur_i, cur_j],
  alpha = model3_fn$alpha.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_1,
    mmhp_par_est$eta_2
  ),
  beta = mmhp_par_est$beta,
  q1 = model3_fn$q1.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  ),
  q2 = model3_fn$q0.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  )
)

state_est_fits <- list.files(m1_path, pattern = "fit123_state_est_M1*")
load(paste0(m1_path, state_est_fits[plot.s]))

state.est.latent.mmhp <- interpolate_state_est_lst[plot.i, plot.j][[1]]
state.est.latent.mmhp.new <- fixInterpolateState(state.est.latent.mmhp,
  termination = 200
)

step.fun.est <- stepfun(
  state.est.latent.mmhp$x.hat,
  2 - state.est.latent.mmhp$z.hat
)

lambda.m3 <- mmhpIntensityNumeric(
  params = object_hat,
  t = temp.t,
  time.vec = time.segment,
  latent.vec = step.fun.est(time.segment)
)

object_true <- lapply(object_matrix, function(x) x[plot.i, plot.j])
names(object_true) <- c("lambda0", "lambda1", "alpha", "beta")

lambda.true <- uniHawkesIntensityNumeric(
  object = object_true,
  events = temp.t,
  time.vec = time.segment
)
```

```{r fig 4b M1 sim, include=extra_plots, eval=extra_plots}
source(here("lib", "mod_drawHawkesIntensityPaper.R"))
y.ub <- c(27, 27, 27) # c(130,72,31)
x_events <- 135 # 150
# model_colors <- c("steelblue", "goldenrod2", "firebrick2")
model_colors <- viridis::viridis(3)
layout(matrix(c(1:6), 6, 1), heights = rep(c(2.5, 1), 3))
layout(matrix(c(1:6), 6, 1), heights = rep(c(7.5, 1), 3))
my.xlim <- 140 # 160#200#130
my.long.x <- 140 # 205#134.5
legend.x <- c(110, 110, 110) # c(135,135,135)
legend.y <- # c(34,32,32)
  legend.y <- c(20, 20, 20) # c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 # 3
line.alpha <- c(0.9, 0.6)
line.wdth <- c(2, 2.5)
negative.col <- add.alpha("lightskyblue", 0.6)
positive.col <- add.alpha("tomato", 0.6)
# negative.col <- viridis::viridis(5,alpha = 0.6)[1]
# positive.col <- viridis::viridis(5,alpha = 0.6)[4]
par(
  mar = c(0, 0.3, 0, 0.1), oma = c(3, 3.5, 4, 0),
  tcl = 0.2, mgp = c(0.5, 0, 0), xpd = TRUE
)

# model1
############
true.object <- object_true
# lambda.true$lambda.t <- lambda.est
# lambda.true$time.vec <- lambda.m1$time.vec

mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawHawkesIntensityPaper(
  lambda0 = model1_par_est$lambda0,
  alpha = model1_fn$alpha.fun(
    model1_par_est$f[plot.i],
    model1_par_est$f[plot.j],
    model1_par_est$eta_1,
    model1_par_est$eta_2,
    model1_par_est$eta_3
  ),
  beta = model1_par_est$beta,
  events = temp.t,
  color = add.alpha(model_colors[1],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)

## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(legend.x[1], legend.y[1], c("True"),
  col = c("black"),
  y.intersp = 0.88, x.intersp = 0.15, bty = "n",
  lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)
legend(legend.x[1] + 18, legend.y[1], c("C-HP"),
  col = c(model_colors[1]),
  y.intersp = 0.88, x.intersp = 0.15, bty = "n",
  lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
  xlim = c(0, my.xlim), ylim = c(-4, 10), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m1$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now ###
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)

axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

# model2
############
mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawHawkesIntensityPaper(
  lambda0 = model2_par_est$gamma[plot.i] +
    model2_par_est$zeta[plot.j],
  alpha = model1_fn$alpha.fun(
    model2_par_est$f[plot.i],
    model2_par_est$f[plot.j],
    model2_par_est$eta_1,
    model2_par_est$eta_2,
    model2_par_est$eta_3
  ),
  beta = model2_par_est$beta,
  events = temp.t,
  color = add.alpha(model_colors[2],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)
legend(legend.x[2], legend.y[2], "C-DCHP",
  col = model_colors[2],
  y.intersp = 0.88, x.intersp = 0.1, bty = "n",
  lty = 1, lwd = 5, cex = legend.cex, seg.len = 1
)

axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(legend.x[2] - 40,
  legend.y[2] + 2, c("State 0/1 events", "State change point"),
  col = c(NA, "red"),
  y.intersp = 0.85, x.intersp = -0.1, bty = "n",
  pch = c(NA, 4), pt.cex = c(NA, 2), cex = legend.cex,
  lty = c(NA, NA), lwd = c(NA, 4)
)
points(legend.x[2] - 35, 16, pch = 1, cex = 2, col = "blue")
points(legend.x[2] - 34, 16, pch = 16, cex = 2, col = "blue")

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t

plot(lambda.true$time.vec, delta.lambda,
  xlim = c(0, my.xlim), ylim = c(-3, 10), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

# model3
########
par(mar = c(0, 0.3, 1, 0.1))
mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawUniMMHPIntensityPaper(object_hat,
  simulation = list(
    x = state.est.latent.mmhp.new$x.hat,
    z = state.est.latent.mmhp.new$z.hat,
    tau = temp.t,
    lambda.max = max(lambda.m3)
  ),
  yupper = y.ub, add = TRUE,
  color = add.alpha(model_colors[3],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)
legend(legend.x[3], legend.y[3], "C-MMHP",
  col = model_colors[3],
  y.intersp = 0.88, x.intersp = 0.1, bty = "n",
  lty = 1, lwd = 5, cex = legend.cex, seg.len = 1
)
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(legend.x[3] - 40,
  legend.y[3], c("Overestimation", "Underestimation"),
  col = c(positive.col, negative.col),
  y.intersp = 0.88, x.intersp = 0.5, bty = "n",
  lty = c(1, 1), lwd = c(10, 10), cex = legend.cex, seg.len = 0.8
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
  xlim = c(0, my.xlim), ylim = c(-12, 8), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m3 - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

## top
mtext(text = "M1 Sim", side = 3, line = 0.8, outer = TRUE, cex = 2.5, font = 2)
## bottom
mtext(text = "Time", side = 1, line = 2, outer = TRUE, cex = 2.5)
## left
mtext(text = "Intensity", side = 2, line = 0.6, outer = TRUE, cex = 2.5)
```

```{r process m2 for fig4b, include=extra_plots, eval=extra_plots}
plot.s <- 1 # 7#9
plot.i <- 4
plot.j <- 5

cur_i <- plot.i
cur_j <- plot.j

load(paste0(m2_path, m2_files[plot.s]))
clean_sim_data_m2 <- cleanSimulationData(
  raw_data = sim_model2_data[[1]],
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)

temp.t <- sim_model2_data[[1]]$day_hour[which(sim_model2_data[[1]]$start == cur_i &
  sim_model2_data[[1]]$end == cur_j)]
# temp.t <- test.mmhp$tau
current.n <- length(temp.t) - 1
time.segment <- seq(0, tail(temp.t, 1), length.out = 5000)

### Preprocess the model inference result
## model 1
#########
model1_par_est <- list(
  lambda0 = mean(sim_model2_stan_sim1_total[[plot.s]]$lambda0),
  eta_1 = mean(sim_model2_stan_sim1_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model2_stan_sim1_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model2_stan_sim1_total[[plot.s]]$eta_3),
  beta = mean(sim_model2_stan_sim1_total[[plot.s]]$beta),
  f = apply(sim_model2_stan_sim1_total[[plot.s]]$f, 2, mean)
)
lambda.m1 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 =
      model1_par_est$lambda0,
    alpha =
      model1_fn$alpha.fun(
        model1_par_est$f[plot.i],
        model1_par_est$f[plot.j],
        model1_par_est$eta_1,
        model1_par_est$eta_2,
        model1_par_est$eta_3
      ),
    beta = model1_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## model2
#############
model2_par_est <- list(
  gamma = apply(
    sim_model2_stan_sim2_total[[plot.s]]$gamma,
    2, mean
  ),
  zeta = apply(sim_model2_stan_sim2_total[[plot.s]]$zeta, 2, mean),
  eta_1 = mean(sim_model2_stan_sim2_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model2_stan_sim2_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model2_stan_sim2_total[[plot.s]]$eta_3),
  beta = mean(sim_model2_stan_sim2_total[[plot.s]]$beta),
  f = apply(sim_model2_stan_sim2_total[[plot.s]]$f, 2, mean)
)

lambda.m2 <- uniHawkesIntensityNumeric(
  object = list(
    lambda0 = model2_par_est$gamma[plot.i] +
      model2_par_est$zeta[plot.j],
    alpha = model1_fn$alpha.fun(
      model2_par_est$f[plot.i],
      model2_par_est$f[plot.j],
      model2_par_est$eta_1,
      model2_par_est$eta_2,
      model2_par_est$eta_3
    ),
    beta = model2_par_est$beta
  ),
  events = temp.t,
  time.vec = time.segment
)

## Model3
#############
lam0_vec <- apply(sim_model2_stan_sim3_total[[plot.s]]$lambda0, 2, mean)
lam1_vec <- apply(sim_model2_stan_sim3_total[[plot.s]]$lambda1, 2, mean)

lam0_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
lam1_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)

for (i in seq_along(lam0_vec)) {
  row_id <- clean_sim_data_m2$I_fit
  col_id <- clean_sim_data_m2$J_fit
  lam0_par_est[row_id, col_id] <- lam0_vec[i]
  lam1_par_est[row_id, col_id] <- lam1_vec[i]
}

mmhp_par_est <- list(
  lambda0 = lam0_par_est,
  lambda1 = lam1_par_est,
  eta_1 = mean(sim_model2_stan_sim3_total[[plot.s]]$eta_1),
  eta_2 = mean(sim_model2_stan_sim3_total[[plot.s]]$eta_2),
  eta_3 = mean(sim_model2_stan_sim3_total[[plot.s]]$eta_3),
  beta = mean(sim_model2_stan_sim3_total[[plot.s]]$beta),
  f = apply(sim_model2_stan_sim3_total[[plot.s]]$f, 2, mean)
)

object_hat <- list(
  lambda0 = mmhp_par_est$lambda0[cur_i, cur_j],
  lambda1 = mmhp_par_est$lambda1[cur_i, cur_j],
  alpha = model3_fn$alpha.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_1,
    mmhp_par_est$eta_2
  ),
  beta = mmhp_par_est$beta,
  q1 = model3_fn$q1.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  ),
  q2 = model3_fn$q0.fun(
    mmhp_par_est$f[cur_i],
    mmhp_par_est$f[cur_j],
    mmhp_par_est$eta_3
  )
)

state_est_fits <- list.files(m2_path, pattern = "fit123_state_est_M2*")
load(paste0(m2_path, state_est_fits[plot.s]))

state.est.latent.mmhp <- interpolate_state_est_lst[plot.i, plot.j][[1]]
state.est.latent.mmhp.new <- fixInterpolateState(state.est.latent.mmhp,
  termination = 200
)

step.fun.est <- stepfun(
  state.est.latent.mmhp$x.hat,
  2 - state.est.latent.mmhp$z.hat
)

lambda.m3 <- mmhpIntensityNumeric(
  params = object_hat,
  t = temp.t,
  time.vec = time.segment,
  latent.vec = step.fun.est(time.segment)
)

object_true <- lapply(object_matrix, function(x) x[plot.i, plot.j])
names(object_true) <- c("lambda0", "lambda1", "alpha", "beta")

lambda.true <- uniHawkesIntensityNumeric(
  object = object_true,
  events = temp.t,
  time.vec = time.segment
)
```

```{r fig 4b M2 sim, include=extra_plots, eval=extra_plots}
source(here("lib", "mod_drawHawkesIntensityPaper.R"))
y.ub <- c(17, 17, 17) # c(130,72,31)
x_events <- 135 # 150
# model_colors <- c("steelblue", "goldenrod2", "firebrick2")
model_colors <- viridis::viridis(3)
layout(matrix(c(1:6), 6, 1), heights = rep(c(2.5, 1), 3))
layout(matrix(c(1:6), 6, 1), heights = rep(c(7.5, 1), 3))
my.xlim <- 140 # 160#200#130
my.long.x <- 140 # 205#134.5
legend.x <- c(50, 50, 50) # c(135,135,135)
legend.y <- # c(34,32,32)
  legend.y <- c(15, 12, 15) # c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 # 3
line.alpha <- c(0.9, 0.6)
line.wdth <- c(2, 2.5)
negative.col <- add.alpha("lightskyblue", 0.6)
positive.col <- add.alpha("tomato", 0.6)
# negative.col <- viridis::viridis(5,alpha = 0.6)[1]
# positive.col <- viridis::viridis(5,alpha = 0.6)[4]
par(
  mar = c(0, 0.3, 0, 0.1), oma = c(3, 3.5, 4, 0),
  tcl = 0.2, mgp = c(0.5, 0, 0), xpd = TRUE
)

# model1
############
true.object <- object_true
# lambda.true$lambda.t <- lambda.est
# lambda.true$time.vec <- lambda.m1$time.vec

mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawHawkesIntensityPaper(
  lambda0 = model1_par_est$lambda0,
  alpha = model1_fn$alpha.fun(
    model1_par_est$f[plot.i],
    model1_par_est$f[plot.j],
    model1_par_est$eta_1,
    model1_par_est$eta_2,
    model1_par_est$eta_3
  ),
  beta = model1_par_est$beta,
  events = temp.t,
  color = add.alpha(model_colors[1],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)

## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(legend.x[1], legend.y[1], c("True"),
  col = c("black"),
  y.intersp = 0.88, x.intersp = 0.15, bty = "n",
  lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)
legend(legend.x[1] + 18, legend.y[1], c("C-HP"),
  col = c(model_colors[1]),
  y.intersp = 0.88, x.intersp = 0.15, bty = "n",
  lty = c(1), lwd = c(5), cex = legend.cex, seg.len = 1
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
  xlim = c(0, my.xlim), ylim = c(-4, 10), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m1$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now ###
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)

axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

# model2
############
mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawHawkesIntensityPaper(
  lambda0 = model2_par_est$gamma[plot.i] +
    model2_par_est$zeta[plot.j],
  alpha = model1_fn$alpha.fun(
    model2_par_est$f[plot.i],
    model2_par_est$f[plot.j],
    model2_par_est$eta_1,
    model2_par_est$eta_2,
    model2_par_est$eta_3
  ),
  beta = model2_par_est$beta,
  events = temp.t,
  color = add.alpha(model_colors[2],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)
legend(legend.x[2], legend.y[2], "C-DCHP",
  col = model_colors[2],
  y.intersp = 0.88, x.intersp = 0.1, bty = "n",
  lty = 1, lwd = 5, cex = legend.cex, seg.len = 1
)

axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(10, legend.y[2] + 2, c("State 0/1 events", "State change point"),
  col = c(NA, "red"),
  y.intersp = 0.85, x.intersp = -0.1, bty = "n",
  pch = c(NA, 4), pt.cex = c(NA, 2), cex = legend.cex,
  lty = c(NA, NA), lwd = c(NA, 4)
)
points(14, 8 + 2, pch = 1, cex = 2, col = "blue")
points(16, 8 + 2, pch = 16, cex = 2, col = "blue")

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t

plot(lambda.true$time.vec, delta.lambda,
  xlim = c(0, my.xlim), ylim = c(-3, 10), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m2$lambda.t - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

# model3
########
par(mar = c(0, 0.3, 1, 0.1))
mod_drawHawkesIntensityPaper(
  lambda0 = true.object$lambda0,
  alpha = true.object$alpha,
  beta = true.object$beta,
  events = temp.t,
  add = FALSE,
  yupper = y.ub[1],
  color = add.alpha("black", alpha = line.alpha[1]),
  line.width = line.wdth[1],
  y.ratio = -0.05, min.y = -2, min.x = 0, max.x = my.xlim,
  title_name = paste(plot.i, "->", plot.j),
  title.cex = 2,
  box.type = "n"
)
drawUniMMHPIntensityPaper(object_hat,
  simulation = list(
    x = state.est.latent.mmhp.new$x.hat,
    z = state.est.latent.mmhp.new$z.hat,
    tau = temp.t,
    lambda.max = max(lambda.m3)
  ),
  yupper = y.ub, add = TRUE,
  color = add.alpha(model_colors[3],
    alpha = line.alpha[2]
  ),
  line.width = line.wdth[2]
)
legend(legend.x[3], legend.y[3], "C-MMHP",
  col = model_colors[3],
  y.intersp = 0.88, x.intersp = 0.1, bty = "n",
  lty = 1, lwd = 5, cex = legend.cex, seg.len = 1
)
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0.5, tick = T, labels = rep("", 2), lwd = 0.5, lwd.ticks = 0, lty = 3
)

legend(10, legend.y[3], c("Overestimation", "Underestimation"),
  col = c(positive.col, negative.col),
  y.intersp = 0.88, x.intersp = 0.5, bty = "n",
  lty = c(1, 1), lwd = c(10, 10), cex = legend.cex, seg.len = 0.8
)

## delta lambda
par(mar = c(0, 0.3, 0, 0.1))
plot(0, 0,
  xlim = c(0, my.xlim), ylim = c(-12, 8), type = "n",
  bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
)
delta.lambda <- lambda.m3 - lambda.true$lambda.t
delta.positive <- delta.lambda
delta.positive[delta.positive < 0] <- 0
delta.negative <- delta.lambda
delta.negative[delta.negative > 0] <- 0
### hiding these for now
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.negative, 0),
  col = negative.col, border = NA
)
polygon(c(0, lambda.true$time.vec, x_events), # 200
  c(0, delta.positive, 0),
  col = positive.col, border = NA
)
# segments(0,0,my.xlim,0,col="gray45")
## draw box
axis(1,
  at = c(-2, my.long.x), col = "gray35",
  line = 0, tick = T, labels = rep("", 2), lwd = 0.7, lwd.ticks = 0
)

## top
mtext(text = "M2 Sim", side = 3, line = 0.8, outer = TRUE, cex = 2.5, font = 2)
## bottom
mtext(text = "Time", side = 1, line = 2, outer = TRUE, cex = 2.5)
## left
mtext(text = "Intensity", side = 2, line = 0.6, outer = TRUE, cex = 2.5)
```

# Figure 5:

To be updated!

```{r data for fig 5,cache=TRUE, eval=revision_plots}
n_sim <- 25

num_nodes <- 10

object_par <- list(f_vec_1 = seq(from = 0.1, to = 0.9,
                                 length.out = num_nodes))
### TO DO

sim_data_path <- here("output", "revisions", "sim_m3", "june_18", "/")

## Initialize
N_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))

## Iteration on 50 simulations
sims <- 1:25
sims <- sims[c(-12, -22, -23, -24)]

for (s in sims) {
  print(s)
  load(paste0(sim_data_path, "sim_model3_fit123_", s, ".RData"))
  load(paste(sim_data_path, "fit123_state_est_", s, ".RData", sep = ""))
  clean_sim_data <- cleanSimulationData(
    raw_data = sim_model3_data,
    cut_off = cut_off,
    N = length(object_par$f_vec_1)
  )

  N_array[s, , ] <- clean_sim_data$N_count

  ## model 1
  #########
  ### transform from cmdstan here
  f_draws <- sim_model3_stan_sim1 %>% 
  select(starts_with("f"))

  ###
  model1_par_est <- list(
    lambda0 = mean(sim_model3_stan_sim1$lambda0),
    eta_1 = mean(sim_model3_stan_sim1$eta_1),
    eta_2 = mean(sim_model3_stan_sim1$eta_2),
    eta_3 = mean(sim_model3_stan_sim1$eta_3),
    beta = mean(sim_model3_stan_sim1$beta),
    f = apply(f_draws, 2, mean)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
      m1_object <- list(
        lambda0 = model1_par_est$lambda0,
        alpha = model1_fn$alpha.fun(
          model1_par_est$f[i],
          model1_par_est$f[j],
          model1_par_est$eta_1,
          model1_par_est$eta_2,
          model1_par_est$eta_3
        ),
        beta = model1_par_est$beta
      )
      Lambda_model1_array[s, i, j] <-
        uniHawkesIntegralIntensity(
          object = m1_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
      Lambda_test <- uniHawkesCompensator(
        object = m1_object,
        events = current_t
      )
      KS_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
      KS_p_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
      PR_model1_array[s, i, j] <- uniHawkesPearsonResidual(
        object = m1_object,
        events = current_t,
        termination =
          tail(current_t, 1)
      )
    }
  }

  ## model2
  #############
  ### convert to cmdstan
  gamma_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("gamma"))
  zeta_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("zeta"))
  f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

  model2_par_est <- list(
    gamma = apply(gamma_draws, 2, mean),
    zeta = apply(zeta_draws, 2, mean),
    eta_1 = mean(sim_model3_stan_sim2$eta_1),
    eta_2 = mean(sim_model3_stan_sim2$eta_2),
    eta_3 = mean(sim_model3_stan_sim2$eta_3),
    beta = mean(sim_model3_stan_sim2$beta),
    f = apply(f_draws, 2, mean)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
      m2_object <- list(
        lambda0 = model2_par_est$gamma[i] +
          model2_par_est$zeta[j],
        alpha = model1_fn$alpha.fun(
          model2_par_est$f[i],
          model2_par_est$f[j],
          model2_par_est$eta_1,
          model2_par_est$eta_2,
          model2_par_est$eta_3
        ),
        beta = model2_par_est$beta
      )
      Lambda_model2_array[s, i, j] <-
        uniHawkesIntegralIntensity(
          object = m1_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
      Lambda_test <- uniHawkesCompensator(
        object = m2_object,
        events = current_t
      )
      KS_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
      KS_p_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
      PR_model2_array[s, i, j] <-
        uniHawkesPearsonResidual(
          object = m2_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
    }
  }

  ## Model3 & true model
  ######
  ### cmdstan part
  lam0_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("lambda0"))
  lam1_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("lambda1"))
  f_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("f"))

  if(ncol(lam0_draws) != sum(clean_sim_data$N_count >2)) {
    stop("Mismatch between params and event pairs")
  }

  lambda_0_est <- apply(lam0_draws, 2, mean)
  lambda_1_est <- apply(lam1_draws, 2, mean)
  ## then put these into a matrix
  lam0_par_est <- matrix(0,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  )
  lam1_par_est <- matrix(0,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  )
  for (i in seq_along(lambda_0_est)) {
    row_id <- clean_sim_data$I_fit[i]
    col_id <- clean_sim_data$J_fit[i]
    lam0_par_est[row_id, col_id] <- lambda_0_est[i]
    lam1_par_est[row_id, col_id] <- lambda_1_est[i]
  }
  mmhp_par_est <- list(
    lambda0 = lam0_par_est,
    lambda1 = lam1_par_est,
    eta_1 = mean(sim_model3_stan_sim3$eta_1),
    eta_2 = mean(sim_model3_stan_sim3$eta_2),
    eta_3 = mean(sim_model3_stan_sim3$eta_3),
    beta = mean(sim_model3_stan_sim3$beta),
    f = apply(f_draws, 2, mean)
  )
  clean_sim_data <- cleanSimulationData(
    raw_data = sim_model3_data,
    cut_off = cut_off,
    N = length(object_par$f_vec_1)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      test.mmhp <- sim_model3_data$mmhp_matrix[i, j][[1]]
      temp.t <- test.mmhp$tau
      current.n <- length(temp.t) - 1
      time.segment <- seq(0, tail(temp.t, 1), length.out = 10000)

      object_hat <- list(
        lambda0 = mmhp_par_est$lambda0[i, j],
        lambda1 = mmhp_par_est$lambda1[i, j],
        alpha = model3_fn$alpha.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_1,
          mmhp_par_est$eta_2
        ),
        beta = mmhp_par_est$beta,
        q1 = model3_fn$q1.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_3
        ),
        q2 = model3_fn$q0.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_3
        )
      )

      object_true <- lapply(object_matrix, function(x) x[i, j])
      names(object_true) <- c("lambda0", "lambda1",
                              "alpha", "beta", "q1", "q2")

      ## est Lmabda and [R]
      Lambda.test <- mmhpCompensator(
        params = object_hat,
        t = temp.t,
        pzt = 2 -
          event_state_est_lst[i, j][[1]]$zt_v,
        if.pzt = FALSE
      )
      KS_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$statistic
      KS_p_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$p.value
      state.est.latent.mmhp <- interpolate_state_est_lst[i, j][[1]]
      step.fun.est <- stepfun(
        state.est.latent.mmhp$x.hat,
        2 - state.est.latent.mmhp$z.hat
      )
      est.intensity <- mmhpIntensityNumeric(
        params = object_hat,
        t = temp.t[-1],
        time.vec = time.segment,
        latent.vec =
          step.fun.est(time.segment)
      )
      est.intensity.events <- mmhpIntensityAtEvents(
        params = object_hat,
        t = temp.t,
        latent_z =
          event_state_est_lst[i, j][[1]]$zt_v
      )
      Lambda_model3_array[s, i, j] <-
        sum(est.intensity) * (time.segment[2] - time.segment[1])
      PR_model3_array[s, i, j] <- sum(1 / sqrt(est.intensity.events)) -
        sum(sqrt(est.intensity)) * (time.segment[2] - time.segment[1])

      ## Under the true model
      Lambda.test.true <- mmhpCompensator(
        params = object_true,
        t = temp.t,
        pzt = 2 -
          test.mmhp$zt[2:length(test.mmhp$zt)],
        if.pzt = FALSE
      )
      KS_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$statistic
      KS_p_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$p.value
      true.intensity <-
        mmhpTrueIntensityNumeric(
          params = object_true,
          t = temp.t,
          latent = list(
            ## these are the truth
            x = fixStateTransition(test.mmhp)$x,
            z = fixStateTransition(test.mmhp)$z
          ),
          time.vec = time.segment
        )
      true.intensity.events <- mmhpIntensityAtEvents(
        params = object_true,
        t = temp.t,
        latent_z = test.mmhp$zt
      )
      Lambda_true_array[s, i, j] <- sum(true.intensity$lambda.t) *
        (time.segment[2] - time.segment[1])
      PR_true_array[s, i, j] <- sum(1 / sqrt(true.intensity.events)) -
        sum(sqrt(true.intensity$lambda.t)) * 
        (time.segment[2] - time.segment[1])
    }
  }
  rm(sim_model3_data)
}

```

```{r fig5a-residual-ks, fig.height=7, eval=original_plots}
# cols_plot <- col_df %>% filter(method %in% c("C-HP","C-DCHP","C-MMHP",
#                                              "True"))

cols_plot <- col_df %>%
  filter(method %in% c(
    "C-HP", "C-DCHP", "C-MMHP",
    "True"
  )) %>%
  arrange(factor(method, levels = c("C-HP", "C-DCHP", "C-MMHP", "True"))) %>%
  pull(cols)
# load("../output/sim_model3_fit123_residuals.RData")
par(mfrow = c(5, 5), mar = c(0.2, 0.2, 0.2, 0.2), oma = c(2.4, 2.4, 2.9, 0.5))
for (i in c(5:1)) {
  for (j in c(5:1)) {
    if (i == j) {
      plot(1, 1,
        type = "n", xlim = c(0, 1), ylim = c(0, 1),
        bty = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", axes = FALSE
      )
      if (i == 5) {
        polygon(c(0, 0, 0.15, 0.15), c(0.72, 0.8, 0.8, 0.72) + 0.04,
          col = cols_plot[1], border = NA
        )
        polygon(c(0, 0, 0.15, 0.15), c(0.72, 0.8, 0.8, 0.72) + 0.04 - 0.22,
          col = cols_plot[2], border = NA
        )
        polygon(c(0, 0, 0.15, 0.15), c(0.72, 0.8, 0.8, 0.72) + 0.04 - 0.44,
          col = cols_plot[3], border = NA
        )
        polygon(c(0, 0, 0.15, 0.15), c(0.72, 0.8, 0.8, 0.72) + 0.04 - 0.66,
          col = cols_plot[4], border = NA
        )
        text(c(0.43, 0.58, 0.61, 0.4), seq(0.8, 0.14, -0.22),
          c("C-HP", "C-DCHP", "C-MMHP", "True"),
          cex = 1.7
        )
      }
    } else {
      # PR <- c((N_array-Lambda_model1_array)[,i,j],(N_array-Lambda_model2_array)[,i,j],
      #         (N_array-Lambda_model3_array)[,i,j],(N_array-Lambda_true_array)[,i,j])
      PR <- c(
        KS_model1_array[, i, j], KS_model2_array[, i, j],
        KS_model3_array[, i, j], KS_true_array[, i, j]
      )
      ### try pr scores instead
      # PR <- c(PR_model1_array[,i,j],PR_model2_array[,i,j],
      #         PR_model3_array[,i,j],PR_true_array[,i,j])
      Type <- c(rep(1, n_sim), rep(2, n_sim), rep(3, n_sim), rep(4, n_sim))
      boxplot(PR ~ Type,
        col = cols_plot,
        # col = c("steelblue", "goldenrod2", "firebrick2", "chartreuse3"),
        xlab = "", ylab = "", xaxt = "n", yaxt = "n"
      )
      # names = c("C-HP", "C-HP-DC", "C-MMHP", "True")
    }
  }
}
mtext(
  text = c(1:5), side = 2, line = 0.5, outer = TRUE, cex = 2, las = 2,
  at = c(0:4) / 5 + 0.1
)
mtext(
  text = c(5:1), side = 1, line = 1.5, outer = TRUE, cex = 2,
  at = c(0:4) / 5 + 0.1
)
mtext(text = "(a)", side = 3, line = 0.5, outer = TRUE, cex = 2.5, font = 2)
```

New potential Figure 5a plots.

```{r log-p-value}
cols_plot <- col_df %>% 
  filter(method %in% c("C-HP","C-DCHP","C-MMHP", "True")) %>%
  arrange(factor(method, levels = c("C-HP", "C-DCHP", "C-MMHP", "True"))) %>%
  pull(cols)

m1_ks_p <- tibble(Method = "C-HP",
                  ks = as.vector(KS_p_model1_array[sims, ,]))

m2_ks_p <- tibble(Method = "C-DCHP",
                  ks = as.vector(KS_p_model2_array[sims, ,]))

m3_ks_p <- tibble(Method = "C-MMHP",
                  ks = as.vector(KS_p_model3_array[sims, ,]))

true_ks_p <- tibble(Method = "True",
                    ks = as.vector(KS_p_true_array[sims, ,]))

all_ks_p <- m1_ks_p %>% 
  bind_rows(m2_ks_p) %>% 
  bind_rows(m3_ks_p) %>% 
  bind_rows(true_ks_p)

all_ks_p$Method <- factor(all_ks_p$Method, levels = c("C-HP",
                                                      "C-DCHP",
                                                      "C-MMHP",
                                                      "True") )

exp_fcn <- function(x) exp(-x)

all_ks_p %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density.., fill = Method), bins = 50) +
  scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value") +
  stat_function(fun = exp_fcn, colour = "red") +
  facet_wrap(~Method) +
  theme_minimal()

```

Inferred ranking boxplots against true rankings.

```{r posterior-mean-ranks}
sim_data_path <- here("output", "revisions", "sim_m3", "june_18", "/")

num_nodes <- 10

# object_par$f_vec_1

sim_ranks_mean <- tibble()
sim_ranks <- tibble()

for(curr_sim in sims) {
  load(paste0(sim_data_path,
              "sim_model3_fit123_",
              curr_sim,
              ".RData"))

  f_m1 <- sim_model3_stan_sim1 %>% select(starts_with("f"))
  m1_ranks <- apply(f_m1, 2, mean)

  f_m1_sims <- f_m1 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")
  f_m1_sims$model <- "C-HP"

  f_m2 <- sim_model3_stan_sim2 %>% select(starts_with("f"))
  m2_ranks <- apply(f_m2, 2, mean)

  f_m2_sims <- f_m2 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m2_sims$model <- "C-DCHP"

  f_m3 <- sim_model3_stan_sim3 %>% select(starts_with("f"))
  m3_ranks <- apply(f_m3, 2, mean)

  f_m3_sims <- f_m3 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m3_sims$model <- "C-MMHP"

  curr_mean_ranks <- tibble(model = c(rep("C-HP", num_nodes),
                                 rep("C-DCHP", num_nodes),
                                 rep("C-MMHP", num_nodes)),
                       ranks = c(m1_ranks,
                                 m2_ranks,
                                 m3_ranks),
                       node = rep(1:num_nodes, 3),
                       true_rank = rep(object_par$f_vec_1, 3),
                       sim_id = curr_sim)
  sim_ranks_mean <- sim_ranks_mean %>% 
    bind_rows(curr_mean_ranks)

  sim_ranks <- sim_ranks %>% 
    bind_rows(f_m1_sims, f_m2_sims, f_m3_sims)

}

sim_ranks$node <- factor(sim_ranks$node, levels = 1:num_nodes)
sim_ranks_mean$node <- factor(sim_ranks_mean$node, levels = 1:num_nodes)

# sim_ranks$node

sim_ranks_mean$model <- factor(sim_ranks_mean$model,
                               levels = c("C-HP", "C-DCHP", "C-MMHP"))

sim_ranks_mean %>% 
  # filter(model == "C-MMHP") %>% 
  ggplot(aes(node, ranks, fill = model)) + 
  geom_boxplot() +
  scale_fill_manual(values = cols_plot, guide = 'none') +
  facet_wrap(~model) +
  geom_point(aes(node, true_rank, colour = "True Value"),
             show.legend = TRUE) +
  labs(title = "Mean rank from each simulation") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")

```

# Figure 5b: Simulation Spearman Rank

Need to update this again for the final simulation parameter setting.

```{r fig-5b-spearman-sim, fig.height=7}
sim_files <- list.files(here("output", "revisions", "rank_sims/"), 
                        pattern = "rank_sim")

all_sims <- list()
for (i in seq_along(sim_files)) {
  single_fit <- readRDS(here(
    "output", "revisions", "rank_sims",
    sim_files[i]
  ))
  all_sims[[i]] <- single_fit
}

spear <- function(sim_data) {
  truth <- sim_data$truth
  sim_data %>%
    select(m1:glicko) %>%
    purrr::map(~ cor.test(truth, .x,
      method = "spearman"
    )$estimate) %>%
    enframe() %>%
    unnest(cols = c(value))
}

sim_data <- map_dfr(all_sims, spear)

name <- sim_data$name

sim_data$name <- case_when(
  name == "isi" ~ "I&SI",
  name == "m1" ~ "C-HP",
  name == "m2" ~ "C-DCHP",
  name == "m3" ~ "C-MMHP",
  name == "agg" ~ "AggRank",
  name == "glicko" ~ "Glicko"
)

sim_data$name <- factor(sim_data$name, levels = c(
  "I&SI",
  "AggRank",
  "Glicko",
  "C-HP",
  "C-DCHP",
  "C-MMHP"
))

plot_cols <- col_df %>%
  filter(method %in% levels(sim_data$name)) %>%
  pull(cols)

sim_data %>%
  rename(Method = name, corr = value) %>%
  ggplot(aes(Method, corr)) +
  geom_boxplot(aes(fill = Method), alpha = 0.7) +
  scale_fill_manual(values = plot_cols) +
  labs(y = "Rank Correlation") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("(b)")
```

# Figure 6: Latent Rank Cohort comparison

## Figure 6a: Rank vs I&SI

```{r fig-6a-rank-vs-isi}
### Rank Comparison for Single Cohort ####
# current_cohort <- 5

load(paste(data_path, cohort_names[current_cohort],
  "/agg_rank_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
load(paste(data_path, cohort_names[current_cohort],
  "/cohort_hp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
load(paste(data_path, cohort_names[current_cohort],
  "/cohort_dchp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
load(paste(data_path, cohort_names[current_cohort],
  "/cohort_mmhp_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))

no_method <- 4
plot_latent_df <- data.frame(
  Methods = rep(1:no_method, each = 1000 * mice_number),
  Mice = rep(rep(1:mice_number, each = 1000), no_method),
  force = rep(0, 1000 * mice_number * no_method)
)
## agg
plot_latent_df$force[(1:(1000 * mice_number))] <-
  as.vector(sim_agg_rank$x[1001:2000, ])

## M1
plot_latent_df$force[(1:(1000 * mice_number)) +
                       1000 * mice_number] <- as.vector(sim_cohort_hp$f)

## M2
plot_latent_df$force[(1:(1000 * mice_number)) + 
                       2000 * mice_number] <- as.vector(sim_cohort_dchp$f)

## M3
plot_latent_df$force[(1:(1000 * mice_number)) + 
                       3000 * mice_number] <- as.vector(sim_cohort_mmhp$f)

plot_latent_df$Mice <- factor(plot_latent_df$Mice,
  levels = expert_rank_10[[current_cohort]]
)
plot_latent_df$Method <- factor(plot_latent_df$Method,
  labels = c("AggRanking", "C-HP", "C-DCHP", "C-MMHP")
)

# png(paste(plot_path,"real_rank_one_cohort.png",sep=""),width=600,height=300)
plot_cols <- col_df %>%
  filter(method %in% c(
    "AggRank", "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

p1 <- ggplot(
  plot_latent_df,
  aes(x = Mice, y = force, fill = Method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_text(size = 10),
    legend.position = "none"
  ) +
  facet_grid(cols = vars(Method)) +
  xlab("Mice (ranked by I&SI method)") +
  ylab("Latent rank")
p1
# dev.off()
```

## Figure 6b: Spearman rank correlation

```{r fig-6b-overall-rank-correlation}
load("../output/weighted_rank_data.RData")
# png(paste(plot_path,"real_rank_all_cohort.png",sep=""),width=600,height=300)
no_method <- 4
plot_cor_df <- data.frame(
  Method = rep(c(1:no_method), each = 10),
  cor = as.vector(spearman_df)
)
plot_cor_df$Method <- factor(plot_cor_df$Method,
  labels = c("Agg-rank", "C-HP", "C-DCHP", "C-MMHP")
)
p2 <- ggplot(plot_cor_df, aes(x = "", y = cor, fill = Method)) +
  geom_boxplot(alpha = 0.7, width = 0.5) +
  scale_fill_manual(values = plot_cols) +
  ggtitle("(b)") +
  ylim(0, 1) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.y = element_text(size = 11),
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  facet_grid(cols = vars(Method)) +
  ylab("Spearman rank correlation")
p2
```

# Figure 7: Pearson residual matrices

```{r fig-7-prMatrix-Cohort3, fig.width=7, fig.height=7}
current_cohort <- 3
## 2 and 8 maybe look better
## 3 probably looks the best

m1_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/chp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

m2_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/dchp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

m3_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/cmmhp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

indep_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/immhp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

matrix_lst_plot <- list(
  indep_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ],
  m1_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ],
  m2_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ],
  m3_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ]
)
# cat("Cohort ",current_cohort,"\n",sep="")
myMultiMatrixPlot(
  X = matrix_lst_plot, no_matrix = 4, n_row = 2,
  xLabels = expert_rank_10[[current_cohort]],
  yLabels = rev(expert_rank_10[[current_cohort]]),
  min = -80, max = 80, axis_cex = 2, title_cex = 1.8,
  colorPalette = "RdBu", if.emp = FALSE,
  # legend.mar=c(0.5,.5,0.5,0.5),
  title_lst = list("I-MMHP", "C-HP", "C-DCHP", "C-MMHP"),
  # title_lst = list("I-MMHP"),
  col_axis = c(-80, -40, 0, 40, 80), fake_matrix = FALSE,
  matrix.mar = c(2.5, 2.5, 2.5, 1)
)
# mtext(text="(a)",side=3,line=0.5,outer=TRUE,cex=2.5,font=2)
```

```{r fig-7b-matrix-structure-score, fig.width=7, fig.height=7}
library(NMF)
source(here("lib", "residualStructureScore.R"))

struct_scores <- tibble()
for (current_cohort in 1:10) {
  m1_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
    "/chp_pr_matrix_", cohort_names[current_cohort],
    ".RDS",
    sep = ""
  ))
  ##
  ## remove animals which are never fought or don't ever fight
  prob_row <- which(rowSums(m1_pr) == 0)
  prob_col <- which(colSums(m1_pr) == 0)
  prob_id <- NA
  prob_id <- c(prob_row, prob_col)
  # print(prob_id)
  if (length(prob_id) != 0) {
    m1_pr <- m1_pr[-prob_id, -prob_id]
  }

  ##
  m2_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
    "/dchp_pr_matrix_", cohort_names[current_cohort],
    ".RDS",
    sep = ""
  ))
  if (length(prob_id) != 0) {
    m2_pr <- m2_pr[-prob_id, -prob_id]
  }

  m3_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
    "/cmmhp_pr_matrix_", cohort_names[current_cohort],
    ".RDS",
    sep = ""
  ))
  if (length(prob_id) != 0) {
    m3_pr <- m3_pr[-prob_id, -prob_id]
  }
  indep_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
    "/immhp_pr_matrix_", cohort_names[current_cohort],
    ".RDS",
    sep = ""
  ))
  # print(range(as.vector(indep_pr)))
  if (length(prob_id) != 0) {
    indep_pr <- indep_pr[-prob_id, -prob_id]
  }
  if (current_cohort %in% c(8, 10)) {
    # one additional row
    # print("in here")
    new_prob_row <- which(rowSums(m1_pr) == 0)
    new_prob_col <- which(colSums(m1_pr) == 0)
    prob_id <- c(new_prob_row, new_prob_col)
    if (length(prob_id) != 0) {
      m1_pr <- m1_pr[-prob_id, -prob_id]
      m2_pr <- m2_pr[-prob_id, -prob_id]
      m3_pr <- m3_pr[-prob_id, -prob_id]
      indep_pr <- indep_pr[-prob_id, -prob_id]
    }
  }

  m1_score <- positiveNegativeApproxError(m1_pr)
  m2_score <- positiveNegativeApproxError(m2_pr)
  m3_score <- positiveNegativeApproxError(m3_pr) 
  ## smallest value which is good
  ind_score <- positiveNegativeApproxError(indep_pr)
  ## this doesn't make sense because all negative
  ## values
  current_score <- tibble(m1_score, m2_score, m3_score, ind_score)
  names(current_score) <- c("C-HP", "C-DCHP", "C-MMHP", "I-MMHP")
  current_score$type <- c("Positive Residuals", "Negative Residuals")
  current_score$Cohort <- current_cohort
  # current_score$resids <- list(m1 = m1_pr,
  #                              m2 = m2_pr,
  #                              m3 = m3_pr,
  #                              ind = indep_pr)
  struct_scores <- struct_scores %>% bind_rows(current_score)
}

struct_plot <- struct_scores %>%
  rename(Pos_Neg = type) %>%
  pivot_longer(
    cols = "C-HP":"I-MMHP",
    names_to = "Method", values_to = "Score"
  )
# filter(Score > 0) %>%
# filter(Score < 0.99) %>%
# filter(!(Method == "I-MMHP" & Pos_Neg == "Positive"))
# mutate(Method = as.factor(Method, levels = c("I-MMHP","C-HP","C-DCHP",
#                                              "C-MMHP"))) %>%
struct_plot$Method <- factor(struct_plot$Method, levels = c(
  "I-MMHP", "C-HP", "C-DCHP",
  "C-MMHP"
))

# box_cols <- col_df %>%
#   filter(method %in% c("I-MMHP","C-HP","C-DCHP","C-MMHP")) %>%
#   pull(cols)

struct_plot %>%
  ggplot(aes(Method, Score, fill = Method)) +
  geom_boxplot() +
  # scale_fill_manual(values = box_cols) +
  ggtitle("(b)") +
  facet_wrap(~Pos_Neg) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )
```

# Figure 8

## Figure 8a: MAE of interaction predictions, single cohort

```{r fig-8a-frob-norm-cohort5, fig.width=8,fig.height=4,eval=TRUE}
load("../output/plot_N_predict.RData")
current_cohort <- 5 # cohort 1 looks worse than rest

temp_plot_df <- predict_day_mae_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c("I-MMHP", "DSNL", "C-HP", "C-DCHP", "C-MMHP")
)
temp_plot_df$day <- factor(temp_plot_df$day)
# png(paste(plot_path,"real_predict_N_one_cohort.png",sep=""),
# width=600,height=300)

plot_cols_pred <- col_df %>%
  filter(method %in% levels(temp_plot_df$method)) %>%
  arrange(factor(method, levels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP", "C-MMHP"
  ))) %>%
  pull(cols)

p3 <- temp_plot_df %>%
  rowwise() %>%
  # mutate(norm = min(norm,60)) %>%
  # rename(Method = method) %>%
  ggplot(aes(x = method, y = mae, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  xlab("Prediction horizon (day)") +
  ylab("MAE") +
  ylim(0, 8) + # aren't infinite but >150 and <200
  NULL
p3 + labs(fill = "Method", title = "Single Cohort")
```

## Figure 8b: Same plot aggregated for all cohorts

```{r fig-7a-pred-all-cohorts-box-mae}
### updated to revision
load("../output/plot_N_predict_revision.RData")

all_cohort_box <- all_cohort_mae_df

all_cohort_box$method <- factor(all_cohort_box$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP",
    "C-MMHP"
  )
)
all_cohort_box$day <- factor(all_cohort_box$day)

plot_cols_pred <- col_df %>%
  filter(method %in% levels(all_cohort_box$method)) %>%
  arrange(factor(method, levels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP", "C-MMHP"
  ))) %>%
  pull(cols)

all_cohort_box %>%
  ggplot(aes(x = method, y = norm, fill = method)) +
  # facet_grid(cols = vars(day)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(
      size = 22,
      hjust = 0.5, face = "bold"
    ),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  # guides(color=guide_legend(title="Method")) +
  xlab("Prediction horizon (day)") +
  ylab("MAE") +
  labs(fill = "Method") +
  # ylim(0, 80) +
  NULL
```

# Figure 9

## Figure 9a: Predicted Spearman rank for single cohort

```{r fig-9a-spearman-rank-cohort1, eval=FALSE}
load(here("output", "plot_rank_predict.RData"))
# current_cohort <- 5
temp_plot_df <- predict_day_rank_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c(
    "I-MMHP", "DSNL", "C-HP",
    "C-DCHP", "C-MMHP"
  )
)
temp_plot_df$day <- factor(temp_plot_df$day)
# png(paste(plot_path,"real_predict_rank_one_cohort.png",sep=""),
# width=600,height=300)
p5 <- ggplot(temp_plot_df, aes(x = method, y = spearman, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  # scale_fill_manual(values =
  # c("gray","yellowgreen","steelblue","goldenrod2","firebrick2")) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  xlab("Prediction horizon (day)") +
  ylab("Spearman rank correlation") +
  # ylim(0, 1) +
  NULL
p5 + labs(fill = "Method")
```

## Figure 9b: Predicted Spearman rank for all cohorts

```{r fig-7b-pred-rank-spearman-all-cohorts}
load("../output/plot_rank_predict.RData")

all_cohort_rank_df$method <- factor(all_cohort_rank_df$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c(
    "I-MMHP", "DSNL", "C-HP",
    "C-DCHP", "C-MMHP"
  )
)

p7 <- ggplot(
  all_cohort_rank_df,
  aes(x = method, y = spearman, fill = method)
) +
  # geom_line(size=1.2, alpha=0.8, position=position_dodge(0.2)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(b)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  xlab("Prediction horizon (day)") +
  ylab("Spearman rank corr.") +
  # coord_cartesian(ylim=c(0,1)) +
  labs(fill = "Method")

p7
```

# Figure 10

## Figure 10a

```{r construct data for plot, results='hide',message=FALSE,cache=TRUE}
library(PlayerRatings)

current_cohort <- 7
# 4 best so far, or 7

load(paste(data_path, cohort_names[current_cohort],
  "/predict_simulation_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]],
  cut_off = 1
)

df <- full_data[[cohort_names[current_cohort]]]
df <- df[df$Actor != "Start" & df$Actor != "End", ]
df1 <- expandrows(df)
df1 <- df1[order(df1$Timestamp), ] # ensure in date order
df1$event <- 1:nrow(df1)
glick.df <- df1[, c("event", "Actor", "Recipient", "score"), with = FALSE]
# need event, actor, recipient, score
gl <- my_glicko(glick.df, history = TRUE, cval = 2)

# ------------------- period
training_end <- nrow(df[df$day <= 18, ])
testing_end <- nrow(df) - 1 # 1077 #929 for cohort 5
newx <- c(training_end:testing_end)

# ----------------- set pars for plot
my_col <- c("black", "grey", "orange", "red")
col_vec <- colorRampPalette(my_col)(mice_number)
ltypes_vec <- rep(1, mice_number) # c(1,2,3,1,2,3,1,2,3,1,2,3)
history_width <- 3
sim_width <- 0.3
shade_vec <- makeTransparent(col_vec, 0.2)
less_shade_vec <- makeTransparent(col_vec, 0.7)

# ---------------- simulation mean and sd
sim_gl_array <- array(NA, dim = c(12, testing_end - training_end + 1, 1000))
for (s in 1:1000) {
  print(s)
  temp_gl <- glickoScoreSimulationWithWindowsReturnAll(
    predict.sim = m3_predict_sim[5:8, s], # 7 for c 2 and 10
    history.df = glick.df[1:training_end, ]
  )
  history_matrix <- temp_gl$history[, , 1]

  if (testing_end > ncol(temp_gl$history[, , 1])) {
    history_matrix <- cbind(
      history_matrix,
      t(matrix(rep(
        history_matrix[, ncol(history_matrix)],
        testing_end - ncol(temp_gl$history[, , 1])
      ),
      nrow = testing_end - ncol(temp_gl$history[, , 1]),
      byrow = TRUE
      ))
    )
  } else {
    # try(
    history_matrix <- history_matrix[, 1:testing_end] 
    # temp_gl$history[,1:testing_end]
    # )
  }
  sim_gl_array[, , s] <- history_matrix[, training_end:testing_end]
}
```

```{r fig-10a-glicko-rating}
par(mar = c(3, 5, 3, 1))
testing_end <- testing_end - 1
plot(1,
  type = "n", xlab = "",
  ylab = "Glicko ranking", xlim = c(0, dim(gl$history)[2]),
  ylim = c(min(gl$history[, , 1]), max(gl$history[, , 1])),
  cex.axis = 2, cex.lab = 2
)
for (m in c(1:mice_number)) {
  # history
  lines(gl$history[m, 1:training_end, 1],
    cex = 0.6, col = col_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )
  lines(c(rep(NA, training_end),
          gl$history[m, (training_end + 1):testing_end, 1]),
    cex = 0.6, col = shade_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )

  # simulation
  cur_end <- testing_end
  mean_vec <- apply(sim_gl_array[m, , ], 1, function(x) mean(na.omit(x)))
  sd_vec <- apply(sim_gl_array[m, , ], 1, function(x) sd(na.omit(x)))
  polygon(c(newx, rev(newx)),
          c(mean_vec - sd_vec, rev(mean_vec + sd_vec)),
          col = shade_vec[m], border = NA)
  # polygon(c(newx), c(mean_vec-sd_vec),col = shade_vec[m], border = NA)
  lines(c(rep(NA, training_end), mean_vec),
    cex = 0.6, col = col_vec[m], lwd = 1, lty = 2
  )
}

# annotate the days
for (d in c(19:21)) {
  abline(v = nrow(df[df$day <= (d - 1), ]), lty = 2, lwd = 0.4)
  text(x = nrow(df[df$day <= (d - 1), ]), y = 2600, d)
}
mtext(text = "(a)", side = 3, line = -2.3,
      outer = TRUE, cex = 2.8, at = 0.5, font = 2)
```

```{r fig-glicko-pred-slides,eval=slides_plots,include=slides_plots}
par(mar = c(3, 5, 3, 1))
plot(1,
  type = "n", xlab = "",
  ylab = "Glicko ranking", xlim = c(0, dim(gl$history)[2]),
  ylim = c(min(gl$history[, , 1]), max(gl$history[, , 1])),
  cex.axis = 2, cex.lab = 2
)
for (m in c(1:mice_number)) {
  # history
  lines(gl$history[m, 1:training_end, 1],
    cex = 0.6, col = col_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )
  lines(c(rep(NA, training_end), 
          gl$history[m, (training_end + 1):testing_end, 1]),
    cex = 0.6, col = shade_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )

  # simulation
  cur_end <- testing_end
  mean_vec <- apply(sim_gl_array[m, , ], 1, function(x) mean(na.omit(x)))
  sd_vec <- apply(sim_gl_array[m, , ], 1, function(x) sd(na.omit(x)))
  polygon(c(newx, rev(newx)), c(mean_vec - sd_vec, rev(mean_vec + sd_vec)),
    col = shade_vec[m], border = NA
  )
  # polygon(c(newx), c(mean_vec-sd_vec),col = shade_vec[m], border = NA)
  lines(c(rep(NA, training_end), mean_vec),
    cex = 0.6, col = col_vec[m], lwd = 1, lty = 2
  )
}

# annotate the days
for (d in c(19:21)) {
  abline(v = nrow(df[df$day <= (d - 1), ]), lty = 2, lwd = 0.4)
  text(x = nrow(df[df$day <= (d - 1), ]), y = 2600, d)
}
# mtext(text="(a)", side=3, line=-2.3, outer=TRUE, cex=2.8, at=0.5, font=2)
```

```{r remove some objects, include=FALSE}
rm(m3_predict_sim)
rm(m2_predict_sim)
rm(m1_predict_sim)
rm(mmhp_predict_sim)
rm(fit_agg_rank)
rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(fit_cohort_mmhp)

# also removing stan fits as they can cause issues in the next chunk
```

## Fig 10b

```{r process for state separation cohort1,eval=FALSE}
# only run this once and save the results
# this doesn't need to be rerun ever, even if the stan fits change
# current_cohort <- 1

load(paste(data_path, cohort_names[current_cohort],
  "/cmmhp_est_zt_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>%
  group_by(initiator, recipient) %>%
  dplyr::summarize(
    count = n(),
    observe = list(observe.id),
    observe.length = list(observe.time),
    no.events = list(no.events)
  )
unique_observe_win <- unique(return_df[,
                                       c("day", "observe.id", "observe.time")])

## 5.  find how many 1s for each pair
print(".......state separation plot........")
total_event_array <- array(0, dim = c(
  mice_number, mice_number,
  max(return_df$observe.id)
))
active_event_array <- array(0, dim = c(
  mice_number, mice_number,
  max(return_df$observe.id)
))

for (i in 1:mice_number) {
  for (j in c(1:mice_number)[-i]) {
    pair <- which(unique_pairs_df$initiator == i & 
                    unique_pairs_df$recipient == j)
    if (length(pair) > 0) {
      current_window_vec <- unique_pairs_df$observe[[pair]]
      for (cur_win in current_window_vec) {
        row_indicator <- return_df$initiator == i & 
          return_df$recipient == j & return_df$observe.id == cur_win
        total_event_array[i, j, cur_win] <- length(return_df[
          row_indicator, "event.times"][[1]])
        active_event_array[i, j, cur_win] <- 
          sum(apply(2 - state_array_list[[pair]][[cur_win]], 1, mean) > 0.5)
      }
    }
  }
}

utility_state_day <- array(0, dim = c(21, mice_number, mice_number))
social_state_day <- array(0, dim = c(21, mice_number, mice_number))
for (cur_day in c(1:21)) {
  cur_wins <- which(unique_observe_win$day == cur_day)
  utility_state_day[cur_day, , ] <- apply(
    active_event_array[, , cur_wins],
    c(1, 2), sum
  )
  social_state_day[cur_day, , ] <- apply(
    total_event_array[, , cur_wins],
    c(1, 2), sum
  ) -
    utility_state_day[cur_day, , ]
}

## utility dsnl
fit_dsnl_active <- stan("../lib/dsnl_poisson.stan",
  data = list(
    day = 21,
    Gt = utility_state_day,
    c = 1,
    sigma = 0.5
  ),
  iter = 1000, chains = 4, control = list(adapt_delta = 0.99)
)
print(fit_dsnl_active, pars = c("x"))
sim_dsnl_active <- rstan::extract(fit_dsnl_active)

## social dsnl
fit_dsnl_inactive <- stan("../lib/dsnl_poisson.stan",
  data = list(
    day = 21,
    Gt = social_state_day,
    c = 1,
    sigma = 1
  ),
  iter = 1000, chains = 4, control = list(adapt_delta = 0.99)
)
print(fit_dsnl_inactive, pars = c("x"))
sim_dsnl_inactive <- rstan::extract(fit_dsnl_inactive)

save(sim_dsnl_active, fit_dsnl_active, utility_state_day,
  sim_dsnl_inactive, fit_dsnl_inactive, social_state_day,
  file = paste(data_path, cohort_names[current_cohort],
    "/dsnl_state_separation_stan_result_",
    cohort_names[current_cohort],
    ".RData",
    sep = ""
  )
)
rm(interpolation_array_list)
rm(state_array_list)
rm(termination_state_list)
rm(initial_state_list)
```

```{r process data for plot 10b, cache=TRUE}
library(PlayerRatings)
# current_cohort <- 5
current_cohort <- 1
load(paste(data_path, cohort_names[current_cohort],
  "/dsnl_poisson_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
load(paste(data_path, cohort_names[current_cohort],
  "/dsnl_state_separation_stan_result_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
sub.types <- c("All", "Active", "Inactive")
# three methods
task.types <- paste0("day", 1:21)
data.tbl <- data.frame(array(NA, c(2000 * 3 * 21, 3)))
colnames(data.tbl) <- c("sub.type", "task.type", "data.value")
source("../lib/myGlicko.R")
#------------ gl score
df <- full_data[[cohort_names[current_cohort]]]
df <- df[df$Actor != "Start" & df$Actor != "End", ]
df1 <- expandrows(df)
df1 <- df1[order(df1$Timestamp), ] # ensure in date order
df1$event <- 1:nrow(df1)
glick.df <- df1[, c("event", "Actor", "Recipient", "score"),
                with = FALSE] # need event, actor, recipient, score
real.gl <- my_glicko(glick.df, history = TRUE, cval = 2)
clean_data_all <- cleanData(full_data[[cohort_names[current_cohort]]],
  cut_off = 1
)
day_end_idx <- unlist(lapply(c(1:21), function(x) {
  max(which(clean_data_all$day == x))
}))

## All aggregate
for (t in 1:length(task.types)) {
  temp_vec <- rep(NA, 2000)
  cur_real_rank <- real.gl$history[, day_end_idx[t], 1]
  for (s in c(1:2000)) {
    temp_result <- cor.test(sim_dsnl$x[s, t, ], cur_real_rank,
      method = "spearman"
    )
    temp_vec[s] <- temp_result$estimate
  }
  data.tbl$sub.type[(1:2000) + (t - 1) * 2000] <- sub.types[1]
  data.tbl$task.type[(1:2000) + (t - 1) * 2000] <- task.types[t]
  data.tbl$data.value[(1:2000) + (t - 1) * 2000] <- temp_vec
}

## Utility aggregate
for (t in 1:length(task.types)) {
  temp_vec <- rep(NA, 2000)
  cur_real_rank <- real.gl$history[, day_end_idx[t], 1]
  for (s in c(1:2000)) {
    temp_result <- cor.test(-sim_dsnl_active$x[s, t, ], cur_real_rank,
                            method = "spearman")
    temp_vec[s] <- temp_result$estimate
  }
  data.tbl$sub.type[(1:2000) + (t - 1) * 2000 + 42000] <- sub.types[2]
  data.tbl$task.type[(1:2000) + (t - 1) * 2000 + 42000] <- task.types[t]
  data.tbl$data.value[(1:2000) + (t - 1) * 2000 + 42000] <- temp_vec
}

## Social aggregate
for (t in 1:length(task.types)) {
  temp_vec <- rep(NA, 2000)
  cur_real_rank <- real.gl$history[, day_end_idx[t], 1]
  for (s in c(1:2000)) {
    temp_result <- cor.test(sim_dsnl_inactive$x[s, t, ],
      cur_real_rank,
      method = "spearman"
    )
    temp_vec[s] <- temp_result$estimate
  }
  data.tbl$sub.type[(1:2000) + (t - 1) * 2000 + 42000 * 2] <- sub.types[3]
  data.tbl$task.type[(1:2000) + (t - 1) * 2000 + 42000 * 2] <- task.types[t]
  data.tbl$data.value[(1:2000) + (t - 1) * 2000 + 42000 * 2] <- temp_vec
}

data.tbl$sub.type <- factor(data.tbl$sub.type)
data.tbl$task.type <- factor(data.tbl$task.type)
```

```{r fig-10b-rank-correlation}
par(mar = c(3, 5, 3, 1))
cx <- 1.75 # cex for axis label text;
# default is 1. bigger numbers make the text bigger, smaller, smaller.
x.ttl <- ""
# blank x-axis label
y.ttl <- "rank corr. (Glicko rank)"
ttl <- "(b)"
# y.lim <- c(min(data.tbl$data.value), max(data.tbl$data.value)+.5);
# y-axis limits. +2 at top to give room for legend
y.lim <- c(-1, 1.2)
x.lim <- c(0.5, (length(task.types)))
cols <- c("paleturquoise4", "lightgreen", "cornsilk")
shifts <- c(-0.25, 0, 0.25)

plot(
  x = 0, y = 0, xlim = x.lim, ylim = y.lim, xaxt = "n",
  col = "white", xlab = x.ttl,
  ylab = y.ttl, main = ttl,
  cex.main = cx * 1.3, cex.axis = cx, cex.lab = cx
)
axis(
  side = 1, at = 1:length(task.types), labels = task.types,
  cex.axis = cx, cex.lab = cx
) # put on the x-axis labels
grid(nx = NA, ny = NULL, col = "darkgrey")
lines(x = c(-1, 100), y = c(0, 0), col = "darkgrey")
for (t.id in 1:length(task.types)) {
  for (i in 1:length(sub.types)) { # t.id <- 1; i <- 1;
    inds <- which(data.tbl$sub.type == sub.types[i] &
      data.tbl$task.type == task.types[t.id])
    boxplot(data.tbl$data.value[inds],
      at = t.id + shifts[i],
      col = cols[i], add = TRUE,
      xaxt = "n", yaxt = "n", bty = "n", boxwex = 0.35,
      outpch = 20, outcex = 0.5
    )
  }
}
legend(
  x = "top", legend = sub.types, fill = cols,
  horiz = TRUE, cex = cx, bg = "white", bty = "n"
)
box()
```