### 2021-07-13 sim_design/sim_design.Rmd
add sim design rmd file

---
title: "Better Designed Simulations"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE
)

func_dir <- "../lib/"
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())

library(here)
library(cmdstanr)
library(ppdiag)
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(bayesplot)
library(posterior)
library(viridis)
library(here)
library(compete)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source(paste(func_dir, "circularPlot.R", sep = ""))

theme_set(theme_minimal())

col_df <- tibble(method = c("I&SI",
                            "AggRank",
                            "DSNL",
                            "Glicko",
                            "I-MMHP",
                            "C-HP",
                            "C-DCHP",
                            "True",
                            "C-MMHP"),
            cols = viridis(9))

model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 + exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)

sim_data_path <- here("output", "revisions", "sim_m3", "sim_3", "/")

```

In this document we want to better develop the properties of
the proposed latent ranking models for animal interactions, leading to
better designed and more robust simulation studies.

# Parameters of the Simulation Model

We specify the following parameters for our simulation model.

```{r specify-pars}
num_nodes <- 10 #20
cut_off <- 3
obs_time <- 100#50 #100

object_par <- list(sim_lambda_1 = 0.6,
                   gamma_var = seq(from = 0.01, to = 0.2,
                                   length.out = num_nodes),
                   zeta_var = rep(0.1, num_nodes),
                   sim_eta_1 = 1, # from 3.5
                   sim_eta_2 = 2,#1.5, # from 2.6
                   sim_eta_3 = 3, # this seems better
                   #sim_eta_3 = 7.5,
                   sim_beta = 1.5, # from 2
                   f_vec_1 = seq(from = 0.2, to = 0.9,
                                 length.out = num_nodes))

object_fn <- list(alpha.fun = function(x,y,eta1,eta2){
  return(eta1*x*y*exp(-eta2*abs(x-y)))},
                  q1.fun = function(x,y,eta3){return(exp(-eta3*x))},
                  q0.fun = function(x,y,eta3){return(exp(-eta3*y))})

```

# Fitting the Simulated Models

We fit each of the 3 proposed models to this simulated data.

# Evaluating the Fit of the Simulated Models

We can explore the fit of a given model by visualizing the 
posterior trace plots for each parameter first, to confirm our
Bayesian model correctly sampled from the posterior.

```{r load-specific-stan-fit}
current_cohort <- 1

print(current_cohort)
### currently 13,14,17 completed

load(paste0(sim_data_path, "sim_model3_fit123_", current_cohort, ".RData"))
load(paste(sim_data_path, "fit123_state_est_", current_cohort,
           ".RData", sep = ""))
clean_sim_data <- cleanSimulationData(
    raw_data = sim_model3_data,
    cut_off = cut_off,
    N = length(object_par$f_vec_1)
  )

```

## Model 1

```{r examine-m1-stan}
mcmc_trace(sim_model3_stan_sim1, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_trace(sim_model3_stan_sim1, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim1, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_dens(sim_model3_stan_sim1, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim1, pars = c("lp__", "lambda0", "beta"))

```

## Model 2

```{r examine-m2-stan}
mcmc_trace(sim_model3_stan_sim2, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_trace(sim_model3_stan_sim2, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim2, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_dens(sim_model3_stan_sim2, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim2, pars = c("lp__", "beta"))

```

## Model 3

```{r examine-m3-stan}
mcmc_trace(sim_model3_stan_sim3, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_trace(sim_model3_stan_sim3, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim3, pars = c("eta_1", "eta_2", "eta_3"))

mcmc_dens(sim_model3_stan_sim3, pars = c("f[1]",
                                          "f[5]",
                                          "f[10]"))

mcmc_dens(sim_model3_stan_sim3, pars = c("lp__", "beta"))

```

These traceplots indicate that M1 has identifiability issues, struggling
with multimodality for multiple parameters, which is not the case in
M2 and M3. Similarly, M3 seems to better recover the model parameters
than M2.

## Model Diagnostics

We can then evaluate the fit of each model further, by 
computing the compensators of the point process.

```{r init-diags}
n_sim <- 1

N_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))

```

```{r compute-diagnostics}
s <- 1
### transform from cmdstan here
f_draws <- sim_model3_stan_sim1 %>% 
select(starts_with("f"))
###
model1_par_est <- list(
  lambda0 = mean(sim_model3_stan_sim1$lambda0),
  eta_1 = mean(sim_model3_stan_sim1$eta_1),
  eta_2 = mean(sim_model3_stan_sim1$eta_2),
  eta_3 = mean(sim_model3_stan_sim1$eta_3),
  beta = mean(sim_model3_stan_sim1$beta),
  f = apply(f_draws, 2, mean)
)

for (i in c(1:num_nodes)) {
  for (j in c(1:num_nodes)[-i]) {
    current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
    m1_object <- list(
      lambda0 = model1_par_est$lambda0,
      alpha = model1_fn$alpha.fun(
        model1_par_est$f[i],
        model1_par_est$f[j],
        model1_par_est$eta_1,
        model1_par_est$eta_2,
        model1_par_est$eta_3
      ),
      beta = model1_par_est$beta
    )
    Lambda_model1_array[s, i, j] <-
      uniHawkesIntegralIntensity(
        object = m1_object,
        events = current_t,
        termination = tail(current_t, 1)
      )
    Lambda_test <- uniHawkesCompensator(
      object = m1_object,
      events = current_t
    )
    KS_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
    KS_p_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
    PR_model1_array[s, i, j] <- uniHawkesPearsonResidual(
      object = m1_object,
      events = current_t,
      termination =
        tail(current_t, 1)
    )
  }
}

## model2
#############
### convert to cmdstan
gamma_draws <- sim_model3_stan_sim2 %>% 
select(starts_with("gamma"))
zeta_draws <- sim_model3_stan_sim2 %>% 
select(starts_with("zeta"))
f_draws <- sim_model3_stan_sim2 %>% 
select(starts_with("f"))

model2_par_est <- list(
  gamma = apply(gamma_draws, 2, mean),
  zeta = apply(zeta_draws, 2, mean),
  eta_1 = mean(sim_model3_stan_sim2$eta_1),
  eta_2 = mean(sim_model3_stan_sim2$eta_2),
  eta_3 = mean(sim_model3_stan_sim2$eta_3),
  beta = mean(sim_model3_stan_sim2$beta),
  f = apply(f_draws, 2, mean)
)

for (i in c(1:num_nodes)) {
  for (j in c(1:num_nodes)[-i]) {
    current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
    m2_object <- list(
      lambda0 = model2_par_est$gamma[i] +
        model2_par_est$zeta[j],
      alpha = model1_fn$alpha.fun(
        model2_par_est$f[i],
        model2_par_est$f[j],
        model2_par_est$eta_1,
        model2_par_est$eta_2,
        model2_par_est$eta_3
      ),
      beta = model2_par_est$beta
    )
    Lambda_model2_array[s, i, j] <-
      uniHawkesIntegralIntensity(
        object = m1_object,
        events = current_t,
        termination = tail(current_t, 1)
      )
    Lambda_test <- uniHawkesCompensator(
      object = m2_object,
      events = current_t
    )
    KS_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
    KS_p_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
    PR_model2_array[s, i, j] <-
      uniHawkesPearsonResidual(
        object = m2_object,
        events = current_t,
        termination = tail(current_t, 1)
      )
  }
}

```

We can then briefly look at some of these fits.

```{r m1-m2-diags}
# hist(as.vector(-log(KS_p_model1_array)))
# hist(as.vector(-log(KS_p_model2_array)))
# 
# hist(as.vector(PR_model1_array))
# hist(as.vector(PR_model2_array))

exp_fcn <- function(x) exp(-x)

tibble(ks = as.vector(KS_p_model1_array)) %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density..)) +
  # scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value", title = "C-HP") +
  stat_function(fun = exp_fcn, colour = "red")

tibble(ks = as.vector(KS_p_model2_array)) %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density..)) +
  # scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value", title = "C-DCHP") +
  stat_function(fun = exp_fcn, colour = "red")

```

```{r}
## Model3 & true model
######
### cmdstan part
lam0_draws <- sim_model3_stan_sim3 %>% 
select(starts_with("lambda0"))
lam1_draws <- sim_model3_stan_sim3 %>% 
select(starts_with("lambda1"))
f_draws <- sim_model3_stan_sim3 %>% 
select(starts_with("f"))

lambda_0_est <- apply(lam0_draws, 2, mean)
lambda_1_est <- apply(lam1_draws, 2, mean)
## then put these into a matrix
lam0_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
lam1_par_est <- matrix(0,
  nrow = length(object_par$f_vec_1),
  ncol = length(object_par$f_vec_1)
)
for (i in seq_along(lambda_0_est)) {
  row_id <- clean_sim_data$I_fit[i]
  col_id <- clean_sim_data$J_fit[i]
  if(clean_sim_data$N_count[row_id, col_id ] >= cut_off) {
    lam0_par_est[row_id, col_id] <- lambda_0_est[i]
    lam1_par_est[row_id, col_id] <- lambda_1_est[i]
  }

}
mmhp_par_est <- list(
  lambda0 = lam0_par_est,
  lambda1 = lam1_par_est,
  eta_1 = mean(sim_model3_stan_sim3$eta_1),
  eta_2 = mean(sim_model3_stan_sim3$eta_2),
  eta_3 = mean(sim_model3_stan_sim3$eta_3),
  beta = mean(sim_model3_stan_sim3$beta),
  f = apply(f_draws, 2, mean)
)
clean_sim_data <- cleanSimulationData(
  raw_data = sim_model3_data,
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)

for (i in c(1:num_nodes)) {
  for (j in c(1:num_nodes)[-i]) {
    test.mmhp <- sim_model3_data$mmhp_matrix[i, j][[1]]
    temp.t <- test.mmhp$tau
    current.n <- length(temp.t) - 1
    time.segment <- seq(0, tail(temp.t, 1), length.out = 10000)

    object_hat <- list(
      lambda0 = mmhp_par_est$lambda0[i, j],
      lambda1 = mmhp_par_est$lambda1[i, j],
      alpha = model3_fn$alpha.fun(
        mmhp_par_est$f[i],
        mmhp_par_est$f[j],
        mmhp_par_est$eta_1,
        mmhp_par_est$eta_2
      ),
      beta = mmhp_par_est$beta,
      q1 = model3_fn$q1.fun(
        mmhp_par_est$f[i],
        mmhp_par_est$f[j],
        mmhp_par_est$eta_3
      ),
      q2 = model3_fn$q0.fun(
        mmhp_par_est$f[i],
        mmhp_par_est$f[j],
        mmhp_par_est$eta_3
      )
    )

    object_true <- lapply(object_matrix, function(x) x[i, j])
    names(object_true) <- c("lambda0", "lambda1",
                            "alpha", "beta", "q1", "q2")

    ## est Lambda and [R]
    if(length(temp.t) > 2){
      Lambda.test <- mmhpCompensator(
        params = object_hat,
        t = temp.t,
        pzt = 2 -
          event_state_est_lst[i, j][[1]]$zt_v,
        if.pzt = FALSE
      )
      KS_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$statistic
      KS_p_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$p.value
      state.est.latent.mmhp <- interpolate_state_est_lst[i, j][[1]]
      step.fun.est <- stepfun(
        state.est.latent.mmhp$x.hat,
        2 - state.est.latent.mmhp$z.hat
      )
      est.intensity <- mmhpIntensityNumeric(
        params = object_hat,
        t = temp.t[-1],
        time.vec = time.segment,
        latent.vec =
          step.fun.est(time.segment)
      )
      est.intensity.events <- mmhpIntensityAtEvents(
        params = object_hat,
        t = temp.t,
        latent_z =
          event_state_est_lst[i, j][[1]]$zt_v
      )
    Lambda_model3_array[s, i, j] <-
      sum(est.intensity) * (time.segment[2] - time.segment[1])

    }
    if(clean_sim_data$N_count[i,j] < cut_off) {
      PR_model3_array[s, i, j] <- 0
    }
    else {
     PR_model3_array[s, i, j] <- sum(1 / sqrt(est.intensity.events)) -
      sum(sqrt(est.intensity)) * (time.segment[2] - time.segment[1]) 
    }

    ## Under the true model
    if(length(temp.t) > 2) {
      Lambda.test.true <- mmhpCompensator(
        params = object_true,
        t = temp.t,
        pzt = 2 -
          test.mmhp$zt[2:length(test.mmhp$zt)],
        if.pzt = FALSE
      )
      KS_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$statistic
      KS_p_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$p.value
      true.intensity <-
        mmhpTrueIntensityNumeric(
          params = object_true,
          t = temp.t,
          latent = list(
            ## these are the truth
            x = fixStateTransition(test.mmhp)$x,
            z = fixStateTransition(test.mmhp)$z
          ),
          time.vec = time.segment
        )
      true.intensity.events <- mmhpIntensityAtEvents(
        params = object_true,
        t = temp.t,
        latent_z = test.mmhp$zt
      )
      Lambda_true_array[s, i, j] <- sum(true.intensity$lambda.t) *
        (time.segment[2] - time.segment[1])
    }

    if(clean_sim_data$N_count[i,j] < cut_off) {
      PR_true_array[s, i, j] <- 0
    }
    else{
      PR_true_array[s, i, j] <- sum(1 / sqrt(true.intensity.events)) -
      sum(sqrt(true.intensity$lambda.t)) * (time.segment[2] - time.segment[1])
    }
  }
}
```

```{r diag-m3}

tibble(ks = as.vector(KS_p_model3_array)) %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density..)) +
  # scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value", title = "C-MMHP") +
  stat_function(fun = exp_fcn, colour = "red")

tibble(ks = as.vector(KS_p_true_array)) %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density..)) +
  # scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value", title = "True Parameters") +
  stat_function(fun = exp_fcn, colour = "red")

```

```{r pr-histograms}
hist(as.vector(PR_model1_array))
hist(as.vector(PR_model2_array))
hist(as.vector(PR_model3_array))
hist(as.vector(PR_true_array))

summary(as.vector(PR_model1_array))
summary(as.vector(PR_model2_array))
summary(as.vector(PR_model3_array))
summary(as.vector(PR_true_array))
```

## Inferred against True Ranks

```{r sim-ranks-plot}
f_m1 <- sim_model3_stan_sim1 %>% select(starts_with("f"))
m1_ranks <- apply(f_m1, 2, mean)

f_m1_sims <- f_m1 %>% 
  rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
  pivot_longer(everything(),
               names_to = "node",
               values_to = "rank")
f_m1_sims$model <- "C-HP"

f_m2 <- sim_model3_stan_sim2 %>% select(starts_with("f"))
m2_ranks <- apply(f_m2, 2, mean)

f_m2_sims <- f_m2 %>% 
  rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
  pivot_longer(everything(),
               names_to = "node",
               values_to = "rank")

f_m2_sims$model <- "C-DCHP"

f_m3 <- sim_model3_stan_sim3 %>% select(starts_with("f"))
m3_ranks <- apply(f_m1, 2, mean)

f_m3_sims <- f_m3 %>% 
  rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
  pivot_longer(everything(),
               names_to = "node",
               values_to = "rank")

f_m3_sims$model <- "C-MMHP"

sim_ranks <- bind_rows(f_m1_sims, f_m2_sims, f_m3_sims)

sim_ranks$node <- factor(sim_ranks$node, levels = 1:num_nodes)
sim_ranks$model <- factor(sim_ranks$model, levels = c("C-HP", 
                                                      "C-DCHP",
                                                      "C-MMHP"))

true_ranks <- tibble(node = 1:num_nodes, true_rank = object_par$f_vec_1)

sim_ranks %>% 
  ggplot(aes(node, rank)) +
  geom_boxplot() +
  facet_wrap(~model) +
  geom_point(data = true_ranks, aes(node, true_rank, colour = "red"),
             show.legend = FALSE)

```

# Diagnostics for all sims

```{r all-diags}
# sim_data_path <- here("output", "revisions", )

n_sim <- 20

num_nodes <- 10 #20
cut_off <- 3
obs_time <- 100 #100

# object_par <- list(sim_lambda_1 = 0.4,
#                    gamma_var = seq(from = 0.01,
#                                     to = .2,
#                                    length.out = num_nodes),
#                    zeta_var = rep(0.1, num_nodes),
#                    sim_eta_1 = 1, # from 3.5
#                    sim_eta_2 = 2, # from 2.6
#                    sim_eta_3 = 3, # this seems better
#                    #sim_eta_3 = 7.5,
#                    sim_beta = 1.5, # from 2
#                    f_vec_1 = seq(from = 0.10,
#                                  to = 0.9,
#                                  length.out = num_nodes))

## Initialize
N_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
Lambda_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
PR_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model1_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model2_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_model3_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))
KS_p_true_array <- array(0, dim = c(
  n_sim, length(object_par$f_vec_1),
  length(object_par$f_vec_1)
))

```

```{r}
## Iteration on 50 simulations

sims <- 1:20
sims <- sims[c(-9,-13,-19,-20)]

for (s in sims) {
  print(s)
  load(paste0(sim_data_path, "sim_model3_fit123_", s, ".RData"))
  load(paste(sim_data_path, "fit123_state_est_", s, ".RData", sep = ""))
  clean_sim_data <- cleanSimulationData(
    raw_data = sim_model3_data,
    cut_off = cut_off,
    N = length(object_par$f_vec_1)
  )

  N_array[s, , ] <- clean_sim_data$N_count

  ## model 1
  #########
  ### transform from cmdstan here
  f_draws <- sim_model3_stan_sim1 %>% 
  select(starts_with("f"))

  ###
  model1_par_est <- list(
    lambda0 = mean(sim_model3_stan_sim1$lambda0),
    eta_1 = mean(sim_model3_stan_sim1$eta_1),
    eta_2 = mean(sim_model3_stan_sim1$eta_2),
    eta_3 = mean(sim_model3_stan_sim1$eta_3),
    beta = mean(sim_model3_stan_sim1$beta),
    f = apply(f_draws, 2, mean)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
      m1_object <- list(
        lambda0 = model1_par_est$lambda0,
        alpha = model1_fn$alpha.fun(
          model1_par_est$f[i],
          model1_par_est$f[j],
          model1_par_est$eta_1,
          model1_par_est$eta_2,
          model1_par_est$eta_3
        ),
        beta = model1_par_est$beta
      )
      Lambda_model1_array[s, i, j] <-
        uniHawkesIntegralIntensity(
          object = m1_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
      Lambda_test <- uniHawkesCompensator(
        object = m1_object,
        events = current_t
      )
      KS_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
      KS_p_model1_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
      PR_model1_array[s, i, j] <- uniHawkesPearsonResidual(
        object = m1_object,
        events = current_t,
        termination =
          tail(current_t, 1)
      )
    }
  }

  ## model2
  #############
  ### convert to cmdstan
  gamma_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("gamma"))
  zeta_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("zeta"))
  f_draws <- sim_model3_stan_sim2 %>% 
  select(starts_with("f"))

  model2_par_est <- list(
    gamma = apply(gamma_draws, 2, mean),
    zeta = apply(zeta_draws, 2, mean),
    eta_1 = mean(sim_model3_stan_sim2$eta_1),
    eta_2 = mean(sim_model3_stan_sim2$eta_2),
    eta_3 = mean(sim_model3_stan_sim2$eta_3),
    beta = mean(sim_model3_stan_sim2$beta),
    f = apply(f_draws, 2, mean)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      current_t <- sim_model3_data$mmhp_matrix[i, j][[1]]$tau
      m2_object <- list(
        lambda0 = model2_par_est$gamma[i] +
          model2_par_est$zeta[j],
        alpha = model1_fn$alpha.fun(
          model2_par_est$f[i],
          model2_par_est$f[j],
          model2_par_est$eta_1,
          model2_par_est$eta_2,
          model2_par_est$eta_3
        ),
        beta = model2_par_est$beta
      )
      Lambda_model2_array[s, i, j] <-
        uniHawkesIntegralIntensity(
          object = m1_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
      Lambda_test <- uniHawkesCompensator(
        object = m2_object,
        events = current_t
      )
      KS_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$statistic
      KS_p_model2_array[s, i, j] <- ks.test(Lambda_test, "pexp")$p.value
      PR_model2_array[s, i, j] <-
        uniHawkesPearsonResidual(
          object = m2_object,
          events = current_t,
          termination = tail(current_t, 1)
        )
    }
  }

  ## Model3 & true model
  ######
  ### cmdstan part
  lam0_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("lambda0"))
  lam1_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("lambda1"))
  f_draws <- sim_model3_stan_sim3 %>% 
  select(starts_with("f"))

  if(ncol(lam0_draws) != sum(clean_sim_data$N_count >2)) {
    stop("Mismatch between params and event pairs")
  }

  lambda_0_est <- apply(lam0_draws, 2, mean)
  lambda_1_est <- apply(lam1_draws, 2, mean)
  ## then put these into a matrix
  lam0_par_est <- matrix(0,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  )
  lam1_par_est <- matrix(0,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  )
  for (i in seq_along(lambda_0_est)) {
    row_id <- clean_sim_data$I_fit[i]
    col_id <- clean_sim_data$J_fit[i]
    lam0_par_est[row_id, col_id] <- lambda_0_est[i]
    lam1_par_est[row_id, col_id] <- lambda_1_est[i]
  }
  mmhp_par_est <- list(
    lambda0 = lam0_par_est,
    lambda1 = lam1_par_est,
    eta_1 = mean(sim_model3_stan_sim3$eta_1),
    eta_2 = mean(sim_model3_stan_sim3$eta_2),
    eta_3 = mean(sim_model3_stan_sim3$eta_3),
    beta = mean(sim_model3_stan_sim3$beta),
    f = apply(f_draws, 2, mean)
  )
  clean_sim_data <- cleanSimulationData(
    raw_data = sim_model3_data,
    cut_off = cut_off,
    N = length(object_par$f_vec_1)
  )

  for (i in c(1:num_nodes)) {
    for (j in c(1:num_nodes)[-i]) {
      test.mmhp <- sim_model3_data$mmhp_matrix[i, j][[1]]
      temp.t <- test.mmhp$tau
      current.n <- length(temp.t) - 1
      time.segment <- seq(0, tail(temp.t, 1), length.out = 10000)

      object_hat <- list(
        lambda0 = mmhp_par_est$lambda0[i, j],
        lambda1 = mmhp_par_est$lambda1[i, j],
        alpha = model3_fn$alpha.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_1,
          mmhp_par_est$eta_2
        ),
        beta = mmhp_par_est$beta,
        q1 = model3_fn$q1.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_3
        ),
        q2 = model3_fn$q0.fun(
          mmhp_par_est$f[i],
          mmhp_par_est$f[j],
          mmhp_par_est$eta_3
        )
      )

      object_true <- lapply(object_matrix, function(x) x[i, j])
      names(object_true) <- c("lambda0", "lambda1",
                              "alpha", "beta", "q1", "q2")

      ## est Lmabda and [R]
      if(length(temp.t) > 2 ) {
        Lambda.test <- mmhpCompensator(
        params = object_hat,
        t = temp.t,
        pzt = 2 -
          event_state_est_lst[i, j][[1]]$zt_v,
        if.pzt = FALSE
        )
        KS_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$statistic
        KS_p_model3_array[s, i, j] <- ks.test(Lambda.test, "pexp")$p.value
        state.est.latent.mmhp <- interpolate_state_est_lst[i, j][[1]]
        step.fun.est <- stepfun(
          state.est.latent.mmhp$x.hat,
          2 - state.est.latent.mmhp$z.hat
        )
        est.intensity <- mmhpIntensityNumeric(
          params = object_hat,
          t = temp.t[-1],
          time.vec = time.segment,
          latent.vec =
            step.fun.est(time.segment)
        )
        est.intensity.events <- mmhpIntensityAtEvents(
          params = object_hat,
          t = temp.t,
          latent_z =
            event_state_est_lst[i, j][[1]]$zt_v
        )
        Lambda_model3_array[s, i, j] <-
          sum(est.intensity) * (time.segment[2] - time.segment[1])
        PR_model3_array[s, i, j] <- sum(1 / sqrt(est.intensity.events)) -
          sum(sqrt(est.intensity)) * (time.segment[2] - time.segment[1])
        }
      ## Under the true model
      if(length(temp.t) > 2 ) {
        Lambda.test.true <- mmhpCompensator(
        params = object_true,
        t = temp.t,
        pzt = 2 -
          test.mmhp$zt[2:length(test.mmhp$zt)],
        if.pzt = FALSE
        )
        KS_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$statistic
        KS_p_true_array[s, i, j] <- ks.test(Lambda.test.true, "pexp")$p.value
        true.intensity <-
          mmhpTrueIntensityNumeric(
            params = object_true,
            t = temp.t,
            latent = list(
              ## these are the truth
              x = fixStateTransition(test.mmhp)$x,
              z = fixStateTransition(test.mmhp)$z
            ),
            time.vec = time.segment
          )
        true.intensity.events <- mmhpIntensityAtEvents(
          params = object_true,
          t = temp.t,
          latent_z = test.mmhp$zt
        )
        Lambda_true_array[s, i, j] <- sum(true.intensity$lambda.t) *
          (time.segment[2] - time.segment[1])
        PR_true_array[s, i, j] <- sum(1 / sqrt(true.intensity.events)) -
          sum(sqrt(true.intensity$lambda.t)) * 
          (time.segment[2] - time.segment[1])
      }

    }
  }
  rm(sim_model3_data)
}
```

```{r plot-diag-output}
cols_plot <- col_df %>% 
  filter(method %in% c("C-HP","C-DCHP","C-MMHP", "True")) %>%
  arrange(factor(method, levels = c("C-HP", "C-DCHP", "C-MMHP", "True"))) %>%
  pull(cols)

m1_ks_p <- tibble(Method = "C-HP",
                  ks = as.vector(KS_p_model1_array[sims, ,]))

m2_ks_p <- tibble(Method = "C-DCHP",
                  ks = as.vector(KS_p_model2_array[sims, ,]))

m3_ks_p <- tibble(Method = "C-MMHP",
                  ks = as.vector(KS_p_model3_array[sims, ,]))

true_ks_p <- tibble(Method = "True",
                    ks = as.vector(KS_p_true_array[sims, ,]))

all_ks_p <- m1_ks_p %>% 
  bind_rows(m2_ks_p) %>% 
  bind_rows(m3_ks_p) %>% 
  bind_rows(true_ks_p)

all_ks_p$Method <- factor(all_ks_p$Method, levels = c("C-HP",
                                                      "C-DCHP",
                                                      "C-MMHP",
                                                      "True") )

exp_fcn <- function(x) exp(-x)

all_ks_p %>% 
  filter(ks != 0) %>% 
  mutate(log_p = -log(ks)) %>% 
  ggplot(aes(x = log_p)) + 
  # geom_freqpoly(binwidth = 20) +
  geom_histogram(aes(y = ..density.., fill = Method), bins = 50) +
  scale_fill_manual(values = cols_plot) +
  labs(y = "Negative Log KS Statistic P Value") +
  stat_function(fun = exp_fcn, colour = "red") +
  facet_wrap(~Method)
```

```{r pr-summary}
summary(as.vector(PR_model1_array[sims, ,]))
summary(as.vector(PR_model2_array[sims, ,]))
summary(as.vector(PR_model3_array[sims, ,][is.finite(PR_model3_array[sims,
                                                                     ,])]))
summary(as.vector(PR_true_array[sims, ,]))
```

### Ranks from all Sims

```{r}
# sim_data_path <- here("output", "revisions", "sim_m3", "sim_2", "/")

num_nodes <- 10

object_par$f_vec_1

sim_ranks_mean <- tibble()
sim_ranks <- tibble()

for(curr_sim in sims) {
  load(paste0(sim_data_path,
              "sim_model3_fit123_",
              curr_sim,
              ".RData"))

  f_m1 <- sim_model3_stan_sim1 %>% select(starts_with("f"))
  m1_ranks <- apply(f_m1, 2, mean)

  f_m1_sims <- f_m1 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")
  f_m1_sims$model <- "C-HP"

  f_m2 <- sim_model3_stan_sim2 %>% select(starts_with("f"))
  m2_ranks <- apply(f_m2, 2, mean)

  f_m2_sims <- f_m2 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m2_sims$model <- "C-DCHP"

  f_m3 <- sim_model3_stan_sim3 %>% select(starts_with("f"))
  m3_ranks <- apply(f_m3, 2, mean)

  f_m3_sims <- f_m3 %>% 
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>% 
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>% 
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m3_sims$model <- "C-MMHP"

  curr_mean_ranks <- tibble(model = c(rep("C-HP", num_nodes),
                                 rep("C-DCHP", num_nodes),
                                 rep("C-MMHP", num_nodes)),
                       ranks = c(m1_ranks,
                                 m2_ranks,
                                 m3_ranks),
                       node = rep(1:num_nodes, 3),
                       true_rank = rep(object_par$f_vec_1, 3),
                       sim_id = curr_sim)
  sim_ranks_mean <- sim_ranks_mean %>% 
    bind_rows(curr_mean_ranks)

  sim_ranks <- sim_ranks %>% 
    bind_rows(f_m1_sims, f_m2_sims, f_m3_sims)

}

sim_ranks$node <- factor(sim_ranks$node, levels = 1:num_nodes)
sim_ranks_mean$node <- factor(sim_ranks_mean$node, levels = 1:num_nodes)
sim_ranks_mean$model <- factor(sim_ranks_mean$model, 
                               levels = c("C-HP", "C-DCHP", "C-MMHP"))
# sim_ranks$node

sim_ranks_mean %>% 
  # filter(model == "C-MMHP") %>% 
  ggplot(aes(node, ranks)) + 
  geom_boxplot() +
  facet_wrap(~model) +
  geom_point(aes(node, true_rank, colour = "red"),
             show.legend = FALSE) +
  labs(title = "Mean rank from each simulation")

## this is just the mean from each simulation, not the posterior draws

true_ranks <- tibble(node = 1:num_nodes, true_rank = object_par$f_vec_1)
true_ranks$node <- factor(true_ranks$node, levels = 1:num_nodes)

sim_ranks %>% 
  ggplot(aes(node, rank)) +
  geom_boxplot() +
  facet_wrap(~model) +
  geom_point(data = true_ranks, aes(node, true_rank, colour = "red"),
             show.legend = FALSE) +
  labs(title = "All Posterior Draws")

```

```{r other-code, include=FALSE}
col_df <- tibble(method = c("I&SI",
                            "AggRank",
                            "DSNL",
                            "Glicko",
                            "I-MMHP",
                            "C-HP",
                            "C-DCHP",
                            "True",
                            "C-MMHP"),
                 cols = viridis(9))

```