### 2021-08-10 output_scripts/investigate_rankings.Rmd
formatting and tidying up results a bit

---
title: "Ranking Stability"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output:
  html_document:
    # toc: true
    # toc_depth: 2
    # toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 5, fig.width = 7,
  warning = FALSE, cache = TRUE
)

slides_plots <- FALSE
extra_plots <- FALSE

revision_plots <- TRUE
original_plots <- FALSE
### original plots for plots which are no longer needed, 
### revision plots are what required

current_cohort <- 5

# don't write to this path
data_path <- "../output/revisions/lapl_check/prior_check/"

func_dir <- "../lib/"

library(here)
# library(rstan)
library(cmdstanr)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(posterior)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

cohort_names <- paste("cohort", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                      sep = "")
cohort_short_names <- paste("C", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                            sep = "")
cut_off <- 3
mice_number <- 12
col_df <- tibble(
  method = c(
    "I&SI", "AggRank", "DSNL", "Glicko", "I-MMHP",
    "C-HP", "C-DCHP", "True", "C-MMHP"
  ),
  cols = viridis(9)
)
theme_set(theme_minimal())
```

```{r load rankings, echo=FALSE, include=FALSE, results='hide', cache=TRUE}
full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for (curr_cohort in fit_cohorts) {
  naive_rank_10[[curr_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})
```

```{r define_functions, eval=revision_plots, include=FALSE}
model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 +
    exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)
```

# Interevents vs rankings

```{r interevents, message=FALSE}
for(current_cohort in 1:10) {
  cat(paste("Cohort ", current_cohort, "\n", sep = ""))
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  stan_input_lst <- prepareDataStan(current_cohort)

  pair_events <- apply(stan_input_lst$Nm, 1, sum)

  pair_matrix <- matrix(0, nrow = 12, ncol = 12)
  for(i in seq_along(pair_events)) {
    i_ind <- stan_input_lst$I_fit[i]
    j_ind <- stan_input_lst$J_fit[i]
    pair_matrix[i_ind, j_ind] <- pair_events[i]
  }
  # pair_matrix <- clean_data$N_count[expert_rank_10[[current_cohort]],
  #                                   expert_rank_10[[current_cohort]]]
  pair_matrix <- pair_matrix[expert_rank_10[[current_cohort]],
                    expert_rank_10[[current_cohort]]]
  colnames(pair_matrix) <- expert_rank_10[[current_cohort]]
  rownames(pair_matrix) <- expert_rank_10[[current_cohort]]
  print(pair_matrix)
  cat("=====\n")

}
```

# Ranking and Events

We first modify the priors from the C-MMHP model which leads to
much improved MCMC performance.

We want to look at the relationship between the inferred variability
of the ranking and the number of interactions a given animal is involved in.

We plot the inferred ranks arranged by decreasing number
of total interactions for all 4 of the comparison models, as in Fig6a.

```{r plot-rank-total-interactions, eval=FALSE}
 for(current_cohort in 1:10) {
   clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
   load(paste(data_path,
              cohort_names[current_cohort],
              "/cohort_mmhp_stan_result_",
              cohort_names[current_cohort],
              ".RData",
              sep = ""
              ))
   total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count)
   f_ranks <- paste("f[", order(total_inter, decreasing = TRUE),
                    "]", sep = "")
   f_draws <- sim_cohort_mmhp %>% select(starts_with("f")) %>%
   pivot_longer(everything())
   f_draws$name <- factor(f_draws$name, levels = f_ranks)
   plot_f <- f_draws %>%
   ggplot(aes(name, value)) +
   geom_boxplot() +
   labs(title = paste("Cohort ", current_cohort, sep = ""),
        subtitle = "Ordered by total interactions")
   print(plot_f)
 }
```

<!-- ```{r all-4-models, warning=FALSE} -->
<!-- for(current_cohort in 1:10) { -->
<!--   clean_data <- cleanData(full_data[[cohort_names[current_cohort]]]) -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/agg_rank_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/cohort_hp_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/cohort_dchp_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count) -->

<!--   no_method <- 4 -->
<!--   plot_latent_df <- data.frame( -->
<!--     Methods = rep(1:no_method, each = 1000 * mice_number), -->
<!--     Mice = rep(rep(1:mice_number, each = 1000), no_method), -->
<!--     force = rep(0, 1000 * mice_number * no_method) -->
<!--   ) -->
<!--   ## agg -->
<!--   x_draws <- sim_agg_rank %>% select(starts_with("x")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number))] <- -->
<!--     as.vector(as.matrix(x_draws)) -->

<!--   ## M1 -->
<!--   f_draws <- sim_cohort_hp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          1000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   ## M2 -->
<!--   f_draws <- sim_cohort_dchp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          2000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   ## M3 -->
<!--   f_draws <- sim_cohort_mmhp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          3000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   plot_latent_df$Mice <- factor(plot_latent_df$Mice, -->
<!--     levels = order(total_inter, decreasing = TRUE) -->
<!--   ) -->
<!--   plot_latent_df$Method <- factor(plot_latent_df$Method, -->
<!--     labels = c("AggRanking", "C-HP", "C-DCHP", "C-MMHP") -->
<!--   ) -->
<!--   plot_cols <- col_df %>% -->
<!--   filter(method %in% c( -->
<!--     "AggRank", "C-HP", -->
<!--     "C-DCHP", "C-MMHP" -->
<!--   )) %>% -->
<!--   pull(cols) -->

<!--   p1 <- ggplot(plot_latent_df, aes(x = Mice, y = force, fill = Method)) + -->
<!--     geom_boxplot(alpha = 0.7) + -->
<!--     scale_fill_manual(values = plot_cols) + -->
<!--     ggtitle("(a)") + -->
<!--     theme_bw() + -->
<!--     theme( -->
<!--       plot.title = element_text(size = 22, hjust = 0.5, face = "bold"), -->
<!--       text = element_text(size = 20), -->
<!--       axis.text.x = element_text(size = 10), -->
<!--       legend.position = "none" -->
<!--     ) + -->
<!--     facet_grid(cols = vars(Method)) + -->
<!--     xlab("Mice (ranked by total interactions)") + -->
<!--     ylab("Latent rank") + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = "")) -->
<!--   print(p1) -->
<!-- } -->

<!-- ``` -->

<!-- Similarly, here are the corresponding plots with the animals ranked by -->
<!-- I&SI score instead. -->

<!-- ```{r isi-ranks, warning=FALSE, eval=FALSE} -->
<!-- for(current_cohort in 1:10) { -->
<!--   clean_data <- cleanData(full_data[[cohort_names[current_cohort]]]) -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count) -->
<!--   f_ranks <- paste("f[", expert_rank_10[[current_cohort]], -->
<!--                    "]", sep = "") -->
<!--   f_draws <- sim_cohort_mmhp %>% select(starts_with("f")) %>% -->
<!--   pivot_longer(everything()) -->
<!--   f_draws$name <- factor(f_draws$name, levels = f_ranks) -->
<!--   plot_f <- f_draws %>% -->
<!--   ggplot(aes(name, value)) + -->
<!--   geom_boxplot() + -->
<!--   labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--        subtitle = "Ordered by I&SI score") -->
<!--   print(plot_f) -->
<!-- } -->

<!-- ``` -->

<!-- ```{r ranked-by-isi} -->
<!-- for(current_cohort in 1:10) { -->
<!--   clean_data <- cleanData(full_data[[cohort_names[current_cohort]]]) -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/agg_rank_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/cohort_hp_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   load(paste(data_path, cohort_names[current_cohort], -->
<!--     "/cohort_dchp_stan_result_", cohort_names[current_cohort], -->
<!--     ".RData", -->
<!--     sep = "" -->
<!--   )) -->
<!--   total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count) -->

<!--   no_method <- 4 -->
<!--   plot_latent_df <- data.frame( -->
<!--     Methods = rep(1:no_method, each = 1000 * mice_number), -->
<!--     Mice = rep(rep(1:mice_number, each = 1000), no_method), -->
<!--     force = rep(0, 1000 * mice_number * no_method) -->
<!--   ) -->
<!--   ## agg -->
<!--   x_draws <- sim_agg_rank %>% select(starts_with("x")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number))] <- -->
<!--     as.vector(as.matrix(x_draws)) -->

<!--   ## M1 -->
<!--   f_draws <- sim_cohort_hp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          1000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   ## M2 -->
<!--   f_draws <- sim_cohort_dchp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          2000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   ## M3 -->
<!--   f_draws <- sim_cohort_mmhp %>% select(starts_with("f")) -->
<!--   plot_latent_df$force[(1:(1000 * mice_number)) + -->
<!--                          3000 * mice_number] <- -->
<!--     as.vector(as.matrix(f_draws)) -->

<!--   plot_latent_df$Mice <- factor(plot_latent_df$Mice, -->
<!--     levels = expert_rank_10[[current_cohort]] -->
<!--   ) -->
<!--   plot_latent_df$Method <- factor(plot_latent_df$Method, -->
<!--     labels = c("AggRanking", "C-HP", "C-DCHP", "C-MMHP") -->
<!--   ) -->
<!--   plot_cols <- col_df %>% -->
<!--   filter(method %in% c( -->
<!--     "AggRank", "C-HP", -->
<!--     "C-DCHP", "C-MMHP" -->
<!--   )) %>% -->
<!--   pull(cols) -->

<!--   p1 <- ggplot(plot_latent_df, aes(x = Mice, y = force, fill = Method)) + -->
<!--     geom_boxplot(alpha = 0.7) + -->
<!--     scale_fill_manual(values = plot_cols) + -->
<!--     ggtitle("(a)") + -->
<!--     theme_bw() + -->
<!--     theme( -->
<!--       plot.title = element_text(size = 22, hjust = 0.5, face = "bold"), -->
<!--       text = element_text(size = 20), -->
<!--       axis.text.x = element_text(size = 10), -->
<!--       legend.position = "none" -->
<!--     ) + -->
<!--     facet_grid(cols = vars(Method)) + -->
<!--     xlab("Mice (ranked by I&SI)") + -->
<!--     ylab("Latent rank") + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = "")) -->
<!--   print(p1) -->
<!-- } -->

<!-- ``` -->

<!-- Cohort 3 win loss -->

<!-- ```{r c3-win-loss} -->
<!-- current_cohort <- 3 -->
<!-- clean_data <- cleanData(full_data[[cohort_names[current_cohort]]]) -->

<!-- clean_data$N_count -->

<!-- load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->

<!-- gamma_draws <- sim_cohort_mmhp %>% -->
<!--   select(starts_with("gamma")) %>% -->
<!--   pivot_longer(everything()) -->

<!-- gamma_draws$name <- factor(gamma_draws$name, -->
<!--                            levels = paste("gamma[", 1:12, "]", sep = "")) -->

<!-- gamma_draws %>% -->
<!--   ggplot(aes(name, value)) + -->
<!--   geom_boxplot() -->

<!-- zeta_draws <- sim_cohort_mmhp %>% -->
<!--   select(starts_with("zeta")) %>% -->
<!--   pivot_longer(everything()) -->

<!-- zeta_draws$name <- factor(zeta_draws$name, -->
<!--                            levels = paste("zeta[", 1:12, "]", sep = "")) -->

<!-- zeta_draws %>% -->
<!--   ggplot(aes(name, value)) + -->
<!--   geom_boxplot() -->

<!-- f_draws <- sim_cohort_mmhp %>% -->
<!--   select(starts_with("f")) %>% -->
<!--   pivot_longer(everything()) %>% -->
<!--   group_by(name) %>% -->
<!--   summarise(med = median(value)) %>% -->
<!--   arrange(-med) -->

<!-- f_draws %>% pull(name) -->

<!-- cmmhp_ranks <- c(10, 12, 7, 6, 3, 8, 2, 4, 1, 9, 11, 5) -->
<!-- clean_data$N_count[cmmhp_ranks, cmmhp_ranks] -->

<!-- c3_count <- clean_data$N_count -->
<!-- names(c3_count) <- 1:12 -->
<!-- rownames(c3_count) <- 1:12 -->
<!-- colnames(c3_count) <- 1:12 -->
<!-- isi98(c3_count) -->
<!-- isi98(c3_count[cmmhp_ranks, cmmhp_ranks]) -->
<!-- ``` -->

<!-- ## Plot Omega Lambda and f draws -->

<!-- ```{r plot-omega-lambda, message=FALSE} -->
<!-- for(current_cohort in 1:10) { -->
<!--   clean_data <- cleanData(full_data[[cohort_names[current_cohort]]]) -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   fit_draws <- as_draws_array(sim_cohort_mmhp) -->
<!--   plot1 <- mcmc_hist(fit_draws, pars = "w_lambda") + -->
<!--     labs(paste("Cohort ", current_cohort, sep = "")) -->
<!--   print(plot1) -->
<!-- } -->
<!-- ``` -->

<!-- # Compare rankings with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3, message=FALSE} -->
<!-- # data_path -->
<!-- other_data_path <- "../output/revisions/lapl_check/ident_check/" -->

<!-- for(current_cohort in 1:10) { -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   sim1 <- sim_cohort_mmhp -->
<!--   load(paste(other_data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   sim2 <- sim_cohort_mmhp -->

<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("f")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta3") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("f")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->

<!--   f_ranks <- paste("f[", expert_rank_10[[current_cohort]], -->
<!--                    "]", sep = "")  -->
<!--   all_data$name <- factor(all_data$name, levels = f_ranks) -->

<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Removing eta3, ordered by ISI", -->
<!--          y = "Latent Rankings") -->
<!--   print(comp_plot) -->

<!--   all_data %>% group_by(name, model) %>%  -->
<!--     summarise(model_mean = mean(value),  -->
<!--               model_sd = sd(value)) -->

<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("eta") | starts_with("w_lam") | starts_with("beta")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>%  -->
<!--                 select(starts_with("eta") |  -->
<!--                          starts_with("w_lam") | -->
<!--                          starts_with("beta")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta3", y = "Eta params") + -->
<!--     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -->
<!--   print(comp_plot) -->
<!--   ### -->
<!--   ## same for zeta, -->
<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("zeta") ) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("zeta")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta3", y = "Zeta params") + -->
<!--     theme(axis.text.x = element_blank()) + -->
<!--     facet_wrap(~name, scales = "free") -->
<!--   print(comp_plot) -->

<!--   ### -->
<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("gamma")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta3") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("gamma")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta3", y = "Out Degree") + -->
<!--     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -->
<!--   print(comp_plot) -->

<!-- } -->

<!-- ``` -->

<!-- # Compare rankings with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1, message=FALSE} -->
<!-- # data_path -->
<!-- other_data_path <- "../output/revisions/lapl_check/ident_check_no_eta1/" -->

<!-- for(current_cohort in 1:10) { -->
<!--   load(paste(data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   sim1 <- sim_cohort_mmhp -->
<!--   load(paste(other_data_path, -->
<!--              cohort_names[current_cohort], -->
<!--              "/cohort_mmhp_stan_result_", -->
<!--              cohort_names[current_cohort], -->
<!--              ".RData", -->
<!--              sep = "" -->
<!--              )) -->
<!--   sim2 <- sim_cohort_mmhp -->

<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("f")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("f")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->

<!--   f_ranks <- paste("f[", expert_rank_10[[current_cohort]], -->
<!--                    "]", sep = "")  -->
<!--   all_data$name <- factor(all_data$name, levels = f_ranks) -->

<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta1, ranked by ISI", -->
<!--          y = "Latent Ranks") -->
<!--   print(comp_plot) -->

<!--   all_data %>% group_by(name, model) %>%  -->
<!--     summarise(model_mean = mean(value),  -->
<!--               model_sd = sd(value)) -->

<!--   ## same for beta_etc, -->
<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("eta") | starts_with("w_lam") | starts_with("beta")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>%  -->
<!--                 select(starts_with("eta") |  -->
<!--                          starts_with("w_lam") | -->
<!--                          starts_with("beta")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta1", y = "Eta params") + -->
<!--     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -->
<!--   print(comp_plot) -->
<!--   ### -->
<!--   ## same for zeta, -->
<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("zeta") ) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("zeta")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta1", y = "Zeta params") + -->
<!--     theme(axis.text.x = element_blank()) + -->
<!--     facet_wrap(~name, scales = "free") -->
<!--   print(comp_plot) -->

<!--   ### -->
<!--   all_data <- sim1 %>% -->
<!--     select(starts_with("gamma")) %>%  -->
<!--     pivot_longer(everything()) %>%  -->
<!--     mutate(model = "with eta1") %>%  -->
<!--     bind_rows(sim2 %>% select(starts_with("gamma")) %>%  -->
<!--                 pivot_longer(everything()) %>%  -->
<!--                 mutate(model = "without")) -->
<!--   all_data$model <- factor(all_data$model) -->
<!--   comp_plot <- ggplot(all_data, aes(name, value, fill = model)) + -->
<!--     geom_boxplot() + -->
<!--     labs(title = paste("Cohort ", current_cohort, sep = ""), -->
<!--          subtitle = "Comparison removing eta1", y = "Out Degree") + -->
<!--     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -->
<!--   print(comp_plot) -->

<!-- } -->

<!-- ``` -->

<!-- # Other Paper plots -->

<!-- ```{r fig-3b} -->
<!-- current_cohort <- 5 -->

<!-- m1_pr <- readRDS(paste(data_path, cohort_names[current_cohort], -->
<!--   "/chp_pr_matrix_", cohort_names[current_cohort], -->
<!--   ".RDS", -->
<!--   sep = "" -->
<!-- )) -->

<!-- m2_pr <- readRDS(paste(data_path, cohort_names[current_cohort], -->
<!--   "/dchp_pr_matrix_", cohort_names[current_cohort], -->
<!--   ".RDS", -->
<!--   sep = "" -->
<!-- )) -->

<!-- m3_pr <- readRDS(paste(data_path, cohort_names[current_cohort], -->
<!--   "/cmmhp_pr_matrix_", cohort_names[current_cohort], -->
<!--   ".RDS", -->
<!--   sep = "" -->
<!-- )) -->

<!-- indep_pr <- readRDS(paste(data_path, cohort_names[current_cohort], -->
<!--   "/immhp_pr_matrix_", cohort_names[current_cohort], -->
<!--   ".RDS", -->
<!--   sep = "" -->
<!-- )) -->

<!-- matrix_lst_plot <- list( -->
<!--   m1_pr[ -->
<!--     rev(expert_rank_10[[current_cohort]]), -->
<!--     expert_rank_10[[current_cohort]] -->
<!--   ], -->
<!--   m2_pr[ -->
<!--     rev(expert_rank_10[[current_cohort]]), -->
<!--     expert_rank_10[[current_cohort]] -->
<!--   ] -->
<!--   , -->
<!--   m3_pr[ -->
<!--     rev(expert_rank_10[[current_cohort]]), -->
<!--     expert_rank_10[[current_cohort]] -->
<!--   ], -->
<!--   indep_pr[ -->
<!--     rev(expert_rank_10[[current_cohort]]), -->
<!--     expert_rank_10[[current_cohort]] -->
<!--   ] -->
<!-- ) -->
<!-- # cat("Cohort ",current_cohort,"\n",sep="") -->
<!-- myMultiMatrixPlot( -->
<!--   X = matrix_lst_plot, no_matrix = 4, n_row = 2, -->
<!--   xLabels = expert_rank_10[[current_cohort]], -->
<!--   yLabels = rev(expert_rank_10[[current_cohort]]), -->
<!--   min = -80, max = 80, axis_cex = 2, title_cex = 1.8, -->
<!--   colorPalette = "RdBu", if.emp = FALSE, -->
<!--   # legend.mar=c(0.5,.5,0.5,0.5), -->
<!--   title_lst = list("C-HP", "C-DCHP", "C-MMHP", "I-MMHP"), -->
<!--   # title_lst = list("I-MMHP"), -->
<!--   col_axis = c(-80, -40, 0, 40, 80), -->
<!--   fake_matrix = FALSE, -->
<!--   matrix.mar = c(2.5, 2.5, 2.5, 1) -->
<!-- ) -->
<!-- mtext(text = "(b)", -->
<!--       side = 3, -->
<!--       line = 0.5, -->
<!--       outer = TRUE, -->
<!--       cex = 2.5, -->
<!--       font = 2) -->

<!-- ``` -->

<!-- ```{r fig-7a} -->
<!-- load("../output/revisions/lapl_check/prior_check/plot_N_predict_revision.RData") -->
<!-- current_cohort <- 5 # cohort 1 looks worse than rest -->

<!-- temp_plot_df <- predict_day_mae_df_lst[[current_cohort]] -->
<!-- temp_plot_df$method <- factor(temp_plot_df$method, -->
<!--   levels = c("mmhp", "dsnl", "m1", "m2", "m3"), -->
<!--   labels = c("I-MMHP", "DSNL", "C-HP", "C-DCHP", "C-MMHP") -->
<!-- ) -->
<!-- temp_plot_df$day <- factor(temp_plot_df$day) -->
<!-- # png(paste(plot_path,"real_predict_N_one_cohort.png",sep=""), -->
<!-- # width=600,height=300) -->

<!-- plot_cols_pred <- col_df %>% -->
<!--   filter(method %in% levels(temp_plot_df$method)) %>% -->
<!--   arrange(factor(method, levels = c( -->
<!--     "I-MMHP", "DSNL", -->
<!--     "C-HP", "C-DCHP", "C-MMHP" -->
<!--   ))) %>% -->
<!--   pull(cols) -->

<!-- p3 <- temp_plot_df %>% -->
<!--   rowwise() %>% -->
<!--   # mutate(norm = min(norm,60)) %>% -->
<!--   # rename(Method = method) %>% -->
<!--   ggplot(aes(x = method, y = mae, fill = method)) + -->
<!--   geom_boxplot(alpha = 0.7) + -->
<!--   scale_fill_manual(values = plot_cols_pred) + -->
<!--   ggtitle("(a)") + -->
<!--   theme_bw() + -->
<!--   theme( -->
<!--     plot.title = element_text(size = 22, hjust = 0.5, face = "bold"), -->
<!--     text = element_text(size = 20), -->
<!--     axis.text.x = element_blank(), -->
<!--     axis.ticks.x = element_blank(), -->
<!--     legend.position = "right" -->
<!--   ) + -->
<!--   facet_grid(cols = vars(day)) + -->
<!--   xlab("Prediction horizon (day)") + -->
<!--   ylab("MAE") + -->
<!--   ylim(0, 5) + # aren't infinite but >150 and <200 -->
<!--   NULL -->
<!-- p3 + labs(fill = "Method", title = "Single Cohort") -->

<!-- ``` -->

<!-- ```{r fig-7b} -->
<!-- load("../output/revisions/lapl_check/prior_check/plot_N_predict_revision.RData") -->

<!-- all_cohort_box <- all_cohort_mae_df -->

<!-- all_cohort_box$method <- factor(all_cohort_box$method, -->
<!--   levels = c("mmhp", "dsnl", "m1", "m2", "m3"), -->
<!--   labels = c( -->
<!--     "I-MMHP", "DSNL", -->
<!--     "C-HP", "C-DCHP", -->
<!--     "C-MMHP" -->
<!--   ) -->
<!-- ) -->
<!-- all_cohort_box$day <- factor(all_cohort_box$day) -->

<!-- plot_cols_pred <- col_df %>% -->
<!--   filter(method %in% levels(all_cohort_box$method)) %>% -->
<!--   arrange(factor(method, levels = c( -->
<!--     "I-MMHP", "DSNL", -->
<!--     "C-HP", "C-DCHP", "C-MMHP" -->
<!--   ))) %>% -->
<!--   pull(cols) -->

<!-- all_cohort_box %>% -->
<!--   ggplot(aes(x = method, y = norm, fill = method)) + -->
<!--   # facet_grid(cols = vars(day)) + -->
<!--   geom_boxplot(alpha = 0.7) + -->
<!--   scale_fill_manual(values = plot_cols_pred) + -->
<!--   ggtitle("(a)") + -->
<!--   theme_bw() + -->
<!--   theme( -->
<!--     plot.title = element_text( -->
<!--       size = 22, -->
<!--       hjust = 0.5, face = "bold" -->
<!--     ), -->
<!--     text = element_text(size = 20), -->
<!--     axis.text.x = element_blank(), -->
<!--     axis.ticks.x = element_blank(), -->
<!--     legend.position = "right" -->
<!--   ) + -->
<!--   facet_grid(cols = vars(day)) + -->
<!--   # guides(color=guide_legend(title="Method")) + -->
<!--   xlab("Prediction horizon (day)") + -->
<!--   ylab("MAE") + -->
<!--   labs(fill = "Method") + -->
<!--   # ylim(0, 80) + -->
<!--   NULL -->

<!-- ``` -->

<!-- ```{r fig-7b-pred-rank-spearman-all-cohorts} -->
<!-- load("../output/revisions/lapl_check/prior_check/plot_rank_predict.RData") -->

<!-- all_cohort_rank_df$method <- factor(all_cohort_rank_df$method, -->
<!--   levels = c("mmhp", "dsnl", "m1", "m2", "m3"), -->
<!--   labels = c( -->
<!--     "I-MMHP", "DSNL", "C-HP", -->
<!--     "C-DCHP", "C-MMHP" -->
<!--   ) -->
<!-- ) -->

<!-- p7 <- ggplot( -->
<!--   all_cohort_rank_df, -->
<!--   aes(x = method, y = spearman, fill = method) -->
<!-- ) + -->
<!--   # geom_line(size=1.2, alpha=0.8, position=position_dodge(0.2)) + -->
<!--   geom_boxplot(alpha = 0.7) + -->
<!--   scale_fill_manual(values = plot_cols_pred) + -->
<!--   ggtitle("(b)") + -->
<!--   theme_bw() + -->
<!--   theme( -->
<!--     plot.title = element_text(size = 22, hjust = 0.5, face = "bold"), -->
<!--     text = element_text(size = 20), -->
<!--     axis.text.x = element_blank(), -->
<!--     axis.ticks.x = element_blank(), -->
<!--     legend.position = "right" -->
<!--   ) + -->
<!--   facet_grid(cols = vars(day)) + -->
<!--   xlab("Prediction horizon (day)") + -->
<!--   ylab("Spearman rank corr.") + -->
<!--   # coord_cartesian(ylim=c(0,1)) + -->
<!--   labs(fill = "Method") -->

<!-- p7 -->
<!-- ``` -->

<!-- ```{r construct data for plot, results='hide',message=FALSE,cache=TRUE} -->
<!-- library(PlayerRatings) -->

<!-- current_cohort <- 7 -->
<!-- # 4 best so far, or 7 -->

<!-- load(paste(data_path, cohort_names[current_cohort], -->
<!--   "/predict_simulation_", cohort_names[current_cohort], -->
<!--   ".RData", -->
<!--   sep = "" -->
<!-- )) -->
<!-- clean_data <- cleanData(full_data[[cohort_names[current_cohort]]], -->
<!--   cut_off = 1 -->
<!-- ) -->

<!-- df <- full_data[[cohort_names[current_cohort]]] -->
<!-- df <- df[df$Actor != "Start" & df$Actor != "End", ] -->
<!-- df1 <- expandrows(df) -->
<!-- df1 <- df1[order(df1$Timestamp), ] # ensure in date order -->
<!-- df1$event <- 1:nrow(df1) -->
<!-- glick.df <- df1[, c("event", "Actor", "Recipient", "score"), with = FALSE] -->
<!-- # need event, actor, recipient, score -->
<!-- gl <- my_glicko(glick.df, history = TRUE, cval = 2) -->

<!-- # ------------------- period -->
<!-- training_end <- nrow(df[df$day <= 18, ]) -->
<!-- testing_end <- nrow(df) - 1 # 1077 #929 for cohort 5 -->
<!-- newx <- c(training_end:testing_end) -->

<!-- # ----------------- set pars for plot -->
<!-- my_col <- c("black", "grey", "orange", "red") -->
<!-- col_vec <- colorRampPalette(my_col)(mice_number) -->
<!-- ltypes_vec <- rep(1, mice_number) # c(1,2,3,1,2,3,1,2,3,1,2,3) -->
<!-- history_width <- 3 -->
<!-- sim_width <- 0.3 -->
<!-- shade_vec <- makeTransparent(col_vec, 0.2) -->
<!-- less_shade_vec <- makeTransparent(col_vec, 0.7) -->

<!-- # ---------------- simulation mean and sd -->
<!-- sim_gl_array <- array(NA, dim = c(12, testing_end - training_end + 1, 1000)) -->
<!-- for (s in 1:1000) { -->
<!--   print(s) -->
<!--   temp_gl <- glickoScoreSimulationWithWindowsReturnAll( -->
<!--     predict.sim = m3_predict_sim[5:8, s], # 7 for c 2 and 10 -->
<!--     history.df = glick.df[1:training_end, ] -->
<!--   ) -->
<!--   history_matrix <- temp_gl$history[, , 1] -->

<!--   if (testing_end > ncol(temp_gl$history[, , 1])) { -->
<!--     history_matrix <- cbind( -->
<!--       history_matrix, -->
<!--       t(matrix(rep( -->
<!--         history_matrix[, ncol(history_matrix)], -->
<!--         testing_end - ncol(temp_gl$history[, , 1]) -->
<!--       ), -->
<!--       nrow = testing_end - ncol(temp_gl$history[, , 1]), -->
<!--       byrow = TRUE -->
<!--       )) -->
<!--     ) -->
<!--   } else { -->
<!--     # try( -->
<!--     history_matrix <- history_matrix[, 1:testing_end]  -->
<!--     # temp_gl$history[,1:testing_end] -->
<!--     # ) -->
<!--   } -->
<!--   sim_gl_array[, , s] <- history_matrix[, training_end:testing_end] -->
<!-- } -->
<!-- ``` -->

<!-- ```{r fig-10a-glicko-rating} -->
<!-- par(mar = c(3, 5, 3, 1)) -->
<!-- testing_end <- testing_end - 1 -->
<!-- plot(1, -->
<!--   type = "n", xlab = "", -->
<!--   ylab = "Glicko ranking", xlim = c(0, dim(gl$history)[2]), -->
<!--   ylim = c(min(gl$history[, , 1]), max(gl$history[, , 1])), -->
<!--   cex.axis = 2, cex.lab = 2 -->
<!-- ) -->
<!-- for (m in c(1:mice_number)) { -->
<!--   # history -->
<!--   lines(gl$history[m, 1:training_end, 1], -->
<!--     cex = 0.6, col = col_vec[m], lwd = history_width, lty = ltypes_vec[m] -->
<!--   ) -->
<!--   lines(c(rep(NA, training_end), -->
<!--           gl$history[m, (training_end + 1):testing_end, 1]), -->
<!--     cex = 0.6, col = shade_vec[m], lwd = history_width, lty = ltypes_vec[m] -->
<!--   ) -->

<!--   # simulation -->
<!--   cur_end <- testing_end -->
<!--   mean_vec <- apply(sim_gl_array[m, , ], 1, function(x) mean(na.omit(x))) -->
<!--   sd_vec <- apply(sim_gl_array[m, , ], 1, function(x) sd(na.omit(x))) -->
<!--   polygon(c(newx, rev(newx)), -->
<!--           c(mean_vec - sd_vec, rev(mean_vec + sd_vec)), -->
<!--           col = shade_vec[m], border = NA) -->
<!--   # polygon(c(newx), c(mean_vec-sd_vec),col = shade_vec[m], border = NA) -->
<!--   lines(c(rep(NA, training_end), mean_vec), -->
<!--     cex = 0.6, col = col_vec[m], lwd = 1, lty = 2 -->
<!--   ) -->
<!-- } -->

<!-- # annotate the days -->
<!-- for (d in c(19:21)) { -->
<!--   abline(v = nrow(df[df$day <= (d - 1), ]), lty = 2, lwd = 0.4) -->
<!--   text(x = nrow(df[df$day <= (d - 1), ]), y = 2600, d) -->
<!-- } -->
<!-- mtext(text = "(a)", side = 3, line = -2.3, -->
<!--       outer = TRUE, cex = 2.8, at = 0.5, font = 2) -->
<!-- ``` -->

<!-- ```{r remove some objects, include=FALSE} -->
<!-- rm(m3_predict_sim) -->
<!-- rm(m2_predict_sim) -->
<!-- rm(m1_predict_sim) -->
<!-- rm(mmhp_predict_sim) -->
<!-- rm(fit_agg_rank) -->
<!-- rm(fit_cohort_hp) -->
<!-- rm(fit_cohort_dchp) -->
<!-- rm(fit_cohort_mmhp) -->

<!-- # also removing stan fits as they can cause issues in the next chunk -->
<!-- ``` -->