### 2021-08-13 output_scripts/Cohort_level_analysis.Rmd
first updates added for the first cohort

---
title: "Cohort Level Analysis"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output:
  html_document:
    # toc: true
    # toc_depth: 1
    number_sections: true
    # toc_float: true
    code_folding: hide
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 5, fig.width = 7,
  warning = FALSE, cache = TRUE
)

slides_plots <- FALSE
extra_plots <- FALSE

revision_plots <- TRUE
original_plots <- FALSE
### original plots for plots which are no longer needed, 
### revision plots are what required

# current_cohort <- 5

# don't write to this path
data_path <- "../output/revisions/lapl_check/prior_check/"
no_eta_3_data_path <- "../output/revisions/lapl_check/ident_check/"
no_eta_1_data_path <- "../output/revisions/lapl_check/ident_check_no_eta1/"

func_dir <- "../lib/"

library(here)
# library(rstan)
library(cmdstanr)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(posterior)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

cohort_names <- paste("cohort", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                      sep = "")
cohort_short_names <- paste("C", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                            sep = "")
cut_off <- 3
mice_number <- 12
col_df <- tibble(
  method = c(
    "I&SI", "AggRank", "DSNL", "Glicko", "I-MMHP",
    "C-HP", "C-DCHP", "True", "C-MMHP"
  ),
  cols = viridis(9)
)
theme_set(theme_minimal())
```

```{r load rankings, echo=FALSE, results='hide', cache=TRUE}
full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for (curr_cohort in fit_cohorts) {
  naive_rank_10[[curr_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})
```

```{r define_functions}
model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 +
    exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)
```

```{r define_plot_functions}
print_total_events <- function(current_cohort) {
  cat(paste("Cohort ", current_cohort, "\n", sep = ""))
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  stan_input_lst <- prepareDataStan(current_cohort)

  pair_events <- apply(stan_input_lst$Nm, 1, sum)

  pair_matrix <- matrix(0, nrow = 12, ncol = 12)
  for(i in seq_along(pair_events)) {
    i_ind <- stan_input_lst$I_fit[i]
    j_ind <- stan_input_lst$J_fit[i]
    pair_matrix[i_ind, j_ind] <- pair_events[i]
  }
  # pair_matrix <- clean_data$N_count[expert_rank_10[[current_cohort]],
  #                                   expert_rank_10[[current_cohort]]]
  pair_matrix <- pair_matrix[expert_rank_10[[current_cohort]],
                    expert_rank_10[[current_cohort]]]
  colnames(pair_matrix) <- expert_rank_10[[current_cohort]]
  rownames(pair_matrix) <- expert_rank_10[[current_cohort]]
  print(pair_matrix)
}

total_events_active <- function(current_cohort) {
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(current_cohort,
                                      raw_df =
                     full_data[[cohort_names[current_cohort]]],
                                      clean_data
  )
  unique_pairs_df <- return_df %>%
    group_by(initiator, recipient) %>%
    dplyr::summarize(
      count = n(),
      observe = list(observe.id),
      observe.length = list(observe.time),
      no.events = list(no.events)
    )
  unique_observe_win <- unique(return_df[,
                                         c("observe.id", "observe.time")])
  load(paste(data_path, cohort_names[current_cohort],
             "/cmmhp_est_zt_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  total_event_array <- array(0,
                             dim = c(mice_number, mice_number,
                                     max(return_df$observe.id)))
  active_event_array <- array(0,
                              dim = c(mice_number, mice_number,
                                      max(return_df$observe.id)))

  for (i in 1:mice_number) {
    for (j in 1:mice_number) {
      if(i != j) {
        pair <- which(unique_pairs_df$initiator == i &
                        unique_pairs_df$recipient == j)
        if (length(pair) > 0) {
          current_window_vec <- unique_pairs_df$observe[[pair]]
          for (cur_win in current_window_vec) {
            row_indicator <- (return_df$initiator == i) &
              (return_df$recipient == j) &
              (return_df$observe.id == cur_win)
            total_event_array[i, j, cur_win] <-
              length(return_df[
                row_indicator,"event.times"][[1]])
            active_event_array[i, j, cur_win] <- sum(apply(2 -
                        state_array_list[[pair]][[cur_win]],
                                                           1, mean) > 0.5)
          }
        }
      }
    }
  }

  total_matrix <- apply(total_event_array, c(1,2), sum)
  active_matrix <- apply(active_event_array, c(1,2), sum)
  rm(interpolation_array_list)
  rm(initial_state_list)
  rm(termination_state_list)
  rm(state_array_list)
  return(active_matrix)
}

###
print_total_less_windows <- function(current_cohort) {
  cat(paste("Cohort ", current_cohort, "\n", sep = ""))
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  stan_input_lst <- prepareDataStan(current_cohort)

  pair_events <- apply(stan_input_lst$Nm, 1, sum) -
    rowSums(stan_input_lst$Nm != 0)

  pair_matrix <- matrix(0, nrow = 12, ncol = 12)
  for(i in seq_along(pair_events)) {
    i_ind <- stan_input_lst$I_fit[i]
    j_ind <- stan_input_lst$J_fit[i]
    pair_matrix[i_ind, j_ind] <- pair_events[i]
  }
  # pair_matrix <- clean_data$N_count[expert_rank_10[[current_cohort]],
  #                                   expert_rank_10[[current_cohort]]]
  pair_matrix <- pair_matrix[expert_rank_10[[current_cohort]],
                    expert_rank_10[[current_cohort]]]
  colnames(pair_matrix) <- expert_rank_10[[current_cohort]]
  rownames(pair_matrix) <- expert_rank_10[[current_cohort]]
  print(pair_matrix)
}

rankings_by_total <- function(current_cohort) {
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  load(paste(data_path,
             cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",
             cohort_names[current_cohort],
             ".RData",
             sep = ""
             ))
  load(paste(data_path, cohort_names[current_cohort],
    "/agg_rank_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_hp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_dchp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count)

  no_method <- 4
  plot_latent_df <- data.frame(
    Methods = rep(1:no_method, each = 1000 * mice_number),
    Mice = rep(rep(1:mice_number, each = 1000), no_method),
    force = rep(0, 1000 * mice_number * no_method)
  )
  ## agg
  x_draws <- sim_agg_rank %>% select(starts_with("x"))
  plot_latent_df$force[(1:(1000 * mice_number))] <-
    as.vector(as.matrix(x_draws))

  ## M1
  f_draws <- sim_cohort_hp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         1000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M2
  f_draws <- sim_cohort_dchp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         2000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M3
  f_draws <- sim_cohort_mmhp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         3000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  plot_latent_df$Mice <- factor(plot_latent_df$Mice,
    levels = order(total_inter, decreasing = TRUE)
  )
  plot_latent_df$Method <- factor(plot_latent_df$Method,
    labels = c("AggRanking", "C-HP", "C-DCHP", "C-MMHP")
  )
  plot_cols <- col_df %>%
  filter(method %in% c(
    "AggRank", "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

  p1 <- ggplot(plot_latent_df, aes(x = Mice, y = force, fill = Method)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = plot_cols) +
    ggtitle("(a)") +
    theme_bw() +
    theme(
      plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
      text = element_text(size = 20),
      axis.text.x = element_text(size = 10),
      legend.position = "none"
    ) +
    facet_grid(cols = vars(Method)) +
    xlab("Mice (ranked by total interactions)") +
    ylab("Latent rank") +
    labs(title = paste("Cohort ", current_cohort, sep = ""))
  print(p1)
}

rankings_by_isi <- function(current_cohort){
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  load(paste(data_path,
             cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",
             cohort_names[current_cohort],
             ".RData",
             sep = ""
             ))
  load(paste(data_path, cohort_names[current_cohort],
    "/agg_rank_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_hp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
    "/cohort_dchp_stan_result_", cohort_names[current_cohort],
    ".RData",
    sep = ""
  ))
  total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count)

  no_method <- 4
  plot_latent_df <- data.frame(
    Methods = rep(1:no_method, each = 1000 * mice_number),
    Mice = rep(rep(1:mice_number, each = 1000), no_method),
    force = rep(0, 1000 * mice_number * no_method)
  )
  ## agg
  x_draws <- sim_agg_rank %>% select(starts_with("x"))
  plot_latent_df$force[(1:(1000 * mice_number))] <-
    as.vector(as.matrix(x_draws))

  ## M1
  f_draws <- sim_cohort_hp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         1000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M2
  f_draws <- sim_cohort_dchp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         2000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M3
  f_draws <- sim_cohort_mmhp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         3000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  plot_latent_df$Mice <- factor(plot_latent_df$Mice,
    levels = expert_rank_10[[current_cohort]]
  )
  plot_latent_df$Method <- factor(plot_latent_df$Method,
    labels = c("AggRanking", "C-HP", "C-DCHP", "C-MMHP")
  )
  plot_cols <- col_df %>%
  filter(method %in% c(
    "AggRank", "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

  p1 <- ggplot(plot_latent_df, aes(x = Mice, y = force, fill = Method)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = plot_cols) +
    ggtitle("(a)") +
    theme_bw() +
    theme(
      plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
      text = element_text(size = 20),
      axis.text.x = element_text(size = 10),
      legend.position = "none"
    ) +
    facet_grid(cols = vars(Method)) +
    xlab("Mice (ranked by I&SI)") +
    ylab("Latent rank") +
    labs(title = paste("Cohort ", current_cohort, sep = ""))
  print(p1)
}

rankings_by_cmmhp <- function(current_cohort) {
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  load(paste(data_path,
             cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",
             cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
             "/agg_rank_stan_result_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
             "/cohort_hp_stan_result_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
             "/cohort_dchp_stan_result_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  total_inter <- rowSums(clean_data$N_count) + colSums(clean_data$N_count)

  no_method <- 4
  plot_latent_df <- data.frame(
    Methods = rep(1:no_method, each = 1000 * mice_number),
    Mice = rep(rep(1:mice_number, each = 1000), no_method),
    force = rep(0, 1000 * mice_number * no_method)
  )
  ## agg
  x_draws <- sim_agg_rank %>% select(starts_with("x"))
  plot_latent_df$force[(1:(1000 * mice_number))] <-
    as.vector(as.matrix(x_draws))

  ## M1
  f_draws <- sim_cohort_hp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         1000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M2
  f_draws <- sim_cohort_dchp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         2000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ## M3
  f_draws <- sim_cohort_mmhp %>% select(starts_with("f"))
  plot_latent_df$force[(1:(1000 * mice_number)) +
                         3000 * mice_number] <-
    as.vector(as.matrix(f_draws))

  ### get median from this
  cmmhp_rank <- f_draws %>% 
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(med_rank = median(value)) %>% 
    arrange(desc(med_rank)) %>% 
    pull(name)
  ## then just get the numbers
  m3_ranks <- str_extract(cmmhp_rank, pattern = "[:digit:]+")

  plot_latent_df$Mice <- factor(plot_latent_df$Mice,
                                levels = m3_ranks
  )
  plot_latent_df$Method <- factor(plot_latent_df$Method,
                                  labels = c("AggRanking",
                                             "C-HP",
                                             "C-DCHP",
                                             "C-MMHP")
  )
  plot_cols <- col_df %>%
    filter(method %in% c(
      "AggRank", "C-HP",
      "C-DCHP", "C-MMHP"
    )) %>%
    pull(cols)
  p1 <- ggplot(plot_latent_df, aes(x = Mice, y = force, fill = Method)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = plot_cols) +
    ggtitle("(a)") +
    theme_bw() +
    theme(
      plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
      text = element_text(size = 20),
      axis.text.x = element_text(size = 10),
      legend.position = "none"
    ) +
    facet_grid(cols = vars(Method)) +
    xlab("Mice (ranked by C-MMHP)") +
    ylab("Latent rank") +
    labs(title = paste("Cohort ", current_cohort, sep = ""))
  print(p1)
}

param_ests <- function(current_cohort){
  load(paste(data_path, cohort_names[current_cohort],
             "/cohort_hp_stan_result_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  load(paste(data_path, cohort_names[current_cohort],
             "/cohort_dchp_stan_result_", cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))

  load(paste(data_path,
             cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",
             cohort_names[current_cohort],
             ".RData",
             sep = ""
  ))
  plot_cols <- col_df %>%
    filter(method %in% c(
      "C-HP",
      "C-DCHP",
      "C-MMHP"
    )) %>%
    pull(cols)
  all_data <- sim_cohort_mmhp %>%
    select(starts_with("eta") | starts_with("w_lam") |
             starts_with("beta")) %>%
    select(! ends_with("delta")) %>% 
    pivot_longer(everything()) %>%
    mutate(Model = "C-MMHP") %>%
    bind_rows(sim_cohort_dchp %>%
                select(starts_with("eta") |
                         starts_with("w_lam") |
                         starts_with("beta")) %>%
                select(!ends_with("delta")) %>% 
                select(!ends_with("detla")) %>% 
                pivot_longer(everything()) %>%
                mutate(Model = "C-DCHP")) %>% 
    bind_rows(sim_cohort_hp %>%
                select(starts_with("eta") |
                         starts_with("w_lam") |
                         starts_with("beta")) %>%
                select(!ends_with("delta")) %>% 
                pivot_longer(everything()) %>%
                mutate(Model = "C-HP"))
  all_data$Model <- factor(all_data$Model,
                           levels = c("C-HP", "C-DCHP", "C-MMHP"))
  comp_plot <- ggplot(all_data, aes(name, value, fill = Model)) +
    geom_boxplot() +
    scale_fill_manual(values = plot_cols) +
    labs(title = paste("Cohort ", current_cohort, sep = ""),
         subtitle = "Estimated Model Parameters") +
    # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~name, scales = "free")
  print(comp_plot)
  ###
  ## same for gamma
  all_data <- sim_cohort_mmhp %>%
    select(starts_with("gamma")) %>%
    pivot_longer(everything()) %>%
    mutate(Model = "C-MMHP") %>%
    bind_rows(sim_cohort_dchp %>%
                select(starts_with("gamma")) %>%
                pivot_longer(everything()) %>%
                mutate(Model = "C-DCHP"))
  all_data$Model <- factor(all_data$Model,
                           levels = c("C-HP", "C-DCHP", "C-MMHP"))
  all_data$name <- factor(all_data$name, 
                          levels = paste0("gamma[",1:12,"]"))
  comp_plot <- ggplot(all_data, aes(name, value, fill = Model)) +
    geom_boxplot() +
    scale_fill_manual(values = plot_cols[2:3]) +
    labs(title = paste("Cohort ", current_cohort, sep = ""),
         subtitle = "Out Degree Parameters", y = "Out Degree") +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~name, scales = "free")
  print(comp_plot)
  ###
  all_data <- sim_cohort_mmhp %>%
    select(starts_with("zeta") ) %>%
    pivot_longer(everything()) %>%
    mutate(Model = "C-MMHP") %>%
    bind_rows(sim_cohort_dchp %>%
                select(starts_with("zeta")) %>%
                pivot_longer(everything()) %>%
                mutate(Model = "C-DCHP"))
  all_data$Model <- factor(all_data$Model,
                           levels = c("C-HP", "C-DCHP", "C-MMHP"))
  all_data$name <- factor(all_data$name, 
                          levels = paste0("zeta[",1:12,"]"))
  comp_plot <- ggplot(all_data, aes(name, value, fill = Model)) +
    geom_boxplot() +
    scale_fill_manual(values = plot_cols[2:3]) +
    labs(title = paste("Cohort ", current_cohort, sep = ""),
         subtitle = "In Degree Parameters", y = "In Degree") +
    facet_wrap(~name, scales = "free")
    # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  print(comp_plot)
}

```

# Results {.tabset .unnumbered #sec1}

## Cohort 1

```{r set_cohort}
current_cohort <- 1
```

### Total Events ordered by I&SI ranking

```{r total-events_1, message=FALSE}
print_total_events(current_cohort)
```

### Total Events less Number of Windows containing events ordered by I&SI

```{r total-events-less-active-windows, message=FALSE}
active_events <- total_events_active(current_cohort)
colnames(active_events) <- expert_rank_10[[current_cohort]]
rownames(active_events) <- expert_rank_10[[current_cohort]]
print(active_events)
```

### Correlation Matrices

### Inferred ranking ordered by I&SI score

```{r ranked-by-isi}
rankings_by_isi(current_cohort)
```

### Inferred ranking order by C-MMHP median ranking

```{r ranked-by-cmmhp}
rankings_by_cmmhp(current_cohort)
```

### Parameter Estimates

```{r param-ests, message=FALSE}
param_ests(current_cohort)
```

<!-- ## Cohort 2 -->

<!-- ```{r set_cohort-2} -->
<!-- current_cohort <- 2 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_2, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c2, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c2, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c2} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ## C 3  -->

<!-- ```{r set_cohort-3} -->
<!-- current_cohort <- 3 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_3, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c3, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c3, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c3} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c3, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c3, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 4 -->

<!-- ```{r set_cohort-4} -->
<!-- current_cohort <- 4 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_4, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c4, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c4, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c4} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c4, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c4, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 5 -->

<!-- ```{r set_cohort-5} -->
<!-- current_cohort <- 5 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_5, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c5, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c5, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c5} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c5, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c5, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 6 -->

<!-- ```{r set_cohort-6} -->
<!-- current_cohort <- 6 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_6, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c6, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c6, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c6} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c6, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c6, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 7 -->

<!-- ```{r set_cohort-7} -->
<!-- current_cohort <- 7 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_7, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c7, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c7, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c7} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c7, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c7, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 8 -->

<!-- ```{r set_cohort-8} -->
<!-- current_cohort <- 8 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_8, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c8, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c8, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c8} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c8, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c8, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 9 -->

<!-- ```{r set_cohort-9} -->
<!-- current_cohort <- 9 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_9, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c9, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c9, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c9} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c9, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c9, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->

<!-- ## C 10 -->

<!-- ```{r set_cohort-10} -->
<!-- current_cohort <- 10 -->
<!-- ``` -->

<!-- ### Total Events ordered by I&SI ranking -->

<!-- ```{r total-events_10, message=FALSE} -->
<!-- print_total_events(current_cohort) -->
<!-- ``` -->

<!-- ### Total Events less Number of Windows containing events ordered by I&SI -->

<!-- ```{r total-events-less-active-windows-c10, message=FALSE} -->
<!-- print_total_less_windows(current_cohort) -->
<!-- ``` -->

<!-- ### Inferred ranking, ordered by total interactions -->

<!-- ```{r all-4-models-c10, warning=FALSE} -->
<!-- rankings_by_total(current_cohort) -->
<!-- ``` -->

<!-- <!-- Similarly, here are the corresponding plots with the animals ranked by --> 
<!-- <!-- I&SI score instead. -->

<!-- ### Inferred ranking ordered by I&SI score -->

<!-- ```{r ranked-by-isi-c10} -->
<!-- rankings_by_isi(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_3 -->

<!-- ```{r plot-rankings-both-models-no-eta3-c10, message=FALSE} -->
<!-- ests_without_eta_3(current_cohort) -->
<!-- ``` -->

<!-- ### Compare parameter estimates with/without eta_1 -->

<!-- ```{r plot-rankings-both-models-no-eta-1-c10, message=FALSE} -->
<!-- ests_without_eta_1(current_cohort) -->
<!-- ``` -->