### 2021-08-18 output_scripts/investigate_rankings.Rmd

---
title: "Ranking Stability"
author: "Owen G. Ward"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 5, fig.width = 7,
  warning = FALSE, cache = TRUE, eval_paper = FALSE
)

paper_plots <- FALSE
# extra_plots <- FALSE

current_cohort <- 5

# don't write to this path
data_path <- "../output/revisions/lapl_check/prior_check/"

func_dir <- "../lib/"

library(here)
# library(rstan)
library(cmdstanr)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(posterior)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

cohort_names <- paste("cohort", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                      sep = "")
cohort_short_names <- paste("C", c(9, 10, 12, 15, 16, 17, 18, 37, 38, 45),
                            sep = "")
cut_off <- 3
mice_number <- 12
col_df <- tibble(
  method = c(
    "I&SI", "AggRank", "DSNL", "Glicko", "I-MMHP",
    "C-HP", "C-DCHP", "True", "C-MMHP"
  ),
  cols = viridis(9)
)
theme_set(theme_minimal())
```

```{r load rankings, echo=FALSE, include=FALSE, results='hide', cache=TRUE}
full_data <- readRDS("../data/mice.RData")
# A=c9, B=c10, C=c12, D=c15, E=c16, F=c17, G=c18, H=c37, I=c38. J=c45

# Define the cohorts will be fitted
fit_cohorts <- c(1:10)
naive_rank_10 <- list()
expert_rank_10 <- list()
out <- captureOutput(for (curr_cohort in fit_cohorts) {
  naive_rank_10[[curr_cohort]] <-
    naiveRankHierarchy(full_data[[cohort_names[curr_cohort]]])
  expert_rank_10[[curr_cohort]] <-
    expertRankHierarchy(full_data[[cohort_names[curr_cohort]]])
})
```

```{r define_functions, include=FALSE}
model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 +
    exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)
```

# Other Paper plots

```{r fig-3b, eval=paper_plots}
current_cohort <- 5

m1_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/chp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

m2_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/dchp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

m3_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/cmmhp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

indep_pr <- readRDS(paste(data_path, cohort_names[current_cohort],
  "/immhp_pr_matrix_", cohort_names[current_cohort],
  ".RDS",
  sep = ""
))

matrix_lst_plot <- list(
  m1_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ],
  m2_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ]
  ,
  m3_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ],
  indep_pr[
    rev(expert_rank_10[[current_cohort]]),
    expert_rank_10[[current_cohort]]
  ]
)
# cat("Cohort ",current_cohort,"\n",sep="")
myMultiMatrixPlot(
  X = matrix_lst_plot, no_matrix = 4, n_row = 2,
  xLabels = expert_rank_10[[current_cohort]],
  yLabels = rev(expert_rank_10[[current_cohort]]),
  min = -80, max = 80, axis_cex = 2, title_cex = 1.8,
  colorPalette = "RdBu", if.emp = FALSE,
  # legend.mar=c(0.5,.5,0.5,0.5),
  title_lst = list("C-HP", "C-DCHP", "C-MMHP", "I-MMHP"),
  # title_lst = list("I-MMHP"),
  col_axis = c(-80, -40, 0, 40, 80),
  fake_matrix = FALSE,
  matrix.mar = c(2.5, 2.5, 2.5, 1)
)
mtext(text = "(b)",
      side = 3,
      line = 0.5,
      outer = TRUE,
      cex = 2.5,
      font = 2)

```

```{r fig-7a, eval=paper_plots}
load("../output/revisions/lapl_check/prior_check/plot_N_predict_revision.RData")
current_cohort <- 5 # cohort 1 looks worse than rest

temp_plot_df <- predict_day_mae_df_lst[[current_cohort]]
temp_plot_df$method <- factor(temp_plot_df$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c("I-MMHP", "DSNL", "C-HP", "C-DCHP", "C-MMHP")
)
temp_plot_df$day <- factor(temp_plot_df$day)
# png(paste(plot_path,"real_predict_N_one_cohort.png",sep=""),
# width=600,height=300)

plot_cols_pred <- col_df %>%
  filter(method %in% levels(temp_plot_df$method)) %>%
  arrange(factor(method, levels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP", "C-MMHP"
  ))) %>%
  pull(cols)

p3 <- temp_plot_df %>%
  rowwise() %>%
  # mutate(norm = min(norm,60)) %>%
  # rename(Method = method) %>%
  ggplot(aes(x = method, y = mae, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  xlab("Prediction horizon (day)") +
  ylab("MAE") +
  ylim(0, 5) + # aren't infinite but >150 and <200
  NULL
p3 + labs(fill = "Method", title = "Single Cohort")

```

```{r fig-7b, eval=paper_plots}
load("../output/revisions/lapl_check/prior_check/plot_N_predict_revision.RData")

all_cohort_box <- all_cohort_mae_df

all_cohort_box$method <- factor(all_cohort_box$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP",
    "C-MMHP"
  )
)
all_cohort_box$day <- factor(all_cohort_box$day)

plot_cols_pred <- col_df %>%
  filter(method %in% levels(all_cohort_box$method)) %>%
  arrange(factor(method, levels = c(
    "I-MMHP", "DSNL",
    "C-HP", "C-DCHP", "C-MMHP"
  ))) %>%
  pull(cols)

all_cohort_box %>%
  ggplot(aes(x = method, y = norm, fill = method)) +
  # facet_grid(cols = vars(day)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(a)") +
  theme_bw() +
  theme(
    plot.title = element_text(
      size = 22,
      hjust = 0.5, face = "bold"
    ),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  # guides(color=guide_legend(title="Method")) +
  xlab("Prediction horizon (day)") +
  ylab("MAE") +
  labs(fill = "Method") +
  # ylim(0, 80) +
  NULL

```

```{r fig-7b-pred-rank-spearman-all-cohorts, eval=paper_plots}
load("../output/revisions/lapl_check/prior_check/plot_rank_predict.RData")

all_cohort_rank_df$method <- factor(all_cohort_rank_df$method,
  levels = c("mmhp", "dsnl", "m1", "m2", "m3"),
  labels = c(
    "I-MMHP", "DSNL", "C-HP",
    "C-DCHP", "C-MMHP"
  )
)

p7 <- ggplot(
  all_cohort_rank_df,
  aes(x = method, y = spearman, fill = method)
) +
  # geom_line(size=1.2, alpha=0.8, position=position_dodge(0.2)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = plot_cols_pred) +
  ggtitle("(b)") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    text = element_text(size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "right"
  ) +
  facet_grid(cols = vars(day)) +
  xlab("Prediction horizon (day)") +
  ylab("Spearman rank corr.") +
  # coord_cartesian(ylim=c(0,1)) +
  labs(fill = "Method")

p7
```

```{r construct data for plot, results='hide',message=FALSE,cache=TRUE, eval=paper_plots}
library(PlayerRatings)

current_cohort <- 7
# 4 best so far, or 7

load(paste(data_path, cohort_names[current_cohort],
  "/predict_simulation_", cohort_names[current_cohort],
  ".RData",
  sep = ""
))
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]],
  cut_off = 1
)

df <- full_data[[cohort_names[current_cohort]]]
df <- df[df$Actor != "Start" & df$Actor != "End", ]
df1 <- expandrows(df)
df1 <- df1[order(df1$Timestamp), ] # ensure in date order
df1$event <- 1:nrow(df1)
glick.df <- df1[, c("event", "Actor", "Recipient", "score"), with = FALSE]
# need event, actor, recipient, score
gl <- my_glicko(glick.df, history = TRUE, cval = 2)

# ------------------- period
training_end <- nrow(df[df$day <= 18, ])
testing_end <- nrow(df) - 1 # 1077 #929 for cohort 5
newx <- c(training_end:testing_end)

# ----------------- set pars for plot
my_col <- c("black", "grey", "orange", "red")
col_vec <- colorRampPalette(my_col)(mice_number)
ltypes_vec <- rep(1, mice_number) # c(1,2,3,1,2,3,1,2,3,1,2,3)
history_width <- 3
sim_width <- 0.3
shade_vec <- makeTransparent(col_vec, 0.2)
less_shade_vec <- makeTransparent(col_vec, 0.7)

# ---------------- simulation mean and sd
sim_gl_array <- array(NA, dim = c(12, testing_end - training_end + 1, 1000))
for (s in 1:1000) {
  print(s)
  temp_gl <- glickoScoreSimulationWithWindowsReturnAll(
    predict.sim = m3_predict_sim[5:8, s], # 7 for c 2 and 10
    history.df = glick.df[1:training_end, ]
  )
  history_matrix <- temp_gl$history[, , 1]

  if (testing_end > ncol(temp_gl$history[, , 1])) {
    history_matrix <- cbind(
      history_matrix,
      t(matrix(rep(
        history_matrix[, ncol(history_matrix)],
        testing_end - ncol(temp_gl$history[, , 1])
      ),
      nrow = testing_end - ncol(temp_gl$history[, , 1]),
      byrow = TRUE
      ))
    )
  } else {
    # try(
    history_matrix <- history_matrix[, 1:testing_end]
    # temp_gl$history[,1:testing_end]
    # )
  }
  sim_gl_array[, , s] <- history_matrix[, training_end:testing_end]
}
```

```{r fig-10a-glicko-rating, eval=paper_plots}
par(mar = c(3, 5, 3, 1))
testing_end <- testing_end - 1
plot(1,
  type = "n", xlab = "",
  ylab = "Glicko ranking", xlim = c(0, dim(gl$history)[2]),
  ylim = c(min(gl$history[, , 1]), max(gl$history[, , 1])),
  cex.axis = 2, cex.lab = 2
)
for (m in c(1:mice_number)) {
  # history
  lines(gl$history[m, 1:training_end, 1],
    cex = 0.6, col = col_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )
  lines(c(rep(NA, training_end),
          gl$history[m, (training_end + 1):testing_end, 1]),
    cex = 0.6, col = shade_vec[m], lwd = history_width, lty = ltypes_vec[m]
  )

  # simulation
  cur_end <- testing_end
  mean_vec <- apply(sim_gl_array[m, , ], 1, function(x) mean(na.omit(x)))
  sd_vec <- apply(sim_gl_array[m, , ], 1, function(x) sd(na.omit(x)))
  polygon(c(newx, rev(newx)),
          c(mean_vec - sd_vec, rev(mean_vec + sd_vec)),
          col = shade_vec[m], border = NA)
  # polygon(c(newx), c(mean_vec-sd_vec),col = shade_vec[m], border = NA)
  lines(c(rep(NA, training_end), mean_vec),
    cex = 0.6, col = col_vec[m], lwd = 1, lty = 2
  )
}

# annotate the days
for (d in c(19:21)) {
  abline(v = nrow(df[df$day <= (d - 1), ]), lty = 2, lwd = 0.4)
  text(x = nrow(df[df$day <= (d - 1), ]), y = 2600, d)
}
mtext(text = "(a)", side = 3, line = -2.3,
      outer = TRUE, cex = 2.8, at = 0.5, font = 2)
```

```{r remove some objects, include=FALSE, eval=paper_plots}
rm(m3_predict_sim)
rm(m2_predict_sim)
rm(m1_predict_sim)
rm(mmhp_predict_sim)
rm(fit_agg_rank)
rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(fit_cohort_mmhp)

# also removing stan fits as they can cause issues in the next chunk
```

# Examine new simulation results

```{r inferred-rankings}
sim_data_path <- here("output",
                      "revisions",
                      "lapl_check",
                      "real_recov", "/")

sim_files <- list.files(sim_data_path, pattern = "sim_model3_fit")

num_nodes <- 12

cohort_est <- readRDS(here("output", "sim_ests.RDS"))

mean_values <- cohort_est %>% select(variable, mean)

f_vec <- mean_values %>%
  pivot_wider(names_from = variable, values_from = mean) %>%
  select(starts_with("f")) %>%
  as.numeric()

f_vec <- seq(from = 0.2, to = 0.9, length.out = num_nodes)

sim_ranks_mean <- tibble()
sim_ranks <- tibble()

model_colors <- col_df %>%
  filter(method %in% c(
    "C-HP",
    "C-DCHP", "C-MMHP"
  )) %>%
  pull(cols)

for(curr_sim in seq_along(sim_files)) {
  load(paste0(sim_data_path,
              sim_files[curr_sim]))

  f_m1 <- sim_model3_stan_sim1 %>% select(starts_with("f"))
  m1_ranks <- apply(f_m1, 2, median)

  f_m1_sims <- f_m1 %>%
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>%
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")
  f_m1_sims$model <- "C-HP"

  f_m2 <- sim_model3_stan_sim2 %>% select(starts_with("f"))
  m2_ranks <- apply(f_m2, 2, median)

  f_m2_sims <- f_m2 %>%
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>%
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m2_sims$model <- "C-DCHP"

  f_m3 <- sim_model3_stan_sim3 %>% select(starts_with("f"))
  m3_ranks <- apply(f_m3, 2, median)

  f_m3_sims <- f_m3 %>%
    rename_with(~gsub("f[", "", .x, fixed = TRUE)) %>%
    rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
    pivot_longer(everything(),
                 names_to = "node",
                 values_to = "rank")

  f_m3_sims$model <- "C-MMHP"

  curr_mean_ranks <- tibble(model = c(rep("C-HP", num_nodes),
                                 rep("C-DCHP", num_nodes),
                                 rep("C-MMHP", num_nodes)),
                       ranks = c(m1_ranks,
                                 m2_ranks,
                                 m3_ranks),
                       node = rep(1:num_nodes, 3),
                       true_rank = rep(f_vec, 3),
                       sim_id = curr_sim)
  sim_ranks_mean <- sim_ranks_mean %>%
    bind_rows(curr_mean_ranks)

  sim_ranks <- sim_ranks %>%
    bind_rows(f_m1_sims, f_m2_sims, f_m3_sims)

}

sim_ranks$node <- factor(sim_ranks$node, levels = 1:num_nodes)
sim_ranks_mean$node <- factor(sim_ranks_mean$node, levels = 1:num_nodes)

# sim_ranks$node

sim_ranks_mean$model <- factor(sim_ranks_mean$model,
                               levels = c( "C-HP", "C-DCHP", "C-MMHP"))

mean_sims <- sim_ranks_mean %>%
  left_join(col_df, by = c("model" = "method"))

mean_sims$model <- factor(mean_sims$model, levels = c("C-HP", "C-DCHP",
                                                      "C-MMHP"))

true_ranks <- tibble(node = rep(1:num_nodes,3),
                     true_rank = rep(f_vec,3),
                     model = c(rep("C-HP", num_nodes),
                               rep("C-DCHP", num_nodes),
                               rep("C-MMHP", num_nodes)))
true_ranks$node <- factor(true_ranks$node, levels = 1:num_nodes)

true_ranks$model <- factor(true_ranks$model,
                           levels = c("C-HP", "C-DCHP","C-MMHP"))

mean_sims %>%
  # filter(model == "C-MMHP") %>%
  ggplot(aes(node, ranks, fill = model)) +
  geom_boxplot(show.legend = FALSE) +
  scale_fill_manual(values = model_colors, guide = FALSE) +
  # geom_point(aes(node, true_rank, colour = "red")) +
  ylim(c(0,1)) +
  labs(title = "(a)",
       subtitle = "Mean rank from each simulation",
       y = "Posterior Mean Rank", x = "Node") +
  facet_wrap(~model) +
  geom_point(data = true_ranks,
             aes(node, true_rank, colour = "True Value")) +
  scale_shape_discrete(name = "True Value") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, size = 26),
        text = element_text(size = 18),
        legend.text = element_text(size = 18))

```

What about the recovery of the other parameters?

```{r eta_recovery}
eta_ests <- tibble()

for(curr_sim in seq_along(sim_files)) {
  load(paste0(sim_data_path,
              sim_files[curr_sim]))

  eta_m1 <- sim_model3_stan_sim1 %>%
    select(starts_with("eta") | starts_with("beta") |
             starts_with("r_lam")) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-HP") %>%
    filter(name != "beta_delta")

  eta_m2 <- sim_model3_stan_sim2 %>%
    select(starts_with("eta") | starts_with("beta") |
             starts_with("r_lam")) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-DCHP") %>%
    filter(name != "beta_delta")

  eta_m3 <- sim_model3_stan_sim3 %>%
    select(starts_with("eta") | starts_with("beta") |
             starts_with("r_lam")) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-MMHP") %>%
    filter(name != "beta_delta")

  curr_sim_ests <- bind_rows(eta_m1,
                             eta_m2,
                             eta_m3)
  eta_ests <- eta_ests %>%
    bind_rows(curr_sim_ests)
}

eta_ests$model <- factor(eta_ests$model, levels = c("C-HP",
                                                    "C-DCHP",
                                                    "C-MMHP"))

tru_values <- tibble(name = c("eta_1", "eta_2", "eta_3",
                              "r_lambda1", "beta"),
                     mean_est = c(object_par$sim_eta_1,
                                  object_par$sim_eta_2,
                                  object_par$sim_eta_3,
                                  object_par$sim_w_lam,
                                  object_par$sim_beta))

eta_ests %>%
  ggplot(aes(name, mean_est)) +
  geom_boxplot(aes(fill = model)) +
  geom_hline(mapping = aes(yintercept = mean_est),
             tru_values, colour = "red") +
  facet_wrap(~name, scales = "free")

```

Similarly, we can check the degree parameters.

```{r gamma-estimates}
gamma_ests <- tibble()

for(curr_sim in seq_along(sim_files)) {
  load(paste0(sim_data_path,
              sim_files[curr_sim]))

  gamm_m2 <- sim_model3_stan_sim2 %>%
    select(starts_with("gamma") ) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-DCHP")

  gamm_m3 <- sim_model3_stan_sim3 %>%
    select(starts_with("gamma") ) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-MMHP")

  curr_sim_ests <- bind_rows(gamm_m2,
                             gamm_m3)
  gamma_ests <- gamma_ests %>%
    bind_rows(curr_sim_ests)
}

gamma_ests$model <- factor(gamma_ests$model, levels = c("C-HP",
                                                    "C-DCHP",
                                                    "C-MMHP"))

tru_values <- tibble(name = paste0("gamma[", 1:num_nodes, "]"),
                     mean_est = object_par$gamma_var)

gamma_ests %>%
  ggplot(aes(name, mean_est)) +
  geom_boxplot(aes(fill = model)) +
  geom_hline(mapping = aes(yintercept = mean_est),
             tru_values, colour = "red") +
  facet_wrap(~name, scales = "free")
```

```{r zeta-estimates}
zeta_ests <- tibble()

for(curr_sim in seq_along(sim_files)) {
  load(paste0(sim_data_path,
              sim_files[curr_sim]))

  zeta_m2 <- sim_model3_stan_sim2 %>%
    select(starts_with("zeta") ) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-DCHP")

  zeta_m3 <- sim_model3_stan_sim3 %>%
    select(starts_with("zeta") ) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    summarise(mean_est = median(value)) %>%
    mutate(model = "C-MMHP")

  curr_sim_ests <- bind_rows(zeta_m2,
                             zeta_m3)
  zeta_ests <- zeta_ests %>%
    bind_rows(curr_sim_ests)
}

zeta_ests$model <- factor(zeta_ests$model, levels = c("C-HP",
                                                    "C-DCHP",
                                                    "C-MMHP"))

tru_values <- tibble(name = paste0("zeta[", 1:num_nodes, "]"),
                     mean_est = object_par$zeta_var)

zeta_ests %>%
  ggplot(aes(name, mean_est)) +
  geom_boxplot(aes(fill = model)) +
  geom_hline(mapping = aes(yintercept = mean_est),
             tru_values, colour = "red") +
  facet_wrap(~name, scales = "free")
```